### 2024-04-28 test.qmd
initial commit

---
title: "Quarto Presentations"
subtitle: "Create beautiful interactive slide decks with Reveal.js"
format:
  revealjs: 
    slide-number: true
    chalkboard: 
      buttons: false
    preview-links: auto
    logo: images/quarto.png
    footer: <http://seiroito.github.io>
    theme: sky
width: 1500
resources:
  - test.pdf
# quarto render test.qmd
---

## Hello, There

This presentation will show you examples of what you can do with Quarto and [Reveal.js](https://revealjs.com), including:

-   Presenting code and LaTeX equations
-   Including computations in slide output
-   Image, video, and iframe backgrounds
-   Fancy transitions and animations
-   Printing to PDF

...and much more

## Pretty Code {auto-animate="true"}

-   Over 20 syntax highlighting themes available
-   Default theme optimized for accessibility

::: footer
Learn more: [Syntax Highlighting](https://quarto.org/docs/output-formats/html-code.html#highlighting)
:::

## Line Highlighting

-   Highlight specific lines for emphasis
-   Incrementally highlight additional lines

## Executable Code

::: footer
Learn more: [Executable Code](https://quarto.org/docs/presentations/revealjs/#executable-code)
:::

## LaTeX Equations

[MathJax](https://www.mathjax.org/) rendering of equations to HTML

::: columns
::: {.column width="40%"}
``` tex
\begin{gather*}
a_1=b_1+c_1\\
a_2=b_2+c_2-d_2+e_2
\end{gather*}

\begin{align}
a_{11}& =b_{11}&
  a_{12}& =b_{12}\\
a_{21}& =b_{21}&
  a_{22}& =b_{22}+c_{22}
\end{align}
```
:::

::: {.column width="60%"}
```{=tex}
\begin{align}
\mbox{売却差損}&\equiv\mbox{購入価格}-\mbox{売却価格}\\
\mbox{債務残高}&\equiv\mbox{(1+r)融資額}-\mbox{返済額}\\
\end{align}
```
```{=tex}
\begin{align}
\mbox{売却価格}&\equiv\mbox{購入価格}-\mbox{売却差損}\\
\mbox{債務残高}-\mbox{売却価格}&\equiv\mbox{(1+r)融資額}-\mbox{返済額}-\mbox{購入価格}+\mbox{売却差損}\\
\end{align}
```
:::
:::

::: footer
Learn more: [LaTeX Equations](https://quarto.org/docs/authoring/markdown-basics.html#equations)
:::

----

::: {.incremental}
$$
\begin{align}
\mbox{売却差益}&\equiv\mbox{売却価格}-\mbox{購入価格}\\
\mbox{債務残高}&\equiv\mbox{(1+r)融資額}-\mbox{返済額}\\
\mbox{返済額}&\equiv\mbox{家賃収入}-\mbox{経費}-\mbox{利潤}\\
\end{align}
$$
:::
::: {.incremental}
$$
\begin{align}
\mbox{売却価格}&\equiv\mbox{購入価格}+\mbox{売却差益}\\
\mbox{債務残高}&\equiv\mbox{(1+r)融資額}-\mbox{家賃収入}+\mbox{経費}+\mbox{利潤}
\end{align}
$$
:::
::: {.incremental}
$$
\begin{align}
\mbox{債務残高}+\mbox{売却価格}&\equiv\mbox{(1+r)融資額}-\mbox{家賃収入}+\mbox{経費}+\mbox{利潤}\\
&\hspace{1em}+\mbox{購入価格}+\mbox{売却差益}\\
\mbox{利潤}+\mbox{売却差益}&\equiv
\mbox{売却価格}+\mbox{債務残高}+\mbox{家賃収入}-\mbox{経費}\\
&\hspace{1em}-\{\mbox{(1+r)融資額}+\mbox{購入価格}\}
\end{align}
$$
:::

----

$$
\begin{align}
\mbox{利潤}+\mbox{売却差益}&\equiv
\underbrace{\mbox{売却価格}+\mbox{債務残高}}_{timing}+\underbrace{\mbox{家賃収入}-\mbox{経費}}_{efforts}\\
&\hspace{1em}\underbrace{-\{\mbox{(1+r)融資額}+\mbox{購入価格}\}}_{fixed}
\end{align}
$$

::: {.incremental}
* timingは時間とともに減少傾向 &rarr; 売却差益  
* effortsは時間とともに増加傾向 &rarr; 利潤  
* 債務残高と売却価格は和(cash stock)で考える、差は考える必要なし  
* 融資額=.5購入価格、債務残高=.5融資額で売却、売却価格=.75購入価格だと、右辺=家賃収入-経費+\{(.75+.25)-(1.5+.5r)\}購入額=家賃収入-経費-(1+r).5購入額  
:::

----

::: {.incremental}
* 売却価格が下がるならば、利潤を高める必要がある  
* しかし、売却価格が下がりすぎるならば、利潤で相殺するのは難しい  
* 売却価格が下がりすぎない物件を探すのが重要  
* 不動産投資では、複数物件の売却差損リスクが強く正相関しているので、リスク分散ができない   
* リスク分散: 不動産とは価値が負相関の資産(国債? 他国資産?)を保有  
:::

### 2024-04-30 posts/UsingGitAndGithub/UsingGitAndGithub.qmd
Moved qmd files under the posts folder

---
title: "GitとGitHubを使う"
subtitle: "version controlとweb based group file-work"
format:
  revealjs: 
    include-in-header: 
      text: |
        <style>
        .center-y {
          margin: 0;
          position: absolute;
          top: 50%;
          -ms-transform: translateY(-50%);
          transform: translateY(-50%);
        }
        </style>
    slide-number: true
    chalkboard: 
      buttons: false
    preview-links: auto
    logo: images/quarto.png
    footer: <http://seiroito.github.io/Miscellaneous/QuartoFiles/UsingGitAndGithub.html>
    theme: sky
width: 1500
resources:
  - UsingGitAndGithub.pdf
# quarto render UsingGitAndGithub.qmd
---
::: {.center-y}

## Gitはmain(自分のPC)、GitHubはorigin(みんなの共有ファイル)

:::

----

### このファイルの前提

* GitHubにアカウントを作ってあります  
* [Git for windows](https://gitforwindows.org/)/Git  をインストールしてあります

----
Gitとは

::: {.incremental}
* ファイルのversion controlプログラム  
   * ファイル: A &rarr; A+B &rarr; A+B-C &rarr; A+...+Z、A+Bに戻りたい! A+B-Cて何だったっけ?  
   * Version control: 全改変記録を保持していつでも閲覧、回帰(とある版に戻ること)が可能  
:::
----
Gitとは

::: {.incremental}
* ファイルのdistributed version controlプログラム  
   * 同じプロジェクトにAさんとBさんが作業: A1 &rarr; A2、A1 &rarr; B2, A2にB2を反映させたい!  
   * Branch (作業の流れ): Aさんの作業の流れ(branch A)から分岐(fork)してBさんの作業の流れ(branch B)がある  
   * Bさん: B2を共有レポジトリにpushする
   * Aさん: B2を共有レポジトリから自分のPCにpullする=B2をベースにして作業する  
   * 差分=A2-A1部分、B2-A1部分: ファイルの違う箇所・機能を作業分担するのが原則  
      * そうしないと手作業で重複を解決しなくてはならず、相当面倒  
   * 複数人の作業内容を自動で合成できる  
:::

----
----
Gitとは

::: {.incremental}
* ファイルのdistributed version controlプログラム  
   * 単一のcentral file(s)を使わず、各自が部分的に貢献したものを共有ファイルとして再構成する  
   * 最新ファイルはサーバではなく、[各自のPCに保存]{style="color:red;"}  
      * Repositoryのあるサーバ管理会社はファイルを所有しない  
      * 所有者は参加者全員  
         * ブロックチェイン(分散型台帳システムdistributed ledger system)に似てます  
      * [リモートで作業]{style="color:red;"}ができる  
      * [独立して試行]{style="color:red;"}できる  
* [Gitの歴史](https://www.welcometothejungle.com/en/articles/btc-history-git)  
:::

----
GitHubとは

::: {.incremental}
* Gitを扱うwebサービス  
* レポジトリをwebに作成、分散版管理しながら共同作業ができる  
* Webページも公開できる  
* 個人アカウント作成、個人単位のファイル容量(上限100MB/ファイル、5GB/repo)なら無料

:::

----

[差は考える必要なし]{style="color:red;"}  
* 融資額=.5購入価格、債務残高=.5融資額で売却、売却価格=.75購入価格だと、右辺=家賃収入-経費+\{(.75+.25)-(1.5+.5r)\}購入額=家賃収入-経費-(1+r).5購入額  
:::

----
左辺を大きくしたい

::: {.incremental}
* 売却価格が下がるならば、利潤を高める必要がある  
* 売却価格が下がりすぎない物件を探すのが重要  
   * 一般に、資産価格は利子率に応じて「率」変化する(「額」変化ではない)  
     * 築浅、高額の物件ほど、上げ下げの幅は大きい  
   * 平時には、実需見通しが上方修正されれば、減少する程度が小さくなる  
   * パニック時には、売り手は速やかに現金を得ようとするため、実需にかかわらず価格が下がる  
* 不動産投資では、複数物件の売却差損リスクが強く正相関しているので、リスク分散ができない   
* リスク分散: 不動産とは価値が負相関の資産(国債? 他国資産?)を保有  
:::

### 2024-04-30 posts/UsingGitAndGithub/UsingGitAndGithub.qmd
Added who should use git in usinggitandgithub.html

@@
 ----
+
 Gitとは
@@
 * ファイルのversion controlプログラム  
-   * ファイル: A &rarr; A+B &rarr; A+B-C &rarr; A+...+Z、A+Bに戻りたい! A+B-Cて何だったっけ?  
-   * Version control: 全改変記録を保持していつでも閲覧、回帰(とある版に戻ること)が可能  
+   * ファイル: A &rarr; A+B &rarr; A+B-C &rarr; A+...+Z、A+Bに戻りたい! A+B-Cて何だっけ?  
+   * Version control: 全改変記録を保持、いつでも閲覧、回帰(とある版に戻ること)が可能  
 :::
+
 ----
+
 Gitとは
@@
    * Branch (作業の流れ): Aさんの作業の流れ(branch A)から分岐(fork)してBさんの作業の流れ(branch B)がある  
-   * Bさん: B2を共有レポジトリにpushする
-   * Aさん: B2を共有レポジトリから自分のPCにpullする=B2をベースにして作業する  
-   * 差分=A2-A1部分、B2-A1部分: ファイルの違う箇所・機能を作業分担するのが原則  
+   * Bさん: B2を自分のPCから共有レポジトリにpush
+   * Aさん: B2を共有レポジトリから自分のPCにpull=B2をベースにして作業  
+   * 差分=A2-A1、B2-A1: ファイルの違う箇所・機能を作業分担するのが原則  
       * そうしないと手作業で重複を解決しなくてはならず、相当面倒  
-   * 複数人の作業内容を自動で合成できる  
+   * 複数人の作業内容を自動で合成・再構成できる  
 :::
@@
 ----
-----
+
 Gitとは
@@
       * 所有者は参加者全員  
-         * ブロックチェイン(分散型台帳システムdistributed ledger system)に似てます  
+         * ブロックチェイン(分散型台帳システムdistributed ledger system)に類似  
       * [リモートで作業]{style="color:red;"}ができる  
       * [独立して試行]{style="color:red;"}できる  
-* [Gitの歴史](https://www.welcometothejungle.com/en/articles/btc-history-git)  
 :::
@@
 ----
+
+Gitとは
+
+::: {.incremental}
+* [Gitの歴史](https://www.welcometothejungle.com/en/articles/btc-history-git)  
+   * Linus Torvald氏はGit開発から一銭も得ていない  
+   * Linus Torvald氏はlinuxt開発から一銭も得ていない
+
+:::
+
+----
+
 GitHubとは
@@
 * Gitを扱うwebサービス  
-* レポジトリをwebに作成、分散版管理しながら共同作業ができる  
+* レポジトリをwebに作成、DVC(分散版管理)しながら共同作業ができる  
 * Webページも公開できる  
@@
 :::
+
+----
+
+Git for windows/は自分のPCでGitを使うプログラム
+
+* 自分のPCの版管理  
+* GitHubとのpush/pullのやりとり  
+* GitHub上でのbranch merge
+
+----
+
+だれがGit/GitHubを使うべきか
+
+* 自分のPCでプロジェクトを実施する人  
+* 頻繁に版更新する人  
+* 他者とプロジェクトを実施する人
 

### 2024-04-30 posts/UsingGitAndGithub/UsingGitAndGithub.qmd
Added who should use git in usinggitandgithub.html
Added target: gh-pages in jobs-steps-with of publish.yml

@@
 # quarto render UsingGitAndGithub.qmd
+# quarto publish gh-pages UsingGitAndGithub.qmd
 ---
@@
 
-## Gitはmain(自分のPC)、GitHubはorigin(みんなの共有ファイル)
+### Gitはmain(自分のPC)、GitHubはorigin(みんなの共有ファイル)
 

### 2024-04-30 posts/UsingGitAndGithub/UsingGitAndGithub.qmd
Centered text

--- 2024-04-30 posts/UsingGitAndGithub/UsingGitAndGithub.qmd
@@
 # quarto render UsingGitAndGithub.qmd
+# quarto publish gh-pages UsingGitAndGithub.qmd # this keeps on deploying and does not stop...
 ---
@@
 
-## Gitはmain(自分のPC)、GitHubはorigin(みんなの共有ファイル)
+### Gitはmain(自分のPC)、GitHubはorigin(みんなの共有ファイル)  {.center}
 

### 2024-06-26 posts/Watabe/CommentsOnRefereeResponse_WatabeYangKanasheuski.qmd
Added CommentsOnRefereeResponse_WatabeYangKanasheuski.html

----
title: "Comments on referee reports and responses"
subtitle: "by Watabe, Yang and Kanasheuski of June, 2024"
author: "Seiro Ito"
date: "`r format(Sys.time(), '%m %d, %Y% %R')`"
format:
  revealjs: 
    include-in-header: 
      text: |
        <style>
        .center-y {
          margin: 0;
          position: absolute;
          top: 50%;
          -ms-transform: translateY(-50%);
          transform: translateY(-50%);
        }
        </style>
    slide-number: true
    chalkboard: 
      buttons: false
    preview-links: auto
    logo: images/quarto.png
    footer: <http://seiroito.github.io/Miscellaneous/QuartoFiles/post/Watabe/CommentsOnRefereeResponse_WatabeYangKanasheuski.html>
    theme: sky
width: 1500
resources:
  - CommentsOnRefereeResponse_WatabeYangKanasheuski.pdf
# quarto render CommentsOnRefereeResponse_WatabeYangKanasheuski.qmd
----
```{r, load_refs, eval = F, include=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("./seiro.bib", check = FALSE)
```

# Editor

::: {.incremental}

I found the editor is helpful and constructive.

I only found comment 8 to be a little problematic.

:::

----
### 8. For R2, comment 2: you might interact the two countries real GDP per capita analogous to Hallak (2006).

You added $\ln GDPPC_{i}*\ln GDPPC_{j}$.

::: {.incremental}
* To me, fn 11 suffices. Table 9 is redundant.  
* As you write in fn 11, 
   * $GDPPC_{i}*GDPPC_{j}$ is collinear with FEs, so income levels are already controlled for, hence results are homothetic over incomes, in the earlier results.  
   * Estimation exploits nonlinearity of logs.  
      * This is why only OLS estimates are affected.  
      * Logged interaction term picks up mostly lower level income variations. Not sure if this is what we want.
   * If the editor really wants to add interaction term, which sounds like one, I would drop country FEs and add $GDPPC_{i}$, $*GDPPC_{j}$, $GDPPC_{i}*GDPPC_{j}$ as covariates.  
:::
----
# Ref 1

I only found a potential issue on Comment 1.

For all other comments, your response is straightforward.

----
### 1.  Given that the access to gaming consoles and PCs is far from exogenous, it would be nice to know whether the results may be affected by using this selected sample.

::: {.incremental}
Your response in short: It is exogenous or ignorably endogenous.

* You are being polite by saying "rightly pointed out" only to deny it later.  
   * This can confuse the ref.  
* You indicate (with reddit): Game seller's choice on platforms is not correlated with geography.  
   * Implies: Platform choice/sample selection is exogenous to/uncorrelated with $X_{in}$ (sales in destination $n$ of origin $i$).  
* Ref 1 does not explicitly state exogenous to "what".  
* On Stern sample, availability is correlated with distance only very weakly. It is actually a different question you are answering but is indicative and shows efforts of doing whatever possible.

The best is to get data of game console sales and show exogenous sample selection. But this is not available.

So you are responding well. 
:::

----
### 1.  Given that the access to gaming consoles and PCs is far from exogenous, it would be nice to know whether the results may be affected by using this selected sample.

Just in case you want to be extra sure:

::: {.incremental}
 * Sales in 75 countries (almost all mid-high income countries).  
 * My guess: Game console sales (my guess: more expensive per "enjoyment") may positively correlate with income levels?  
 * If so, studio HQ location (my guess: mostly in higher income countries) and game console sales in destination may have some geographical correlations, because higher income countries tend to sit next to each other. In lower income countries, more availability of Steam than console games? 
   * You can partly test this by plotting sales $X_{n}=\sum_{i}X_{in}$ and income levels.  
:::

----
# Ref 2

Nothing to add for Ref 2.

### 2024-06-26 posts/Watabe/CommentsOnRefereeResponse_WatabeYangKanasheuski.qmd
Corrected typos on CommentsOnRefereeResponse_WatabeYangKanasheuski.html

@@
-----
+---
 title: "Comments on referee reports and responses"
@@
 author: "Seiro Ito"
-date: "`r format(Sys.time(), '%m %d, %Y% %R')`"
 format:
@@
     logo: images/quarto.png
-    footer: <http://seiroito.github.io/Miscellaneous/QuartoFiles/post/Watabe/CommentsOnRefereeResponse_WatabeYangKanasheuski.html>
+    footer: <http://seiroito.github.io/Miscellaneous/QuartoFiles/posts/Watabe/CommentsOnRefereeResponse_WatabeYangKanasheuski.html>
     theme: sky
@@
 # quarto render CommentsOnRefereeResponse_WatabeYangKanasheuski.qmd
-----
-```{r, load_refs, eval = F, include=FALSE, cache=FALSE}
-library(RefManageR)
-BibOptions(check.entries = FALSE,
-           bib.style = "authoryear",
-           cite.style = "alphabetic",
-           style = "markdown",
-           hyperlink = FALSE,
-           dashed = FALSE)
-myBib <- ReadBib("./seiro.bib", check = FALSE)
-```
-
-# Editor
+---
+## Editor
 
@@
 
-I found the editor is helpful and constructive.
-
-I only found comment 8 to be a little problematic.
+* I found the editor is helpful and constructive.  
+* I only found comment 8 to be a little problematic.
 
@@
 ----
-### 8. For R2, comment 2: you might interact the two countries real GDP per capita analogous to Hallak (2006).
+
+#### For R2, comment 2: you might interact the two countries real GDP per capita analogous to Hallak (2006).
 
@@
 ::: {.incremental}
+
 * To me, fn 11 suffices. Table 9 is redundant.  
 * As you write in fn 11,  
-   * $GDPPC_{i}*GDPPC_{j}$ is collinear with FEs, so income levels are already controlled for, hence results are homothetic over incomes, in the earlier results.  
+  * $GDPPC_{i}-GDPPC_{j}$ is collinear with FEs, so income levels are already controlled for, hence results are homothetic over incomes, in the earlier results.
+
+:::
+
+----
+
+#### For R2, comment 2: you might interact the two countries real GDP per capita analogous to Hallak (2006).
+
+You added $\ln GDPPC_{i}*\ln GDPPC_{j}$.
+
+::: {.incremental}
+
+* As you write in fn 11,  
   * Estimation exploits nonlinearity of logs.  
@@
      * Logged interaction term picks up mostly lower level income variations. Not sure if this is what we want.
-   * If the editor really wants to add interaction term, which sounds like one, I would drop country FEs and add $GDPPC_{i}$, $*GDPPC_{j}$, $GDPPC_{i}*GDPPC_{j}$ as covariates.  
+  * If the editor really wants to add interaction term, which sounds like one, I would drop country FEs and add $GDPPC_{i}$, $GDPPC_{j}$, $GDPPC_{i}*GDPPC_{j}$ as covariates.
+
 :::
+
 ----
+
 # Ref 1
@@
 ----
-### 1.  Given that the access to gaming consoles and PCs is far from exogenous, it would be nice to know whether the results may be affected by using this selected sample.
+
+#### Given that the access to gaming consoles and PCs is far from exogenous, it would be nice to know whether the results may be affected by using this selected sample.
+
+Your response in short: It is exogenous or ignorably endogenous.
 
 ::: {.incremental}
+
+- You are being polite by saying "rightly pointed out" only to deny it later.  
+   - This can confuse the ref.  
+- You indicate (with reddit): Game seller's choice on platforms is not correlated with geography.  
+   - Implies: Platform choice/sample selection is exogenous to/uncorrelated with $X_{in}$ (sales in destination $n$ of origin $i$).  
+- Ref 1 does not explicitly state exogenous to "what".
+
+:::
+
+----
+
+#### Given that the access to gaming consoles and PCs is far from exogenous, it would be nice to know whether the results may be affected by using this selected sample.
+
+::: {.incremental}
+
 Your response in short: It is exogenous or ignorably endogenous.
 
-* You are being polite by saying "rightly pointed out" only to deny it later.  
-   * This can confuse the ref.  
-* You indicate (with reddit): Game seller's choice on platforms is not correlated with geography.  
-   * Implies: Platform choice/sample selection is exogenous to/uncorrelated with $X_{in}$ (sales in destination $n$ of origin $i$).  
-* Ref 1 does not explicitly state exogenous to "what".  
-* On Stern sample, availability is correlated with distance only very weakly. It is actually a different question you are answering but is indicative and shows efforts of doing whatever possible.
+- On Stern sample, availability is correlated with distance only very weakly. It is actually a different question you are answering but is indicative and shows efforts of doing whatever possible.
 
@@
 So you are responding well.
+
 :::
@@
 ----
-### 1.  Given that the access to gaming consoles and PCs is far from exogenous, it would be nice to know whether the results may be affected by using this selected sample.
+
+#### Given that the access to gaming consoles and PCs is far from exogenous, it would be nice to know whether the results may be affected by using this selected sample.
 
@@
 ::: {.incremental}
- * Sales in 75 countries (almost all mid-high income countries).  
- * My guess: Game console sales (my guess: more expensive per "enjoyment") may positively correlate with income levels?  
- * If so, studio HQ location (my guess: mostly in higher income countries) and game console sales in destination may have some geographical correlations, because higher income countries tend to sit next to each other.
+
+- Sales in 75 countries (almost all mid-high income countries).  
+- My guess: Game console sales (my guess: more expensive per "enjoyment") may positively correlate with income levels?  
+- If so, studio HQ location (my guess: mostly in higher income countries) and game console sales in destination may have some geographical correlations, because higher income countries tend to sit next to each other.
 In lower income countries, more availability of Steam than console games?
-   * You can partly test this by plotting sales $X_{n}=\sum_{i}X_{in}$ and income levels.  
+ - You can partly test this by plotting sales $X_{n}=\sum_{i}X_{in}$ and income levels.
+
 :::
 
-----
+---
+
 # Ref 2

### 2024-06-26 posts/Watabe/CommentsOnRefereeResponse_WatabeYangKanasheuski.qmd
Added typo photos

@@
   revealjs: 
-    include-in-header: 
-      text: |
-        <style>
-        .center-y {
-          margin: 0;
-          position: absolute;
-          top: 50%;
-          -ms-transform: translateY(-50%);
-          transform: translateY(-50%);
-        }
-        </style>
     slide-number: true
@@
 
-For all other comments, your response is straightforward.
+For all other comments, your response is straightforward and clean.
 
@@
 Nothing to add for Ref 2.
+
+---
+
+## Typos
+
+![](IMG20240626163539.jpg)
+
+---
+
+## Typos
+
+![](IMG20240626163619.jpg)
+
+---
+
+## Typos
+
+![](IMG20240626163552.jpg){height="600"}

### 2024-07-30 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
added BarrEggelstonSmith slides

---
title: "Investing in infants: the lasting effects of cash transfers to new families"
subtitle: "By Andrew Barr, Jonathan Eggleston, Alexander A. Smith"
author: "Seiro Ito"
format:
  revealjs: 
    slide-number: true
    chalkboard: 
      buttons: false
    preview-links: auto
    logo: images/quarto.png
    footer: <http://seiroito.github.io/Miscellaneous/QuartoFiles/posts/BarrEgglestonSmith/BarrEgglestonSmith2022.html>
    theme: sky
include-in-header:
  - text: |
      <style>
      #title-slide
       .title {
        font-size: 2.0em;
      }
      </style>
#width: 1000
bibliography: c:/seiro/settings/TeX/seiro.bib
resources:
  - BarrEgglestonSmith2022.pdf
# quarto render BarrEgglestonSmith2022.qmd
---
<style>
    :root{--r-main-font-size: 36px;}
    .reveal h1, .reveal h2, .reveal h3, .reveal h4, .reveal h5 {
                  text-transform: none;
          }
.spacer {
    margin-top: 20px;
}
br {
  /* change <br> space length*/
  display: block; /* makes it have a width */
  content: ""; /* clears default height */
  margin-top: 10px; /* change this to whatever height you want it */
}
</style>

## ECD Long-run effects

::: {.incremental}
* Much affirmative evidence of health and education (in-kind) interventions.  
   * Food stamps are different from cash [@Hoynes2016; @Baileyetal2023].  
   * Lottery cash distributions are given at older children, results are mixed [@Akeeetal2010; @Cesarinietal2016].  
* Cash intervention impacts are little known.  
:::

----

## Data

**IRS1040 data at U.S. Census**

::: {.incremental}
* Main tax form used to file a U.S. individual income tax return  
* Every filer  
* Years: 1979, 1984, 1989, 1994-95, 1998-2018  
* Various incomes, dependents (Social Security Numbers)  
:::

----

## Data

**Social Security Administration Numident File**

::: {.incremental}
* Every person born after 1969  
<!-- * Merged with IRS1040 through Social Security Number  -->
* Date of birth, sex, state of birth  
   * Gives: Size of cash transfer in early childhood, family composition  
* $\pm 1$ month of Jan 1, 1982, 1986, 1992  
<!--    * Only these years can be merged with IRS1040 (1979, 1984, 1989)  -->
<!--   * IRS1040 (1979, 1984, 1989) used to predict eligibility for EITC  -->
:::

----

## Data

**North Carolina Education Data**

::: {.incremental}
* Students born in 1992-1999 (1993-98 recentred years)  
* Birth dates, academic achievements, behaioral records, free and reduced-price lunch (FRL) eligibility  
   * Income thresholds are similar between EITC and FRL  
      * \$25,600 (FRL) vs. \$27,400 (EITC) in 2000  
      * EITC eligible rate $\simeq$ 75% of all FRL eligible  
* Born within $\pm$0928 day windows of Jan 1 (excluding closest 8 days on both sides): 44,992 students  
:::

----

## Data

::: {.incremental}
* IRS1040: Parent-dependents linkage  
   * Child's incomes in 23-25, 26-28, 29-31, 32-34 (individual+spouse, 3-year averages)  
* IRS1040+NumidentFile: Date of birth of dependent &rarr; Pre-birth parental incomes &rarr; Pre-birth adjusted gross income (AGI) &rarr; EITC amount  
* Use 1979 incomes (in 1980 tax record) to predict 1980 AGI &rarr; Pre-birth AGI for a Jan 1981 born child  
:::

----

## Data

::: {.incremental}
* Predicted AGI using lagged incomes 
   * Attenuation bias  
   * Less concerns of endogeneity (parental ability correlates with AGI, chance of birth date selection)  
* Create an index: Weighted normalised $z$ scores or each outcome $k=\sum\limits_{k}w_{k}\frac{x_{k, treated}-\hat{\mu}_{x_{k, control}}}{\hat{\sigma}_{x_{k, control}}}$ [@Kling2007]

:::

----

## Empirical strategy

<p class="spacer">Birthday RDD [@SchulkindShapiro2014]</p>

* Dec 31: Eligible in +1 day  
* Jan 1: Eligible in +1 year

----

## Empirical strategy

::: {.incremental}
Strength

* Pure cash transfer impacts, not "coupled with changes to work incentives"

<br>

Concerns

1. Some may deliver earlier, confirmed by [@BucklesHungerman2013;  with a c-section, @SchulkindShapiro2014]  
   * No jump in density for firstborns around new year (Fig II)  
   * Balance in covariates (Tab II) 
   * "Donut-hole" RDD: Drop observations within $\pm$ 8 days of Jan 1  
:::

----

## Empirical strategy

<br>

Concerns

2. Discontinuity in other variables  
   * School year begins near Jan 1 in some states, drop these states  
   * Larger tax benefits in later years: Expect larger impacts in later years

----

## Empirical strategy

<br>
<p style="font-size:30px;"> 
\begin{equation}
Y_{it}=\beta_{0}+\beta_{1}\mathcal 1[z_{i}<0]+\beta_{2}z_{i}+\beta_{3}\mathcal 1[z_{i}<0]z_{i}+\theta_{t}+\epsilon_{it}.
\tag{2}
\end{equation}
</p>

<p style="color:red;"> 
```{r variable description}
#| label: tbl-VariableDescription
library(knitr)
desc <- matrix(c(
    "$z_{i}$",  "Days before Jan 1.", 
    "$\\mathcal 1[z_{i}<0]$", "Treatment assignment.",
    "$\\beta_{1}$", "Main interest."
  ), byrow = T, ncol = 2)
kableExtra::kbl(desc, align = c("r", "l"), label = NA, caption = NA, 
  table.attr = list(style="font-size:26px; color:red;"), format = "markdown")
```
</p>

----

## Results

:::: {layout="[40, 40 ]"}

::: {#first-column}
![](FigIII-1.jpg){height=80%}
:::

::: {#second-column}
![](FigIII-2.jpg){height=80%}
:::

::::

* $295 for 23-25, $433 for 26-28, or 1.2%, 1.6% of adult incomes

----

## Results

:::: {layout="[40, 40 ]" layout-valign="bottom"}

::: {#first-column}
![](FigAVa.jpg){height=80%}
:::

::: {#second-column}
![](FigAVb.jpg){height=80%}
:::

::::

* Parents may shift delivery by 4 days (just after Christmas)...

----

## Results

* Later cohorts have larger impacts (Tab IV)  
* Much smaller impacts for nonfirstborn given transfer is $306 vs. $1,291 (Tab A.VI)  
* Smaller income impacts for women, because they are more likely to file jointly with spouse, and their income shares are lower (Fig IV)  
* Larger impacts in later years (Fig V)

----

## Results

:::: {.columns}

::: {.column width="70%"}
![](FigV.jpg){height="550"}
:::

::: {.column width="30%"}
* $+319 (23-25), $+456 (26-28), larger at 29+  
* Earnings by 26 is predictive of future earnings [@HaiderSolon2006; @Chettyetal2011]  
:::

::::

----

## Mechanism: Parents and family environment

:::: {.columns}

::: {.column width="65%"}
![](Fig6A.jpg){height="550"}
:::

::: {.column width="35%"}
* Family earnings&uarr; by 4% in 3, 4 years (Fig VI)  
:::

::::

----

## Mechanism: Parents and family environment

:::: {.columns}

::: {.column width="65%"}
![](FigA13.jpg){height="550"}
:::

::: {.column width="35%"}
* Dependents&uarr;  
* Poverty status&darr;  
* Filing tax returns&uarr;  
* Being married  
:::

::::

----

## Mechanism: Parents and family environment

1. Liquidity&uarr;  by 10%-20% during < 12 months old  
   * Hightened expenses and reduced incomes for age < 1 year  
   * Employment continuation through mobility: Car purchases, car parts/repair spending (existing studies)  
   * Intergenerational elasticicty of earning (IGE) $\simeq$ .3. So parental earnings impact ($903.7/58250=1.55$%) $\times$ IGE (1/3)$\simeq 5.16$% or 1/3 of impacts on child's future income is through parental earnings&uarr;

----

## Mechanism: Parents and family environment

2. Stress&darr; (no direct evidence shown in this paper)  
   * Confirmed when additional incomes are provided to single mothers [@SchmidtShoreSheppardWatson2023]  
   * Reduced stress changes in interactions with children [@MilliganStabile2011; @EvansGarthwaite2014]  
   * Marital status stability

----

## Mechanism: Child human capital

:::: {.columns}

::: {.column width="55%"}
![](Fig7.jpg){height="550"}
:::

::: {.column width="45%"}
Outcome index&uarr; by .03$\sigma$ per $1000 EITC

* Test scores grades 3-8&uarr; by .037$\sigma$  
* HS graduation&uarr; by .02$\sigma$  
* Suspension&darr; by .02$\sigma$

Robust to donut size and RDD bandwidth (Tab A12)

See also Tab VIII, Tab A13  
:::

::::
----

## Mechanism: Child human capital

:::: {.columns}

::: {.column width="55%"}
![](FigA20.jpg){height="550"}
:::

::: {.column width="45%"}
* Test scores impacts do not fade  
* Suggests human capital channel at play  
* Little gender difference  
   * Due probably to less noise in outcomes than tax filing  
:::

::::
----

## Mechanism: Child human capital

::: {.incremental}
* $\sigma$&uarr; = $2500&uarr; at age 28 per year [@ChettyFriedmanRockoff2014b]  
* .037$\sigma$ (per \$1000)$\times$ $2500=$92 per year per $1000 spent  
   * 9.2% is pretty high...   
* "(C)ould entirely account for our observed wage effects ... $353 per $1000" (?)  
:::

----

## Conclusions

::: {.incremental}
* Long lasting positive impacts of liquidity increase within 12 months after the birth of firstborns  
   * Adult incomes, test scores, behavioral problems, HS graduation  
   * Not evidence for later transfers or non-firstborns  
:::

----

## Conclusions

::: {.incremental}
* NPV of tax receipts from added incomes > initial transfer  
* From Sec IV.B: Impact as % earnings per $1000 spent  
   * Food stamps: .3% to 1% 
   * Perry: Under 1%  
   * EITC: 1% to 2%  
* Previous studies on casino wins, lottery, adoption are on wealthier familes, so their impacts are small  
:::

----

## 感想

* 成人所得への効果? かねてからの疑問に答える画期的な研究  
* さらに、NCデータを使って人的資本経由というメカニズムもある程度明らかにした  
* 現金を親に渡すと子どもに使わない等の懸念が伝統的にあるが、現金でも結果が出てます、と示した   
* 現金移転の方が効率的なので、保健や教育などで現物移転ばかり強調しすぎない方が良いかも  
   * 現金移転は教育に効果、現金で自動車を買うと非難されそう  
* 外的妥当性: EITC無知層(ヒスパニックやアフリカ系)、途上国  
   * 親の知識水準に依存しそう

----

#### References

::: {#refs}
:::

### 2024-07-30 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
added bib file in BarrEggelstonSmith

@@
 #width: 1000
-bibliography: c:/seiro/settings/TeX/seiro.bib
+bibliography: ~../posts/seiro.bib
 resources:
@@
 * Much affirmative evidence of health and education (in-kind) interventions.  
+* Cash intervention impacts are little known.  
    * Food stamps are different from cash [@Hoynes2016; @Baileyetal2023].  
    * Lottery cash distributions are given at older children, results are mixed [@Akeeetal2010; @Cesarinietal2016].  
-* Cash intervention impacts are little known.  
 :::
@@
 * Jan 1: Eligible in +1 year
+
+<br>
+ITT effects: Some eligible (mostly Hispanics, African Americans) do not apply
 
@@
 * Later cohorts have larger impacts (Tab IV)  
-* Much smaller impacts for nonfirstborn given transfer is $306 vs. $1,291 (Tab A.VI)  
+* Much smaller impacts for nonfirstborn &larr; transfer is $306 vs. $1,291 (Tab A.VI)  
 * Smaller income impacts for women, because they are more likely to file jointly with spouse, and their income shares are lower (Fig IV)  
-* Larger impacts in later years (Fig V)
+* Larger impacts in later years (Fig V) &larr; more generous transfers
 
@@
    * Confirmed when additional incomes are provided to single mothers [@SchmidtShoreSheppardWatson2023]  
-   * Reduced stress changes in interactions with children [@MilliganStabile2011; @EvansGarthwaite2014]  
+   * Reduced stress changes the interactions with children [@MilliganStabile2011; @EvansGarthwaite2014]  
    * Marital status stability

### 2024-07-30 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
added bib file in BarrEggelstonSmith

@@
 #width: 1000
-bibliography: ~../posts/seiro.bib
+bibliography: seiro.bib
 resources:

### 2024-07-30 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
Added R installation and dependency installation in publish.yml
Corrected bottom link to BarrEgglestonSmith2022.html

@@
     logo: images/quarto.png
-    footer: <http://seiroito.github.io/Miscellaneous/QuartoFiles/posts/BarrEgglestonSmith/BarrEgglestonSmith2022.html>
+    footer: <http://seiroito.github.io/QuartoFiles/BarrEgglestonSmith/BarrEgglestonSmith2022.html>
     theme: sky

### 2024-07-30 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
Corrected bottom link to BarrEgglestonSmith2022.html

@@
       * EITC eligible rate $\simeq$ 75% of all FRL eligible  
-* Born within $\pm$0928 day windows of Jan 1 (excluding closest 8 days on both sides): 44,992 students  
+* Born within $\pm$ 0928 day windows of Jan 1 (excluding closest 8 days on both sides): 44,992 students  
 :::

### 2024-07-31 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
Added research question page to BarrEgglestonSmith2022.html

@@
    * Lottery cash distributions are given at older children, results are mixed [@Akeeetal2010; @Cesarinietal2016].  
+:::
+
+----
+
+## Research question
+
+::: {.incremental}
+* Does providing cash at infancy (< 1 year) have a long run impact on welfare?  
+* Use fiscal year cutoff for EITC. Born on:  
+   * Dec 31 receives after +1 day  
+   * Jan 1 receives after +1 year  
+* Examine with tax records + social security numident file (birth day)  
 :::
@@
 ::: {.incremental}
-* $\sigma$&uarr; = $2500&uarr; at age 28 per year [@ChettyFriedmanRockoff2014b]  
+* Exam score $\sigma$&uarr; \Rarr; $2500&uarr; at age 28 per year [@ChettyFriedmanRockoff2014b]  
 * .037$\sigma$ (per \$1000)$\times$ $2500=$92 per year per $1000 spent  

### 2024-07-31 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
Added research question page to BarrEgglestonSmith2022.html

@@
 ::: {.incremental}
-* Exam score $\sigma$&uarr; \Rarr; $2500&uarr; at age 28 per year [@ChettyFriedmanRockoff2014b]  
+* Exam score $\sigma$&uarr; \rArr; $2500&uarr; at age 28 per year [@ChettyFriedmanRockoff2014b]  
 * .037$\sigma$ (per \$1000)$\times$ $2500=$92 per year per $1000 spent  

### 2024-07-31 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
Added research question page to BarrEgglestonSmith2022.html

@@
 ::: {.incremental}
-* Exam score $\sigma$&uarr; \rArr; $2500&uarr; at age 28 per year [@ChettyFriedmanRockoff2014b]  
+* Exam score $\sigma$&uarr; &rArr; $2500&uarr; at age 28 per year [@ChettyFriedmanRockoff2014b]  
 * .037$\sigma$ (per \$1000)$\times$ $2500=$92 per year per $1000 spent  

### 2024-07-31 posts/WebPageExample/WebPageExample.qmd
Added WebPageExample.html

--- 2024-07-31 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
@@
 format:
-  revealjs: 
-    slide-number: true
-    chalkboard: 
-      buttons: false
-    preview-links: auto
-    logo: images/quarto.png
-    footer: <http://seiroito.github.io/QuartoFiles/BarrEgglestonSmith/BarrEgglestonSmith2022.html>
-    theme: sky
-include-in-header:
-  - text: |
-      <style>
-      #title-slide
-       .title {
-        font-size: 2.0em;
-      }
-      </style>
-#width: 1000
-bibliography: seiro.bib
+  html:
+    toc: true
+    toc-location: left
+    number-sections: true
+    number-depth: 2
+    embed-resources: true
+    smooth-scroll: true
+    anchor-sections: true
+bibliography: ../seiro.bib
 resources:
-  - BarrEgglestonSmith2022.pdf
-# quarto render BarrEgglestonSmith2022.qmd
+  - WebPageExample.pdf
+# quarto render WebPageExample.qmd
 ---
 <style>
-:root{--r-main-font-size: 36px;}
-.reveal h1, .reveal h2, .reveal h3, .reveal h4, .reveal h5 {
-              text-transform: none;
-      }
-.spacer {
-    margin-top: 20px;
-}
 br {
@@
 
-::: {.incremental}
 * Much affirmative evidence of health and education (in-kind) interventions.  
@@
    * Lottery cash distributions are given at older children, results are mixed [@Akeeetal2010; @Cesarinietal2016].
-:::
-
-----
-
-## Research question
-
-::: {.incremental}
-* Does providing cash at infancy (< 1 year) have a long run impact on welfare?  
-* Use fiscal year cutoff for EITC. Born on:  
-   * Dec 31 receives after +1 day  
-   * Jan 1 receives after +1 year  
-* Examine with tax records + social security numident file (birth day)  
-:::
-
-----
 
@@
 
-::: {.incremental}
 * Main tax form used to file a U.S. individual income tax return  
@@
 * Various incomes, dependents (Social Security Numbers)
-:::
-
-----
-
-## Data
 
@@
 
-::: {.incremental}
 * Every person born after 1969  
@@
 <!--   * IRS1040 (1979, 1984, 1989) used to predict eligibility for EITC  -->
-:::
-
-----
-
-## Data
 
@@
 
-::: {.incremental}
 * Students born in 1992-1999 (1993-98 recentred years)  
@@
 * Born within $\pm$ 0928 day windows of Jan 1 (excluding closest 8 days on both sides): 44,992 students
-:::
 
-----
+#### How data sets are combined
 
-## Data
-
-::: {.incremental}
 * IRS1040: Parent-dependents linkage  
@@
 * Use 1979 incomes (in 1980 tax record) to predict 1980 AGI &rarr; Pre-birth AGI for a Jan 1981 born child
-:::
 
-----
-
-## Data
-
-::: {.incremental}
 * Predicted AGI using lagged incomes 
@@
 * Create an index: Weighted normalised $z$ scores or each outcome $k=\sum\limits_{k}w_{k}\frac{x_{k, treated}-\hat{\mu}_{x_{k, control}}}{\hat{\sigma}_{x_{k, control}}}$ [@Kling2007]
-
-:::
-
-----
 
@@
 
-----
-
-## Empirical strategy
-
-::: {.incremental}
 Strength
@@
    * "Donut-hole" RDD: Drop observations within $\pm$ 8 days of Jan 1  
-:::
-
-----
-
-## Empirical strategy
-
-<br>
-
-Concerns
-
 2. Discontinuity in other variables  
@@
    * Larger tax benefits in later years: Expect larger impacts in later years
-
-----
-
-## Empirical strategy
 
@@
 
-----
-
 ## Results
 
-:::: {layout="[40, 40 ]"}
+::: {#fig-res1 layout-ncol=2}
 
-::: {#first-column}
-![](FigIII-1.jpg){height=80%}
+![](FigIII-1.jpg){height=80%, #fig3-1}
+
+![](FigIII-2.jpg){height=80%, #fig3-2}
+
+Impacts on child incomes in adulthood
 :::
-
-::: {#second-column}
-![](FigIII-2.jpg){height=80%}
-:::
-
-::::
 
@@
 
-----
+::: {#fig-res2 layout-ncol=2}
 
-## Results
+![](FigAVa.jpg){width="350"}
 
-:::: {layout="[40, 40 ]" layout-valign="bottom"}
+![](FigAVb.jpg){width="350"}
 
-::: {#first-column}
-![](FigAVa.jpg){height=80%}
+Impacts by donut size
 :::
 
-::: {#second-column}
-![](FigAVb.jpg){height=80%}
-:::
-
-::::
-
 * Parents may shift delivery by 4 days (just after Christmas)...
-
-----
-
-## Results
 
@@
 
-----
-
-## Results
-
-:::: {.columns}
-
-::: {.column width="70%"}
-![](FigV.jpg){height="550"}
+::: {.column-margin}
+![](FigV.jpg){height="250"}
 :::
 
-::: {.column width="30%"}
 * $+319 (23-25), $+456 (26-28), larger at 29+  
 * Earnings by 26 is predictive of future earnings [@HaiderSolon2006; @Chettyetal2011]
-:::
-
-::::
-
-----
 
@@
 
-:::: {.columns}
-
-::: {.column width="65%"}
-![](Fig6A.jpg){height="550"}
+::: {.column-margin}
+![](Fig6A.jpg){height="250"}
 :::
 
-::: {.column width="35%"}
 * Family earnings&uarr; by 4% in 3, 4 years (Fig VI)
-:::
 
-::::
-
-----
-
-## Mechanism: Parents and family environment
-
-:::: {.columns}
-
-::: {.column width="65%"}
+::: {.column-margin}
 ![](FigA13.jpg){height="550"}
@@
 
-::: {.column width="35%"}
 * Dependents&uarr;  
@@
 * Being married
-:::
-
-::::
-
-----
-
-## Mechanism: Parents and family environment
 
@@
 So parental earnings impact ($903.7/58250=1.55$%) $\times$ IGE (1/3)$\simeq 5.16$% or 1/3 of impacts on child's future income is through parental earnings&uarr;
-
-----
-
-## Mechanism: Parents and family environment
-
 2. Stress&darr; (no direct evidence shown in this paper)  
@@
 
-----
-
 ## Mechanism: Child human capital
 
-:::: {.columns}
-
-::: {.column width="55%"}
-![](Fig7.jpg){height="550"}
+::: {.column-margin}
+![](Fig7.jpg){height="250"}
 :::
 
-::: {.column width="45%"}
 Outcome index&uarr; by .03$\sigma$ per $1000 EITC
@@
 See also Tab VIII, Tab A13
+
+::: {.column-margin}
+![](FigA20.jpg){height="250"}
 :::
 
-::::
-----
-
-## Mechanism: Child human capital
-
-:::: {.columns}
-
-::: {.column width="55%"}
-![](FigA20.jpg){height="550"}
-:::
-
-::: {.column width="45%"}
 * Test scores impacts do not fade  
@@
    * Due probably to less noise in outcomes than tax filing
-:::
 
-::::
-----
-
-## Mechanism: Child human capital
-
-::: {.incremental}
 * Exam score $\sigma$&uarr; &rArr; $2500&uarr; at age 28 per year [@ChettyFriedmanRockoff2014b]  
@@
 * "(C)ould entirely account for our observed wage effects ... $353 per $1000" (?)
-:::
-
-----
 
@@
 
-::: {.incremental}
 * Long lasting positive impacts of liquidity increase within 12 months after the birth of firstborns  
@@
    * Not evidence for later transfers or non-firstborns  
-:::
-
-----
-
-## Conclusions
-
-::: {.incremental}
 * NPV of tax receipts from added incomes > initial transfer  
@@
 * Previous studies on casino wins, lottery, adoption are on wealthier familes, so their impacts are small
-:::
-
-----
 
@@
 
-----
-
 #### References

### 2024-07-31 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
Added link to WebPageExample.html in readme.md

@@
 #width: 1000
-bibliography: seiro.bib
+bibliography: ../seiro.bib
+# If it is 2 levels up, ../../seiro.bib
 resources:

### 2024-08-19 posts/ABIDE/ab.qmd
Added analysis.html in ABIDE and papers in BarrEgglestonSmith

---
title: "Produce random sequence in [.3, .7]"
subtitle: "A/Bテスト用のランダム表示比率"
author: "Seiro Ito"
format:
  html:
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 2
    embed-resources: true
    smooth-scroll: true
    anchor-sections: true
bibliography: ../seiro.bib
resources:
  - ab.pdf
# setwd("C:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/ABIDE/"); quarto::quarto_render("ab.qmd")
# quarto render ab.qmd
---
<style>
br {
  /* change <br> space length*/
  display: block; /* makes it have a width */
  content: ""; /* clears default height */
  margin-top: 10px; /* change this to whatever height you want it */
}
</style>

## A/Bテスト用の乱数

```{r ab test}
#| echo: false
asc <- function(x)
{
  if (!is.factor(x)) x <- as.factor(unlist(x))
  return(levels(x)[as.integer(x)])
}
seqrep <- function(seqn, repvec)
{
  Vec <- NULL
  for (i in 1:length(repvec)) Vec <- c(Vec, paste(rep(seqn, repvec[i]), sep = "", collapse = ""))
  return(Vec)
}
putzeroontop <- function(z, autozero = T, totaldigits = F)
#  autozero: if numeric, add the given digit of zero on top
#  totaldigits: if numeric, total digits will be set (001 is 3) if maximum 0s to be added is smaller
#    modified to accept NA in z: July 18, 2016
{
  iic <- grep("\\D", z, perl = T)
  iina <- which(is.na(z))
  iin <- grep("\\d", z, perl = T)
  digs <- nchar(asc(z))
  adddig <- max(digs, na.rm = T) - digs
  if (is.numeric(autozero)) 
  {
    maxadd <- max(max(digs, na.rm = T) - digs, na.rm = T)
    if (autozero > maxadd) adddig <- adddig + (autozero - maxadd)
  }
  if (is.numeric(totaldigits))
    if (totaldigits > max(digs, na.rm = T)) 
    adddig <- adddig + (totaldigits - max(digs, na.rm = T))
  adddig[c(iic, iina)] <- 0
  z <- paste(seqrep("0", adddig), z, sep = "")
  z[grep("NA", z)] <- NA
  return(z)
}
library(knitr)
```

最初の50日分の乱数

* 0.5近傍が少ない  
   * 悪いことではないが、少し極端な値が多いかも  
* 緑線は平均値

```{r plot sequence for first 50 days first trial}
#| echo: false
library(data.table)
library(ggplot2)
set.seed(20240731)
rn <- runif(10000)
rn <- rn[rn >= .25 & rn <= .75]
rn2 <- data.table(days=1:length(rn), val=c(rn))
library(ggplot2)
ggplot(rn2[1:50, ], aes(x = val)) +
  geom_histogram(bins = length(seq(.25, .75, .02))) + 
  geom_vline(xintercept = mean(rn2[1:50, val]), color = "green")
```

* 何度も乱数を発生させても良いけど、均一分布に近くなるような基準が必要かも  
   * 最初の50日おける最小頻度と最大頻度の差が一定範囲内(2以下)にする制約

以下のコードをRで読み込みます。

```{r generate sequence for first 1400 days}
#| warning: false
#| output: false
set.seed(20240731)
eye <- 0
dh <- 10
while (dh > 2.0) {
  eye <- eye + 1
  rn <- runif(10000)
  rn <- rn[rn >= .24 & rn <= .76]
  h <- hist(rn[1:50], breaks=seq(min(rn), max(rn), length.out=20))
  imax <- which.max(h$density)
  imin <- which.min(h$density)
  dh <- h$density[imax] - h$density[imin]
}
rn2 <- data.table(days=1:length(rn), val=c(rn))
rn2[, more := "B"]
rn2[val > .5, more := "A"]
rn2[1:50, ]
```
```{r display sequence for first 1400 days}
rn3 <- matrix(round(rn[1:(200*7)], 2), ncol = 7, byrow = T)
dimnames(rn3) <- list(putzeroontop(1:nrow(rn3)), c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
kable(rn3)
write.table(formatC(rn3, digits = 2, format = "f"), "ab.tsv", quote = F, sep = "\t")
```

`r formatC(eye, digits = 0, format = "f")`回目の乱数発生で基準を満たしました。dh=`r dh`, 中央点は`r h$mids`、各中央点の確率密度は`r h$density`。

```{r plot sequence for first 50 days}
#| warning: false
#| echo: false
g <- ggplot(rn2, aes(x = days, y = val, colour = as.numeric(val), fill = as.numeric(val))) +
  geom_point() + geom_line(aes(colour = as.numeric(val)))
g + scale_x_continuous(limits = c(1, 50)) + 
  theme(
    legend.position = "bottom"
  ) + 
  geom_hline(data = rn2, yintercept = mean(rn2[1:50, val]), color = "green")
```
```{r plot density for first 50 days, warning = F}
#| echo: false
#| warning: false
hist(rn[1:50], breaks=seq(min(rn), max(rn), length.out=20))
rn2 <- data.table(days=1:length(rn), val=rn)
#ggplot(rn2[1:50, ], aes(x=val)) +
#  geom_histogram(bins = 20) + 
#  geom_vline(xintercept = mean(rn2[1:50, val]), color = "green")
```

最初の200日の乱数とそのヒストグラム。

```{r plot sequence for first 200 days, warning = F}
#| echo: false
rn2 <- data.table(days=1:length(rn), val=c(rn))
rn2[, more := "B"]
rn2[val > .5, more := "A"]
library(ggplot2)
g <- ggplot(rn2, aes(x = days, y = val, colour = as.numeric(val), 
    fill = as.numeric(val))) +
  geom_point() + geom_line(aes(colour = as.numeric(val)))
g + scale_x_continuous(limits = c(1, 200)) + 
  theme(
    legend.position = "bottom"
  ) + 
  geom_hline(data = rn2, yintercept = mean(rn2[1:200, val]), color = "green")
```

```{r plot density for first 200 days, warning = F}
#| echo: false
hist(rn[1:200], breaks=seq(min(rn), max(rn), length.out=20))
```

最初の100週の乱数とそのヒストグラム。

```{r plot sequence for first 100 weeks, warning = F}
#| echo: false
library(ggplot2)
g <- ggplot(rn2, aes(x = days, y = val, colour = as.numeric(val), 
    fill = as.numeric(val))) +
  geom_point() + geom_line(aes(colour = as.numeric(val)))
g + scale_x_continuous(limits = c(1, 700)) + 
  theme(
    legend.position = "bottom"
  ) + 
  geom_hline(data = rn2, yintercept = mean(rn2[1:700, val]), color = "green")
```

```{r plot density for first 100 weeks, warning = F}
#| echo: false
library(ggplot2)
hist(rn[1:1400], breaks=seq(min(rn), max(rn), length.out=20))
```

```{r reshape and attach dates, eval = F}
rn0 <- fread("ab.tsv")
library(lubridate)
abdates <- seq(ymd('2024-08-01'), length.out = 1400, by = '1 day')
rn0[, V1 := NULL]
ab <- data.table(date=abdates, dow = weekdays(abdates), 
  value=formatC(c(t(as.matrix(rn0))), digits = 2, format = "f"))
write.table(ab, "abLong.tsv", quote = F, sep = "\t")
```

### 2024-08-19 posts/ABIDE/analysis.qmd
Added analysis.html in ABIDE and papers in BarrEgglestonSmith

---
title: "A/Bテストの分析"
author: "Seiro Ito"
execute: 
  echo: false
format:
  html:
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 2
    embed-resources: true
    smooth-scroll: true
    anchor-sections: true
bibliography: ../seiro.bib
resources:
  - analysis.pdf
# setwd("C:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/ABIDE/"); quarto::quarto_render("analysis.qmd")
# quarto render analysis.qmd
---
<style>
br {
  /* change <br> space length*/
  display: block; /* makes it have a width */
  content: ""; /* clears default height */
  margin-top: 10px; /* change this to whatever height you want it */
}
</style>

# データ

```{r list files}
fn <- list.files(path="data/", full.names = T)
```
分析対象の日付
```{r list dates}
library(data.table)
dts <- unique(gsub(".tsv", "", gsub("data..", "", fn)))
dt1 <- as.IDate(
  as.POSIXct(paste0("2024", substring(dts, 1, 4)), format = "%Y%m%d")
  )
dt2 <- as.IDate(
  as.POSIXct(paste0("2024", substring(dts, 5, 8)), format = "%Y%m%d")
  )
paste(dt1, weekdays(dt1), "~", dt2, weekdays(dt2))
```
```{r read data}
#| echo: false
#| results: hide
dInList <- lapply(fn, fread, skip = 6)
nmInList <- lapply(dInList, function(x) as.character(x[1, ]))
nmInList <- lapply(nmInList, function(x) c("EventPath", "DateHour", x[-c(1:2, length(x))], "total"))
dInList <- lapply(fn, fread, skip = 9)
lapply(1:length(dInList), function(i) setnames(dInList[[i]], nmInList[[i]]))
```{r aggregate data by date}
dt <- rbindlist(dInList, use.names = T, fill = T)
dt[, Date := as.POSIXct(paste0(DateHour, ":00"), format = "%Y%m%d%H:%M")]
dt[, date := as.IDate(Date)]
#### read randomisation file
rnd <- fread("abLong.tsv")
dtr <- merge(dt, rnd, by = "date", all.x = T)
dtr[, AB := "A"]
dtr[grepl("B", EventPath), AB := "B"]
dtr[, AB := factor(AB)]
dtr[, Denominator := round(value, 1)]
dtr[grepl("B", EventPath), Denominator := round(1-value, 1)]
DtForG <- dtr[, .(UEngage = sum(user_engagement), Denominator = Denominator[1]), by = .(date, AB)][, 
  .(date, AB, UEngage, Denominator, UERate = UEngage/Denominator)]
```
発生したデータ

* user_engagement: 一定時間以上滞在閲覧数  
* click: クリック数  
* value: Aを表示する確率

```{r }
dtr[, .(AB, Date, page_view, user_engagement, scroll, click, total, Denominator)]
```

日ごとに平均したデータ
```{r }
DtForG
```
日ごと平均の記述統計

```{r }
summary(DtForG)
```

UERate=一定時間以上滞在閲覧数/表示比率=1日表示させたときの一定時間以上滞在閲覧数
```{r plot sequence for first 2 weeks}
#| warning: false
#| echo: false
library(ggplot2)
g <- ggplot(DtForG, aes(x = date, y = UERate, colour = AB, fill = AB, group = AB)) +
  geom_point() + geom_line(aes(colour = AB))
g <- g + scale_x_date(date_labels = "%B%d日(%a)") +
  theme(
    legend.position = "bottom"
  )
####   +  geom_hline(data = DtForG, yintercept = 
####    DtForG[, .(mean = mean(UERate)), by = AB]
####  , color = c(""))
g
```
`{r} DtForG[, min(date)]` ~ `{r} DtForG[, max(date)]`における	AとBのUERateの平均値: `{r} round(DtForG[, .(mean = mean(UERate)), by = AB][, mean], 2)`

# ここまでの結論

* ドングリの背比べ、という感じでしょうか...  
* 異常値を落として分析するか?

一般的な傾向として

* 金曜-日曜は低調  
* エンゲージメントが異常に増える日がある(なぜ?): 8月1日(木)、8月5日(月)

分析上の課題

* データの発生メカニズムを正確に理解して、結果を正しく解釈する(宿題)  
* 急変動を引き起こす社会の要因を見出す方法がない  
   * XのTrendingなどのデータを参照すべき?

# 今後

* clickを分析したいが、18日間で1カウントのみなので分析不能  
* scrollは18日間AB合わせて56カウントなので、分析単位を週(もしくは平日、週末)にすれば何かできるかも  
* 8月末に実験続行の是非を決めるが、データ発生が少ないときは、分析単位を日ごとではなく週ごとにするために延長すべきか

```{r plot density for first 100 weeks, warning = F}
#| echo: false
#| eval: false
rn0 <- fread("ab.tsv")
library(lubridate)
abdates <- seq(ymd('2024-08-01'), length.out = 1400, by = '1 day')
rn0[, V1 := NULL]
ab <- data.table(date=abdates, dow = weekdays(abdates), 
  value=formatC(c(t(as.matrix(rn0))), digits = 2, format = "f"))
write.table(ab, "abLong.tsv", quote = F, sep = "\t")
```

### 2024-08-28 posts/ABIDE/analysis.qmd
2nd trial in rendering.

@@
   echo: false
+  freeze: auto
 format:

### 2024-08-29 posts/ABIDE/ab.qmd
Deleted thousands of png files in ab.qmd and used plot = F in hist command that generated them

@@
   rn <- rn[rn >= .24 & rn <= .76]
-  h <- hist(rn[1:50], breaks=seq(min(rn), max(rn), length.out=20))
+  h <- hist(rn[1:50], breaks=seq(min(rn), max(rn), length.out=20), plot = F)
   imax <- which.max(h$density)

### 2024-08-29 posts/ABIDE/analysis.qmd
Deleted thousands of png files in ab.qmd and used plot = F in hist command that generated them

@@
 </style>
+
+# What's new `{r} Sys.Date()`
+
+* user engagement rateの振幅幅: $A>B$ の傾向あり、AはUERate=0が2日あり  
+   * 異常値(UERate$\geqslant 50$)抜きだと、平均は両者ほぼ同じ、標準偏差はAが3割ほど高い
+
+# A/B
+
+::: {layout-ncol=2}
+
+![A](A_08-29-2024.jpg){#fig-A}
+
+![B](B_08-29-2024.jpg){#fig-B}
+
+IDEスクエアのトップページ: A/B 
+:::
 
@@
 ```
-`{r} DtForG[, min(date)]` ~ `{r} DtForG[, max(date)]`における	AとBのUERateの平均値: `{r} round(DtForG[, .(mean = mean(UERate)), by = AB][, mean], 2)`
+```{r mean and std}
+#| echo: false
+ms <- DtForG[, .(mean = mean(UERate), std = var(UERate)^(.5)), by = AB]
+ms
+ms1 <- DtForG[date > as.IDate("2024-08-01", format = "%Y-%m-%d"), 
+  .(mean = mean(UERate), std = var(UERate)^(.5)), by = AB]
+ms2 <- DtForG[UERate < 50, 
+  .(mean = mean(UERate), std = var(UERate)^(.5)), by = AB]
+```
+
+`{r} DtForG[, min(date)]` ~ `{r} DtForG[, max(date)]`における	AとBのUERateの平均値: `{r} round(ms[, mean], 2)`
+
+* 一見してAは振幅幅が大きい(ゼロが2回)と思ったが、標準偏差は小さく、30越えが`{r} nrow(DtForG[UERate>30 & grepl("A", AB), ])`回、Bは`{r} nrow(DtForG[UERate>30 & grepl("B", AB), ])`回なので上振れもあまりない  
+* 変動係数=標準偏差/平均: A=`{r} round(ms[grepl("A", AB), std], 2)`/`{r} round(ms[grepl("A", AB), mean], 2)`=`{r} round(ms[grepl("A", AB), std]/ms[grepl("A", AB), mean], 2)`、B=`{r} round(ms[grepl("B", AB), std], 2)`/`{r} round(ms[grepl("B", AB), mean], 2)`=`{r} round(ms[grepl("B", AB), std]/ms[grepl("B", AB), mean], 2)`
+* 実は初日の影響: 初日抜きの変動係数: A=`{r} round(ms1[grepl("A", AB), std], 2)`/`{r} round(ms1[grepl("A", AB), mean], 2)`=`{r} formatC(ms1[grepl("A", AB), std]/ms1[grepl("A", AB), mean], digits=2, format = "f")`、B=`{r} round(ms1[grepl("B", AB), std], 2)`/`{r} round(ms1[grepl("B", AB), mean], 2)`=`{r} round(ms1[grepl("B", AB), std]/ms1[grepl("B", AB), mean], 2)`
+* 異常値除外後(UERate$<50$): 変動係数: A=`{r} round(ms2[grepl("A", AB), std], 2)`/`{r} round(ms2[grepl("A", AB), mean], 2)`=`{r} formatC(ms2[grepl("A", AB), std]/ms2[grepl("A", AB), mean], digits=2, format = "f")`、B=`{r} round(ms2[grepl("B", AB), std], 2)`/`{r} round(ms2[grepl("B", AB), mean], 2)`=`{r} round(ms2[grepl("B", AB), std]/ms2[grepl("B", AB), mean], 2)`、標準偏差はAがBの`{r} formatC(ms2[grepl("A", AB), std]/ms2[grepl("B", AB), std], digits=2, format = "f")`倍
 

### 2024-09-19 posts/ABIDE/analysis.qmd
Added figure in BarrEgglestonSmith and moved ab.qmd to other folder

--- 2024-08-28 posts/ABIDE/analysis.qmd
@@
 
+# What's new `{r} Sys.Date()`
+
+* user engagement rateの振幅幅: $A>B$ の傾向が続いている  
+   * 平均値はほぼ同じ  
+* scrollではA/Bの差はあまりない  
+* 前回指摘した「週後半に閲覧数が落ちる傾向」がより鮮明になってきた  
+   * なんとかすべきかも
+
+# A/B
+
+::: {layout-ncol=2}
+
+![A](A_08-29-2024.jpg){#fig-A}
+
+![B](B_08-29-2024.jpg){#fig-B}
+
+IDEスクエアのトップページ: A/B 
+:::
+
 # データ
@@
 ```{r list files}
-fn <- list.files(path="data/", full.names = T)
+fn <- list.files(path="data/", pattern = "^[ab]", full.names = T)
 ```
@@
 dtr[grepl("B", EventPath), Denominator := round(1-value, 1)]
-DtForG <- dtr[, .(UEngage = sum(user_engagement), Denominator = Denominator[1]), by = .(date, AB)][, 
-  .(date, AB, UEngage, Denominator, UERate = UEngage/Denominator)]
+DtForG <- dtr[, .(
+  UEngage = sum(user_engagement), 
+  Scroll = sum(scroll), Denominator = Denominator[1]), by = .(date, AB)][, 
+  .(date, AB, UEngage, Scroll, Denominator, 
+  UERate = UEngage/Denominator, ScrollRate = Scroll/Denominator)]
 ```
@@
 
-* user_engagement: 一定時間以上滞在閲覧数  
-* click: クリック数  
-* value: Aを表示する確率
+* user_engagement: 一定時間以上滞在閲覧数 (incl. scrollers)  
+* scroll: 最下部まで90%以上スクロールした閲覧数  
+* click: ドメイン外部へのクリック数
 
@@
 dtr[, .(AB, Date, page_view, user_engagement, scroll, click, total, Denominator)]
+summary(dtr[, .(AB, Date, page_view, user_engagement, scroll, click, total, Denominator)])
 ```
@@
 
+```{r mean and std}
+#| echo: false
+#| warning: false
+library(kableExtra)
+grepout <- function(str, x)
+  # returns element of match (not numbers)
+  x[grep(str, x, perl = T)]
+oldcolnames <- colnames(DtForG)
+setnames(DtForG, c("UEngage", "Scroll"), paste0("val.", c("UEngage", "Scroll")))
+DtForGL <- reshape(DtForG, direction = "long", idvar = c("date", "AB", "Denominator"),
+  varying = grepout("val", colnames(DtForG)))
+DtForGL[, rate := val/Denominator]
+setnames(DtForGL, "time", "variable")
+DtForGL[, c("UERate", "ScrollRate") := NULL]
+DtForGL
+msL <- DtForGL[, .(mean = mean(rate), std = var(rate)^(.5), cofvar = var(rate)^(.5)/mean(rate)), by = .(variable, AB)]
+numcols <- c("mean", "std", "cofvar")
+msL[, (numcols) := lapply(.SD, round, 3), .SDcols = numcols]
+#msL
+kt <- kbl(msL, align = "rcccc",
+  caption = "UERate、ScrollRateの平均、標準偏差、変動係数")
+#kt <- column_spec(kt, 1, width = "2cm; min-width:2cm;")
+#kt <- column_spec(kt, 2, width = ".5cm; min-width:.5cm;")
+#kt <- column_spec(kt, 3:5, width = "1.0cm; min-width:1.0cm;")
+setnames(DtForG, oldcolnames)
+ms <- DtForG[, .(mean = mean(UERate), std = var(UERate)^(.5)), by = AB]
+ms1 <- DtForG[date > as.IDate("2024-08-01", format = "%Y-%m-%d"), 
+  .(mean = mean(UERate), std = var(UERate)^(.5)), by = AB]
+ms2 <- DtForG[UERate < 50, 
+  .(mean = mean(UERate), std = var(UERate)^(.5)), by = AB]
+```
+
+```{r, results = "asis"}
+kable_styling(kt,
+        bootstrap_options = "striped", 
+        full_width = F, 
+        position = "left"
+        )
+```
+
 UERate=一定時間以上滞在閲覧数/表示比率=1日表示させたときの一定時間以上滞在閲覧数
-```{r plot sequence for first 2 weeks}
+```{r plot UERate sequence for first weeks}
 #| warning: false
@@
 ```
-`{r} DtForG[, min(date)]` ~ `{r} DtForG[, max(date)]`における	AとBのUERateの平均値: `{r} round(DtForG[, .(mean = mean(UERate)), by = AB][, mean], 2)`
+
+`{r} DtForG[, min(date)]` ~ `{r} DtForG[, max(date)]`における	AとBのUERateの平均値: `{r} round(ms[, mean], 2)`
+
+* 一見してAは振幅幅が大きい(ゼロが2回)ように見えるが、実際にこの印象は正しい  
+   * 30越えも`{r} nrow(DtForG[UERate>30 & grepl("A", AB), ])`回ある(Bは`{r} nrow(DtForG[UERate>30 & grepl("B", AB), ])`回で上振れも少ない)が、計算すると標準偏差は小さいので、この印象は誤っているように見える  
+   * 変動係数=標準偏差/平均: A=`{r} round(ms[grepl("A", AB), std], 2)`/`{r} round(ms[grepl("A", AB), mean], 2)`=`{r} round(ms[grepl("A", AB), std]/ms[grepl("A", AB), mean], 2)`、B=`{r} round(ms[grepl("B", AB), std], 2)`/`{r} round(ms[grepl("B", AB), mean], 2)`=`{r} round(ms[grepl("B", AB), std]/ms[grepl("B", AB), mean], 2)`
+   * 実は初日の影響: 初日抜きの変動係数: A=`{r} round(ms1[grepl("A", AB), std], 2)`/`{r} round(ms1[grepl("A", AB), mean], 2)`=`{r} formatC(ms1[grepl("A", AB), std]/ms1[grepl("A", AB), mean], digits=2, format = "f")`、B=`{r} round(ms1[grepl("B", AB), std], 2)`/`{r} round(ms1[grepl("B", AB), mean], 2)`=`{r} round(ms1[grepl("B", AB), std]/ms1[grepl("B", AB), mean], 2)`
+   * 異常値除外後(UERate$<50$): 変動係数: A=`{r} round(ms2[grepl("A", AB), std], 2)`/`{r} round(ms2[grepl("A", AB), mean], 2)`=`{r} formatC(ms2[grepl("A", AB), std]/ms2[grepl("A", AB), mean], digits=2, format = "f")`、B=`{r} round(ms2[grepl("B", AB), std], 2)`/`{r} round(ms2[grepl("B", AB), mean], 2)`=`{r} round(ms2[grepl("B", AB), std]/ms2[grepl("B", AB), mean], 2)`、変動係数はAがBの`{r} formatC(ms2[grepl("A", AB), std]/ms2[grepl("B", AB), std], digits=2, format = "f")`倍
+
+ScrollRate=scrollした閲覧数/表示比率=1日表示させたときのscroll閲覧数
+```{r plot ScrollRate sequence for first weeks}
+#| warning: false
+#| echo: false
+library(ggplot2)
+g <- ggplot(DtForG, aes(x = date, y = ScrollRate, colour = AB, fill = AB, group = AB)) +
+  geom_point() + geom_line(aes(colour = AB))
+g <- g + scale_x_date(date_labels = "%B%d日(%a)") +
+  theme(
+    legend.position = "bottom"
+  )
+####  , color = c(""))
+g
+```
+`{r} DtForG[, min(date)]` ~ `{r} DtForG[, max(date)]`における	AとBのScrollRateの平均値: `{r} round(msL[grepl("S", variable), mean], 2)`
+
+* スクロールは、平均でも標準偏差でも、A/Bの差があまりない
 
@@
 * 金曜-日曜は低調
-* エンゲージメントが異常に増える日がある(なぜ?): 8月1日(木)、8月5日(月)
 
@@
 
-* clickを分析したいが、18日間で1カウントのみなので分析不能  
 * scrollは18日間AB合わせて56カウントなので、分析単位を週(もしくは平日、週末)にすれば何かできるかも  
-* 8月末に実験続行の是非を決めるが、データ発生が少ないときは、分析単位を日ごとではなく週ごとにするために延長すべきか
+* 週ごとの特徴の違いを組み入れた分析  
+* 9月末まで実験継続、特筆すべき新たな傾向がなければ終了
+
+# Past news
+
+### What's new 2024-08-29
+
+* user engagement rateの振幅幅: $A>B$ の傾向あり、AはUERate=0が2日あり  
+   * 異常値(UERate$\geqslant 50$)抜きだと、平均は両者ほぼ同じ、標準偏差はAが3割ほど高い
 

### 2024-10-14 posts/ABIDE/analysis.qmd
Added sky_add_FilePointer.html to apply, if want to, sky_add.css to all quarto posts

@@
 
-* user engagement rateの振幅幅: $A>B$ の傾向が続いている  
-   * 平均値はほぼ同じ  
-* scrollではA/Bの差はあまりない  
-* 前回指摘した「週後半に閲覧数が落ちる傾向」がより鮮明になってきた  
-   * なんとかすべきかも
+* A/B: いずれの結果指標においても、どちらが優れているとは言えない
 
@@
 
-* 一見してAは振幅幅が大きい(ゼロが2回)ように見えるが、実際にこの印象は正しい  
+* 一見してAは振幅幅が大きい(ゼロが`r nrow(DtForG[UERate==0 & grepl("A", AB), ])`回)ように見えるが、実際にこの印象は正しい  
    * 30越えも`{r} nrow(DtForG[UERate>30 & grepl("A", AB), ])`回ある(Bは`{r} nrow(DtForG[UERate>30 & grepl("B", AB), ])`回で上振れも少ない)が、計算すると標準偏差は小さいので、この印象は誤っているように見える  
@@
 
-# ここまでの結論
+# 推計
 
-* ドングリの背比べ、という感じでしょうか...  
-* 異常値を落として分析するか?
+曜日効果
 
-一般的な傾向として
+```{r regress on weekdays, warning = F, message = F}
+clx <-function(fm, dfcw = 1, cluster, returnV = F, dfc2 = F, DFC = F,
+  Estfun = "sandwich", deviation = F)
+#  1 way clustering robust covariance matrix
+#  following Arai 2010
+#  fm: fitted lm model
+#  cluster: 1 column matrix
+#  dfcw: degree of freedom adjustment when using deviation data
+#   Note: When using the FE estimator with N obs and M groups, 
+#   lm uses dof as N-K, but it should be N-M-K so dfcw = (N-K-M)/(N-K).
+#   In FD estimator, there is no need of dof correction.
+#  dfc2: some suggests this dfc, but not sure if this is correct
+#  DFC: User provided dof
+#  returnV: if T, aslo returns cluster-robust covariance
+#  ESTfun: use estfun provided by "sandwich" package or "seiro"
+#  deviation: T if using deviations data, if F, dfcw is set to 1
+{
+  require(sandwich); require(lmtest)
+  M <- length(unique(cluster))
+  N <- length(cluster)
+  dfc <- (M / (M - 1)) * ((N - 1) / (N - fm$rank))
+  if (dfc2) dfc <- (N -  fm$rank) / (N - M - fm$rank)
+  if (DFC) dfc <- DFC
+  #  switch estfun according to specified by estfun = "sandwich" (default) or "seiro"
+  esfun <- function(fm, ESTfun) switch(ESTfun, sandwich = estfun(fm), seiro = estfun2(fm))
+  #u <- apply(estfun2(fm), 2, function(x) tapply(x, cluster, sum))
+  # This sums estimating function [resid*x (in FOC of OLS)] by cluster.
+  u <- apply(esfun(fm, Estfun), 2, function(x) tapply(x, cluster, sum))
+  # There is no attribute called "df": lm$df.residual, glm$residual.
+  # But leave it for backward compatibility.
+  if (any(grepl("^df$", names(fm))))  
+    dfcw <-  fm$df /(fm$df - (M-1)) else
+  if (deviation)  
+    dfcw <-  fm$df.residual /(fm$df.residual - (M-1))
+  vcovCL <- dfc * sandwich(fm, meat = crossprod(u) / N) * dfcw
+  #diag(vcovCL)^.5
+  if (returnV) list(est = lmtest::coeftest(fm, vcov. = vcovCL), 
+      ci = lmtest::coefci(fm, vcov. = vcovCL),
+      V = vcovCL, dfc = dfc, dfcw = dfcw, reg = fm, 
+      clusters.M = M, cluster = cluster, N = length(fm$fit)) else
+  coeftest(fm, vcovCL) 
+}
 
-* 金曜-日曜は低調
+DtForG[, wkdy := strftime(date, format = "%a")]
+DtForG[, wkdy := factor(wkdy, levels = c("日", "月", "火", "水", "木", "金", "土"))]
+DtForG[, dywk := wday(date)]
+DtForG[, dywk := factor(dywk)]
 
-分析上の課題
+tc <- textConnection(
+"20240801 ページ公開
+20240801 トップページお知らせ欄掲載
+20240805 Xにリニューアル情報投稿
+20240813 おすすめ記事差し替え
+20240813 おすすめ記事差し替え（トップ：ベネズエラ）
+20240829 おすすめ記事差し替え（トップ：バングラデシュ）"
+)
+tc <- data.table(read.table(tc))
+tc[, date := as.POSIXct(paste0(V1, "01:00"), format = "%Y%m%d%H:%M")]
+tc[, date := as.IDate(date)]
+setcolorder(tc, c("V1", "date", "V2"))
+tc[, V1 := NULL]
+setnames(tc, c("date", "event"))
+setkey(DtForG, date)
+setkey(tc, date)
+DtForG <- tc[DtForG]
+DtForG[, Sashikae := 0L]
+DtForG[grepl("差し替え", event), Sashikae := 1L]
+DtForG[, Sashikae1 := shift(Sashikae, n=1, "lead")]
+DtForG[, Sashikae2 := shift(Sashikae, n=2, "lead")]
+DtForG[, Sashikae3 := shift(Sashikae, n=3, "lead")]
+DtForG[, Sashikae3 := Sashikae1+Sashikae2+Sashikae3]
+DtForG[is.na(Sashikae3), Sashikae3 := 0L]
+DtForG[, c("Sashikae1", "Sashikae2") := NULL]
+DtForG[, event := factor(event)]
+#### UERate
+lm1 <- lm(UERate ~ AB+wkdy+Sashikae3, data = DtForG)
+lm2 <- lm(UERate ~ -1+AB:wkdy, data = DtForG)
+lm2 <- lm(UERate ~ -1+AB:wkdy+AB*Sashikae3, data = DtForG)
+cr1 <- clx(lm1, cluster = DtForG[, dywk])
+cr2 <- clx(lm2, cluster = DtForG[, dywk])
+library(fixest)
+fe1 <- feols(UERate ~AB+Sashikae3 | wkdy, DtForG)
+fe2 <- feols(UERate ~AB+AB*wkdy | wkdy, DtForG)
+fe2 <- feols(UERate ~AB+AB*wkdy+AB*Sashikae3 | wkdy, DtForG)
+####summary(fe1, cluster = "wkdy")
+####summary(fe2, cluster = "wkdy")
+library(modelsummary)
+#### UERate
+lm3 <- lm(ScrollRate ~ AB+wkdy+Sashikae3, data = DtForG)
+lm4 <- lm(ScrollRate ~ -1+AB:wkdy, data = DtForG)
+lm4 <- lm(ScrollRate ~ -1+AB:wkdy+AB*Sashikae3, data = DtForG)
+cr3 <- clx(lm3, cluster = DtForG[, dywk])
+cr4 <- clx(lm4, cluster = DtForG[, dywk])
+library(fixest)
+fe3 <- feols(ScrollRate ~AB+Sashikae3 | wkdy, DtForG)
+fe4 <- feols(ScrollRate ~AB+AB*wkdy | wkdy, DtForG)
+fe4 <- feols(ScrollRate ~AB+AB*wkdy+AB*Sashikae3 | wkdy, DtForG)
+####summary(fe3, cluster = "wkdy")
+###summary(fe4, cluster = "wkdy")
+library(modelsummary)
+crlist <- list("UE: base" = cr1, "UE: with interactions" = cr2, "SR: base" = cr3, "SR: with interactions" = cr4)
+####modelsummary(crlist, statistic = "p.value")
+modelplot(crlist[1:2], coef_omit = 'Interc')
+```
 
-* データの発生メカニズムを正確に理解して、結果を正しく解釈する(宿題)  
-* 急変動を引き起こす社会の要因を見出す方法がない  
-   * XのTrendingなどのデータを参照すべき?
+曜日効果のB-A差
+```{r}
+felist <- list("UE: base" = fe1, "UE: with interactions" = fe2, "SR: base" = fe3, "SR: with interactions" = fe4)
+felist1 <- dvnames(felist[c(1, 3)])
+felist2 <- dvnames(felist[c(2, 4)])
+modelsummary(felist, statistic = "p.value")
+####modelplot(felist1, draw = F)
+library(ggplot2)
+modelplot(felist1, color = "blue") + facet_grid(~model)
+modelplot(felist2, color = "blue") + facet_grid(~model)
+```
 
-# 今後
+user engagement rates:
 
-* scrollは18日間AB合わせて56カウントなので、分析単位を週(もしくは平日、週末)にすれば何かできるかも  
-* 週ごとの特徴の違いを組み入れた分析  
-* 9月末まで実験継続、特筆すべき新たな傾向がなければ終了
+* 曜日効果(曜日によるアクセス多寡)はあるが、解釈できない  
+* A/B: どちらが優れているとは言えない  
+* トップ記事差し替え効果: 認められない
+
+scroll rates:
+
+* 曜日効果(曜日によるアクセス多寡)はあるが、解釈できない  
+    * UEもscrollも似たような曜日効果  
+* A/B: どちらが優れているとは言えない  
+* トップ記事差し替え効果: 認められない
+
+# 結論
+
+```{r }
+library(kableExtra)
+tc <- textConnection(
+"variables mean variability
+UserEngagementRate 0 +A
+ScrollRate 0 0")
+tc <- data.table(read.table(tc, header = T))
+tc[, mean := as.character(mean)]
+tc[, mean := ""]
+tc[2, variability := ""]
+####setnames(tc, c("結果指標", "平均", "振幅幅"))
+kt <- kbl(tc, align = "rcc", row.names = F, caption = "Summary of A/B tests")
+kt <- kable_styling(kt, full_width = F, position = "left", fixed_thead = T, 
+    html_font = "Cambria", bootstrap_options = c("condensed"))
+kt <- row_spec(kt, 0, align = "rcc", background = "#e6f2ff")
+kt <- column_spec(kt, 1, width = "3cm; min-width:3cm;")
+kt <- column_spec(kt, 2:3, width = "2.25cm; min-width:2.25cm;")
+ktftnt <- c("A/B are randomly shown to all visitors.
+Probability of A/B changes by the hour.
+User engagement rate is expected number of browsers per day that stay longer than the threshold duration.
+Scroll rate is expected number of browsers per day that scroll the page to the bottom.", "OLS estimates.
+Regressions with interaction terms incorporate day-of-the-week fixed effects.
+")
+kt <- footnote(kt, 
+           general = "Periods: August 1 - September 30, 2024. ",
+           number = ktftnt
+           )
+```
+```{r}
+#| label: tblsummary
+#| tbl-cap: "Summary of A/B tests"
+library(tinytable)
+####tt(tc, width = c(2, 1, 2), notes = list("1"=ktftnt[1], "2"=ktftnt[2]))
+ttc <- tt(tc)
+ttc <- style_tt(ttc, j = 2:3, align = "c") 
+ttc <- style_tt(ttc, j = 1, align = "r") 
+ttc <- style_tt(ttc, i = 0, background = "#e6f2ff")
+ttc
+```
+
+* A/B: いずれの結果指標においても、どちらが優れているとは言えない
 
 # Past news
+
+### What's new 2024-09-05
+
+* user engagement rateの振幅幅: $A>B$ の傾向が続いている  
+   * 平均値はほぼ同じ  
+* scrollではA/Bの差はあまりない  
+* 前回指摘した「週後半に閲覧数が落ちる傾向」がより鮮明になってきた  
+   * なんとかすべきかも
 

### 2024-10-14 posts/BarthPapageorgeThom/BPT.qmd
Added sky_add_FilePointer.html to apply, if want to, sky_add.css to all quarto posts

---
title: "Genetic endowments and wealth inequality"
subtitle: |
  Daniel Barth, Nicholas W. Papageorge, Kevin Thom\
  Journal of Political Economy, 2020, vol. 128, No.4
author: "伊藤成朗"
engine: knitr
execute:
  echo: false
  freeze: auto
format:
  revealjs:
    standalone: true
    embed-resources: true
    slide-number: true
    incremental: true
    width: 1600
    height: 900
    preview-links: auto
    logo: GrootbergGiraffeHead.jpg
    footer: <http://seiroito.github.io/QuartoFiles/posts/BarthPapageorgeThom/BPT.html>
    theme: [sky, sky-add.scss]
    linkcolor: red
    code-line-numbers: true
    code-copy: true
    code-block-border-left: "#31BAE9"
    toc: true
    toc-depth: 1
    template-partials: 
      - ../toc-slide.html
    include-in-header: ../sky_add_FilePointer.html
    include-after-body: ../toc-add.html
    toc-bg-image: "../../GrootbergGiraffe3.jpg"
lightbox: auto
filters:
  - parse-latex
  - imagify
imagify:
  header-includes: |
    \usetikzlibrary{calc}
    \usetikzlibrary{positioning}
    \usetikzlibrary{math}
    \usetikzlibrary{decorations.pathmorphing}
    \usetikzlibrary{shapes}
    \usetikzlibrary{arrows}
    \usetikzlibrary{backgrounds}
    \usetikzlibrary{fit}
    \usepackage{pgfplots}
  pdf-engine: lualatex
include-in-header:
  - text: |
      <style>
       .title {
        font-size: 2.0em;
      }
      </style>
#### From https://github.com/quarto-dev/quarto-cli/discussions/2951
#### I am happy with the left-indenting of the section line
#### Horizontal centering can be challenging (!) in quarto: stupid
#### https://github.com/quarto-dev/quarto-cli/issues/1231
#### It turns out an easy solution is :::{style="text-align: center;"}
  - text: |
      <style>
      .center-xy {
        margin: 0;
        position: absolute;
        top: 50%;
        left: 20%;
        -ms-transform: translateY(-50%), translateX(-50%);
        transform: translateY(-50%), translateX(-50%);
      }
      </style>
bibliography: c:/seiro/docs/seiro.bib
#### ../../../seiro.bib
resources:
  - BPT.pdf
# quarto render BPT.qmd
---
```{css, echo = F, results = "hide"}
Red {
  color: red;
}
Orange {
  color: #e94e2c;
}
Blue {
  color: blue;
}
.reveal .slides pre code {
    max-height: 800px;
}
#TOC {
  font-color: #e94e2c;
}
.big-code{
  font-size: 110%  
}
#### Below is suggested by Claude
#### .small-font-list {
####   font-size: 70%;
####   line-height: 1.2;
#### }
#### .small-font-list {
####   font-size: 70%;
####   line-height: 1.2;
#### }
#### .small-font-list ol {
####   list-style-type: none;
####   padding-left: 0;
####   margin-left: 0;
####   counter-reset: item;
#### }
#### .small-font-list ol li {
####   margin-bottom: 0.5em;
####   display: flex;
#### }
#### .small-font-list ol li::before {
####   content: counter(item) ". ";
####   counter-increment: item;
####   margin-right: 0.5em;
####   flex-shrink: 0;
#### }
#### .small-font-list ol li > p:first-child {
####   margin: 0;
####   display: inline;
#### }
.small-font-list {
  font-size: 70%;
  line-height: 1.2;
}
.small-font-list ol {
  list-style-type: none;
  padding-left: 0;
  margin-left: 0;
  counter-reset: item;
}
.small-font-list ol li {
  counter-increment: item;
  display: flex;
  align-items: flex-start;
  margin-bottom: 0.5em;
}
.small-font-list ol li::before {
  content: counter(item) ". ";
  flex: 0 0 2.5em;  /* Adjust this value to change the width of the number area */
  text-align: right;
  margin-right: 0.5em;
}
#### description list by Claude
/* Main container for the description list */
.description-list {
  font-size: 50% !important;  /* Reduces font size to 70% of the parent element */
  line-height: 1.2 !important;  /* Sets line height to 1.2 times the font size */
}
/* Styles for the description list element */
.description-list dl {
  margin-left: 0;  /* Removes default left margin */
  padding-left: 0;  /* Removes default left padding */
}
/* Styles for the term (header) in the description list */
.description-list dt {
  font-weight: bold;  /* Makes the term text bold */
  display: inline;  /* Displays the term inline. Uncomment for a line break after item header. */
  margin-right: 0.5em;  /* Adds some space after the term */
}
/* Styles for the description in the description list */
.description-list dd {
  display: inline;  /* Displays the description inline. Uncomment for a line break after item header. */
  margin-left: 0;  /* Removes default left margin */
}
/* Styles for paragraphs within the description */
.description-list dd p {
  display: inline;  /* Keeps paragraph content inline. Uncomment for a line break after item header. */
  margin: 0;  /* Removes default paragraph margins */
}
/* Creates a new line and adds space between list items */
.description-list dd::after {
  content: "";  /* Necessary for the pseudo-element to render */
  display: block;  /* Creates a line break */
  margin-bottom: 0.5em;  /* Adds space between items. Adjust as needed */
}
#### description with a line break between item header and texts
.description-lb {
  font-size: 70%;
  line-height: 1.2;
}
.description-lb dl {
  margin-left: 0;
  padding-left: 0;
}
.description-lb dt {
  font-weight: bold;
  margin-bottom: 0.2em;
}
.description-lb dd {
  margin-left: 0;
  margin-bottom: 1em;  /* Adjust this value to control spacing between items */
}
.description-lb dd p {
  margin: 0;
}
```

----

Genome

::::: {layout="[10,-2,20]" layout-valign="bottom"}

::::{.fragment}
![Chromosone...DNA](DNAMacrostructure.jpg)

::: {style="font-size: 70%;"}
<https://courses.lumenlearning.com/suny-ulster-ap1/chapter/the-nucleus-and-dna-replication/>
:::
::::

::::{.fragment}
![DNA structure](DNAStructure.jpg)

::: {style="font-size: 70%;"}
<https://www.haibunda.com/bundapedia/20230214003129-210-297249/deoxyribonucleic-acid--dna->
:::
::::

:::::

----

<https://mammothmemory.net/biology/dna-genetics-and-inheritance/dna-base-pairing/dna-double-helix.html>

<https://studylib.net/doc/5786293/snp-presentation>

----

### 2024-10-14 posts/BarthPapageorgeThom/BPT.qmd
Added sky_add_FilePointer.html to apply, if want to, sky_add.css to all quarto posts
Added sky_add_FilePointer.html to apply, if want to, seiro.css to all quarto posts

@@
 ---
-title: "Genetic endowments and wealth inequality"
+title: |
+  | Genetic endowments
+  | and wealth inequality
 subtitle: |
   Daniel Barth, Nicholas W. Papageorge, Kevin Thom\
-  Journal of Political Economy, 2020, vol. 128, No.4
+  *Journal of Political Economy*, 2020, vol. 128, No.4
 author: "伊藤成朗"
@@
     footer: <http://seiroito.github.io/QuartoFiles/posts/BarthPapageorgeThom/BPT.html>
-    theme: [sky, sky-add.scss]
+    theme: [sky]
     linkcolor: red
@@
       - ../toc-slide.html
-    include-in-header: ../sky_add_FilePointer.html
-    include-after-body: ../toc-add.html
+    includes:
+      in_header: ../sky_add_FilePointer.html
+      after-body: ../toc-add.html
     toc-bg-image: "../../GrootbergGiraffe3.jpg"
@@
   - parse-latex
-  - imagify
-imagify:
-  header-includes: |
-    \usetikzlibrary{calc}
-    \usetikzlibrary{positioning}
-    \usetikzlibrary{math}
-    \usetikzlibrary{decorations.pathmorphing}
-    \usetikzlibrary{shapes}
-    \usetikzlibrary{arrows}
-    \usetikzlibrary{backgrounds}
-    \usetikzlibrary{fit}
-    \usepackage{pgfplots}
-  pdf-engine: lualatex
+include-before-body:
+#### latex math shorthand
+  file: ../MathShorthand.html 
 include-in-header:
@@
 
+@BarthPapageorgeThom2020
+
 ----
 
-Genome
+## SNP, polygenic score
 
@@
 ::::{.fragment}
+Genome
+
 ![Chromosone...DNA](DNAMacrostructure.jpg)
@@
 ::: {style="font-size: 70%;"}
-<https://courses.lumenlearning.com/suny-ulster-ap1/chapter/the-nucleus-and-dna-replication/>
+[https://courses.lumenlearning.com/...](https://courses.lumenlearning.com/suny-ulster-ap1/chapter/the-nucleus-and-dna-replication)
 :::
@@
 ::: {style="font-size: 70%;"}
-<https://www.haibunda.com/bundapedia/20230214003129-210-297249/deoxyribonucleic-acid--dna->
+_[https://www.haibunda.com/...](https://www.haibunda.com/bundapedia/20230214003129-210-297249/deoxyribonucleic-acid--dna-)_
 :::
@@
 
-<https://mammothmemory.net/biology/dna-genetics-and-inheritance/dna-base-pairing/dna-double-helix.html>
+::::: {layout="[10,-2,20]" layout-valign="bottom"}
 
-<https://studylib.net/doc/5786293/snp-presentation>
+::::{.fragment}
+![Double helix and rung](DoubleHelixRung.jpg)
+
+::: {style="font-size: 70%;"}
+[https://mammothmemory.net/...](https://mammothmemory.net/biology/dna-genetics-and-inheritance/dna-base-pairing/dna-double-helix.html)
+:::
+::::
+
+::::{.fragment}
+![SNP](SNP.png)
+
+::: {style="font-size: 70%;"}
+_[https://studylib.net/...](https://studylib.net/doc/5786293/snp-presentation)_
+:::
+::::
+
+:::::
 
 ----
+
+<!-- tree.nathanfriend.io -->
+
+```{md}
+.
+└── Chromosome
+    └── DNA
+        └── Gene
+            └── Genome
+```
+
+:::::{.columns}
+
+::::{.column width="55%"}
+![](genome.jpg)
+
+::: {style="font-size: 70%;"}
+[_Source: L. C. Douglas_](https://historiesofthingstocome.blogspot.com/2010/08/age-of-genome-4.html){height=5cm}
+:::
+::::
+
+::::{.column width="45%"}
+* Chromosomes: 1/2 each from both parents 
+* DNA: Rolled up and shapes like an X
+* Gene: DNA segments
+* Genome: 3 billion nucletiodes (letters)
+   * Letter sequence = genome sequence
+   * 1.5 billion base pairs
+::::
+
+:::::
+
+* Base pairs differ between people in less than 1% (15 million) locations  
+* Single-nucleotide polymorphism (SNP) = such location
+   * 2 chromosomes per person: AT-AT, GC-GC, AT-GC or 0, 2, 1 (reference = GC)
+
+----
+
+Twin studies
+
+* Estimate how much genetic factors *collectively* matter for **explaining variations** of a trait  
+   * Collectively: Unit = group level aggregation
+   * Explaining variations: Variance decomposition 
+* Do not reveal which SNP is correlated
+
+. . .
+
+Genome wide association studies (GWASs)
+
+* Collect $J$ observable (?) SNPs of individual $i$.
+
+. . .
+
+Regress outcome $Y_{i}$ on each SNP using $J$ estimating equations:
+
+$$
+Y_{i}=\bfmu'\bfx_{i}+\beta_{j}SNP_{ij}+\epsilon_{ij}, \quad j=1,\dots,J.
+$$
+
+----
+
+Polygenic score of $i$ for the outcome $Y$ = *"Educational attainment (EA) score"*
+
+$$
+PGS_{i}=\sum_{j=1}^{J}\tilde{\beta} SNP_{ij}
+$$
+
+* Use Bayesian LDpred procedure to correct for correlations in $\tilde{\beta}_{j}$
+
+* Use all SNPs: Better out-of-sample results than using only SNPs with genome-wide significance $p$ value $< 5*10^{-8} =$ .0000005%
+
+* PGS is considered to be a predictor of individual fixed effects
+   * HRS sample: covariance with EA score / total schooling years variance = 10.6%
+   * Edu attainment SNPs $\propto$ biological process of brain development, cognition [@Okbay2016; @Lee2018]
+   * Edu attainment SNPs $\propto$ cognition SNPs [@Okbay2016]
+
+----
+
+## Interpretation
+
+* Correlation, not causality: Outcomes &lArr; gene + environment (endogenous)
+* Attenuation due to measurement errors in $\hat{\beta}_{j}$
+* Linearity assumption &rarr; underestimation of gene contributions (relative to twin studies)
+* Population stratification: Ethnic-group fixed effects &lArr; gene, environment  
+   * Including principal components of SNP data as covariates is shown to controll geographic variations
+* External validity is up to HRS ancestral composition
+
+----
+
+## Data
+
+Health and retirement survey: Age > 50 + partners, genetic samples 2006, 2008
+
+* All households with at least one individual with genetic European 
+* "Retired households" in 1996, 1998, 2002, 2004, 2006, 2008, 2010
+* Households with 1-2 members, drop same sex HHs
+* Household-year observations of both members 65-75 years old
+
+. . .
+
+2590 HHs, 5701 HH-year observations (Table 1)
+
+----
+
+----

### 2024-10-14 posts/BarthPapageorgeThom/BPT.qmd
Added up to mortality section in BPT.html

@@
   font-size: 110%  
+}
+.divCenterCol
+{
+  text-align:center;
 }
@@
 
-::::{.column width="55%"}
+::::{.column width="45%"}
 ![](genome.jpg)
@@
 * Attenuation due to measurement errors in $\hat{\beta}_{j}$
-* Linearity assumption &rarr; underestimation of gene contributions (relative to twin studies)
+* Linearity assumption &rarr; likely underestimation of gene contributions (relative to twin studies)
 * Population stratification: Ethnic-group fixed effects &lArr; gene, environment  
-   * Including principal components of SNP data as covariates is shown to controll geographic variations
+   * Including first (10) principal components of SNP data as covariates is shown to control geographic variations
 * External validity is up to HRS ancestral composition
@@
 
-Health and retirement survey: Age > 50 + partners, genetic samples 2006, 2008
-
+1. Health and retirement survey: Age > 50 + partners, genetic samples 2006, 2008 
    * All households with at least one individual with genetic European 
@@
    * Household-year observations of both members 65-75 years old
+   * Self-reported earnings  
+1. Social security administration data  
+   * Income data, top coded
 
@@
 
+> our sample comprises households for whom wealth data are most likely to be both accurate and
+comprehensive.
+
+* All pension, annuity, social security income, defined-contribution retirement plans 
+* Net value of housing, private businesses, vehicles, financial assets (cash, checking account, saving account, CD, stocks, mutual funds, trusts, other)
+* Winsorize at 1st and 99th percentiles of log real total wealth
+
 ----
+
+## Selection biases
+
+1. Into genotyping: Older, more educated, more females, wealthier [(Table S1)](BarthPapgergeThome_SNPWealth_JPE2020Appendix.pdf#page=5) 
+   * Education $\propto$ genotyping &rArr; individuals with low EA scores have higher than average education  &rArr; attenuates EA score $\propto$ education gradient 
+1. Of using retired households: More precise wealth but smaller sample size/more selection [(Table 3)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=17) 
+   * Selection results in more males, older in high EA scores &lArr; (Male) higher EA scorers live longer, but modest in size (Panel A)
+   * Age band of retired HHs &uarr; than main sample &rArr; q4-q1 differences (more selection) &uarr; &rArr; EA score $\propto$ education gradient &uarr; (Panel A vs. B)
+   * $+$non-retired HHs &rArr; q4-q1 differences &darr; (less selection) &rArr; EA score $\propto$ education gradient &darr; (Panel A vs. C)
+   * Smaller degree of selection leads to smaller q4-q1 differences
+   * But larger measurement errors in wealth ($\because$ pensions, etc. are not measured)
+
+----
+
+> However, the magnitudes of these differences are similar and relatively modest across alternate samples.
+Restricting our sample to retired households balances concerns about sample selection and measurement error.
+[(p.19)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=19)
+
+. . .
+
+Meaning:
+
+* We could have used a wider age band retired HH sample 
+   * More precision (larger sample size, smaller measurement errors) 
+   * Larger EA-edu gradient
+* But we did not
+   * Feel free to accept our results!
+
+----
+
+## Results
+
+::::: {.columns}
+
+::: {.column width="5%"}
+:::{style="font-size: 45%;line-height: 1.1;"}
+Fig 2A
+:::
+:::
+::: {.column width="25%"}
+![](Fig2A.jpg){width="70%"}
+:::
+::: {.column width="70%"}
+EA score $\propto$ wealth
+
+*  EA score -1 to 1: $+$ $20,000
+
+:::{.fragment}
+Table 2 also shows EA score $\propto$ wealth ($475K, q4-q1), lifetime labor income ($380K, q4-q1)  
+:::
+:::
+
+::: {.column width="5%"}
+:::{.fragment}
+:::{style="font-size: 45%;line-height: 1.1;"}
+Fig 2B
+:::
+:::
+:::
+::: {.column width="25%"}
+:::{.fragment}
+![](Fig2B.jpg){width="70%"}
+:::
+:::
+
+::: {.column width="70%"}
+:::{.fragment}
+EA score $\propto$ wealth similar between high schoolers vs college grads, up to q3
+:::
+:::
+
+::: {.column width="5%"}
+:::{.fragment}
+:::{style="font-size: 45%;line-height: 1.1;"}
+Tab 4
+:::
+:::
+:::
+::: {.column width="25%"}
+:::{.fragment}
+![](Tab4.jpg){width="50%"}
+:::
+:::
+
+::: {.column width="70%"}
+:::{.fragment}
+EA score $\propto$ wealth gradient: .246 (raw)&rarr;.070 (+edu)&rarr;.047 (+labor income)
+:::
+
+:::{.fragment}
+Gradient: "ability" + school quality &larr; income measure error
+:::
+:::
+
+:::::
+
+----
+
+Robust to:
+
+::::: {.columns}
+
+::: {.column width="50%"}
+* $+$Non financial respondent EA scores
+   * NFR EA scores is not partial-correlated
+* $+$SSA income (more objective)
+* $+$Non-retired HHs  
+* $+$Self-reported income of HRS
+:::
+
+::: {.column width="40%"}
+<div class="divCenterCol">
+![](Tab5.jpg){width="50%"}
+</div>
+:::
+:::::
+
+* Changes in sample and definitions: HRS sampling weights, only 1 HH-year per sample, only coupled HHs, different age, complicated function of incomes, excluding retirement and housing wealth, different version of EA score
+* Adding: cognitive ability, number of children, death of HH member, years since retirement
+
+----
+
+## Gene-environment correlations
+
+::::: {.columns}
+
+::: {.column width="50%"}
+Proxy of environment
+
+* Direct: Bequests
+* Indirect: Parental education
+
+::: {.fragment}
+Does not change the gene-wealth gradient
+:::
+
+* Parental investments are captured by respondent's schooling and labor incomes
+:::
+
+::: {.column width="40%"}
+<div class="divCenterCol">
+![](Tab6.jpg){width="50%"}
+</div>
+:::
+:::::
+
+----
+
+## Additional mechanisms
+
+EA score &uarr;
+
+::::: {.columns}
+
+::: {.column width="50%"}
+* One-year mortality &darr; for females 
+* Expected longeviity &uarr; for females 
+:::
+
+::: {.column width="40%"}
+<div class="divCenterCol">
+![](Tab7.jpg){width="50%"}
+</div>
+:::
+
+::: {.column width="50%"}
+* Risk aversion &darr; for females 
+* Expected longeviity &uarr; for females 
+:::
+
+::: {.column width="40%"}
+<div class="divCenterCol">
+![](Tab7.jpg){width="50%"}
+</div>
+:::
+
+:::::
+
+----

### 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
Added up to mortality section in BPT.html
Completed BPT.html for RG2024/10/15

@@
       </style>
-bibliography: c:/seiro/docs/seiro.bib
-#### ../../../seiro.bib
+bibliography: ../seiro.bib
+#### ../../../../../seiro.bib
+#### copy c:/seiro/docs/seiro.bib ../seiro.bib
 resources:
   - BPT.pdf
-# quarto render BPT.qmd
+#### quarto render BPT.qmd
 ---
@@
 
-@BarthPapageorgeThom2020
+@BarthPapageorgeThom2020 take aways
+
+* Gene-wealth gradient exists: EA score predicts wealth at retirement
+* Gene-wealth gradient&lArr; education, income, bequests, parents’ education, but does not explain fully
+* Gene-wealth gradient&rarr;0 when stock ownership is added
+   * EA score $\propto$ risk torelance, longevity, stock ownership, precision in macroeconomic risk perception
+* HHs without DB pension: gene-wealth gradient&uarr;
 
@@
 
-```{md}
+```
 .
@@
 * Winsorize at 1st and 99th percentiles of log real total wealth
+
+<br>
+
+Standard controls
+: age, sex of FR, calendar time, family member age-sex dummies and interactions
+
+Full education controls
+: education-age-sex dummies and interactions
 
@@
 
-EA score &uarr;
+EA score&uarr;
 
@@
 
-::: {.column width="50%"}
-* One-year mortality &darr; for females 
-* Expected longeviity &uarr; for females 
+::: {.column width="60%"}
+* One-year mortality&darr; for females 
+* Expected longeviity&uarr; for females 
 :::
 
-::: {.column width="40%"}
+::: {.column width="30%"}
 <div class="divCenterCol">
@@
 
-::: {.column width="50%"}
-* Risk aversion &darr; for females 
-* Expected longeviity &uarr; for females 
+::: {.column width="60%"}
+* Elicited risk aversion&darr; in gambles on incomes, business, but not on inheritance 
 :::
 
-::: {.column width="40%"}
+::: {.column width="30%"}
 <div class="divCenterCol">
-![](Tab7.jpg){width="50%"}
+![](Tab8.jpg){width="50%"}
+</div>
+:::
+
+::: {.column width="60%"}
+* Stock ownership&uarr; but not of other assets 
+* EA score $\propto$ wealth gradient &rarr;0 if we add stock ownership  
+:::
+
+::: {.column width="30%"}
+<div class="divCenterCol">
+![](Tab9.jpg){width="50%"}
 </div>
@@
 ----
+
+EA score&uarr;
+
+::::: {.columns}
+
+::: {.column width="60%"}
+* Deviation from objective probability&darr; 
+* Indication of complete certainty&darr;  
+* Financial planning time horizon&uarr; (viz. a few months) (except longer than 10 years)
+   * Better able to think about complex and abstract decisions
+   * @GabaixLaibson2022: Precise signals &rArr; behaves as if more patient
+:::
+
+::: {.column width="30%"}
+<div class="divCenterCol">
+![](Tab10.jpg){width="90%"}
+</div>
+:::
+
+::: {.column width="60%"}
+* Elicited stock returns&uarr; &rArr; stock ownership&uarr;, wealth&uarr; 
+:::
+
+::: {.column width="30%"}
+<div class="divCenterCol">
+![](TabS9.jpg){width="50%"}
+</div>
+:::
+
+::: {.column width="60%"}
+* Robust to adding cognitive test scores 
+:::
+
+::: {.column width="30%"}
+<div class="divCenterCol">
+![](TabS15.jpg){width="40%"}
+</div>
+:::
+
+:::::
+
+----
+
+::::: {.columns}
+
+::: {.column width="60%"}
+* EA score $\propto$ wealth, conditional on DB pension participation
+* EA score $\propto$ wealth&uarr; if not participating in DB pension
+:::
+
+::: {.column width="30%"}
+<div class="divCenterCol">
+![](Tab11.jpg){width="90%"}
+</div>
+:::
+
+:::::
+
+:::{.fragment}
+Gene-wealth gradient &lArr; a household's <Red>difficulty in making wise financial decisions</Red>
+:::
+
+----
+
+## Conclusions
+
+* Previous studies: EA scores $\propto$ education
+* This study: EA scores $\propto$ wealth
+* EA scores $\propto$ probabilistic thinking &rArr; human capital accumulation heterogeneity  
+   * Teaching probability may make human capital more homogenous?
+   * Not necessarily, gene-wealth gradient is a biased estimate &larr; environment
+   * But it hints something can be done there
+
+. . .
+
+::: {style="font-size: 80%;"}
+   > For example, it may be that children with lower polygenic scores begin to face challenges at particular ages or struggle to meet specific educational milestones.
+In that case,we could better target educational policies to help alleviate these road-blocks.
+:::
+
+こうした学習の遅れが格差の源泉と認識できるようになった
+
+----
+
+## Conclusions
+
+* "Ignoring heterogeneity in EA scores may place too much weight on education and inheritance"
+   * Using HH (parent-multiple children) fixed effect estimator should prevent this? 
+* Intervention changes the nature contribution (gradient size): DB pensions
+
+----
+
+## 感想
+
+* Gene-wealth gradient = corr[PGS, wealth|endogenous variables]  
+* 偏相関が小さくなるポイント: 株式所有 &lArr; longer time horizon, optimism
+* 金融教育?
+* 株式所有と確定給付年金を加えたら偏相関は?
+
+----

### 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
Completed BPT.html for RG2024/10/15

@@
 
+* 「貧しさは遺伝する」?...PGSを使って資産との相関を計測した研究
+* 強い根拠はないにせよ、介入ポイントを議論したのは素晴らしい
 * Gene-wealth gradient = corr[PGS, wealth|endogenous variables]  
 * 偏相関が小さくなるポイント: 株式所有 &lArr; longer time horizon, optimism
+    * 株式所有と確定給付年金を加えたら偏相関は?
 * 金融教育? 
-* 株式所有と確定給付年金を加えたら偏相関は?
+* 確定給付年金をもっと広く手厚く(できるのか?)
 

### 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
Completed BPT.html for RG2024/10/15

--- 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
@@
 
-* Gene-wealth gradient = corr[PGS, wealth|endogenous variables]  
+* 「貧しさは遺伝する」?...PGSを使って資産との相関を計測した研究
+* 強い根拠はないにせよ、介入ポイントを議論したのは素晴らしい
+* Gene-wealth gradient = corr[PGS, wealth|(endogenous) variables]  
 * 偏相関が小さくなるポイント: 株式所有 &lArr; longer time horizon, optimism
+* 偏相関が大きくなるポイント: 確定給付年金
+    * 株式所有と確定給付年金を加えたら偏相関は?
 * 金融教育? 
-* 株式所有と確定給付年金を加えたら偏相関は?
+* 確定給付年金をもっと広く手厚く(できるのか?)
 

### 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
Completed BPT.html for RG2024/10/15

@@
 * Gene-wealth gradient&rarr;0 when stock ownership is added
-   * EA score $\propto$ risk torelance, longevity, stock ownership, precision in macroeconomic risk perception
+   * EA score $\SIM$ risk torelance, longevity, stock ownership, precision in macroeconomic risk perception
 * HHs without DB pension: gene-wealth gradient&uarr;
@@
 
-![Chromosone...DNA](DNAMacrostructure.jpg)
+![](DNAMacrostructure.jpg){height="100%"}
 
@@
 ::::{.fragment}
-![DNA structure](DNAStructure.jpg)
+![](DNAStructure.jpg){height="80%"}
 
@@
    * HRS sample: covariance with EA score / total schooling years variance = 10.6%
-   * Edu attainment SNPs $\propto$ biological process of brain development, cognition [@Okbay2016; @Lee2018]
-   * Edu attainment SNPs $\propto$ cognition SNPs [@Okbay2016]
+   * Edu attainment SNPs $\SIM$ biological process of brain development, cognition [@Okbay2016; @Lee2018]
+   * Edu attainment SNPs $\SIM$ cognition SNPs [@Okbay2016]
 
@@
 1. Into genotyping: Older, more educated, more females, wealthier [(Table S1)](BarthPapgergeThome_SNPWealth_JPE2020Appendix.pdf#page=5) 
-   * Education $\propto$ genotyping &rArr; individuals with low EA scores have higher than average education  &rArr; attenuates EA score $\propto$ education gradient 
-1. Of using retired households: More precise wealth but smaller sample size/more selection [(Table 3)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=17) 
+   * Education $\SIM$ genotyping &rArr; individuals with low EA scores have higher than average education  &rArr; attenuates EA score $\SIM$ education gradient
+
+----
+
+## Selection biases
+
+2. Of using retired households: More precise wealth but smaller sample size/more selection [(Table 3)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=17) 
    * Selection results in more males, older in high EA scores &lArr; (Male) higher EA scorers live longer, but modest in size (Panel A)
-   * Age band of retired HHs &uarr; than main sample &rArr; q4-q1 differences (more selection) &uarr; &rArr; EA score $\propto$ education gradient &uarr; (Panel A vs. B)
-   * $+$non-retired HHs &rArr; q4-q1 differences &darr; (less selection) &rArr; EA score $\propto$ education gradient &darr; (Panel A vs. C)
+   * Age band of retired HHs &uarr; than main sample &rArr; q4-q1 differences (more selection) &uarr; &rArr; EA score $\SIM$ education gradient &uarr; (Panel A vs. B)
+   * $+$non-retired HHs &rArr; q4-q1 differences &darr; (less selection) &rArr; EA score $\SIM$ education gradient &darr; (Panel A vs. C)
    * Smaller degree of selection leads to smaller q4-q1 differences
@@
 ::: {.column width="70%"}
-EA score $\propto$ wealth
+EA score $\SIM$ wealth
 
@@
 :::{.fragment}
-Table 2 also shows EA score $\propto$ wealth ($475K, q4-q1), lifetime labor income ($380K, q4-q1)  
+Table 2 also shows EA score $\SIM$ wealth ($475K, q4-q1), lifetime labor income ($380K, q4-q1)  
 :::
@@
 :::{.fragment}
-EA score $\propto$ wealth similar between high schoolers vs college grads, up to q3
+EA score $\SIM$ wealth similar between high schoolers vs college grads, up to q3
 :::
@@
 :::{.fragment}
-EA score $\propto$ wealth gradient: .246 (raw)&rarr;.070 (+edu)&rarr;.047 (+labor income)
+EA score $\SIM$ wealth gradient: .246 (raw)&rarr;.070 (+edu)&rarr;.047 (+labor income)
 :::
@@
 * Stock ownership&uarr; but not of other assets 
-* EA score $\propto$ wealth gradient &rarr;0 if we add stock ownership  
+* EA score $\SIM$ wealth gradient &rarr;0 if we add stock ownership  
 :::
@@
 ::: {.column width="60%"}
-* EA score $\propto$ wealth, conditional on DB pension participation
-* EA score $\propto$ wealth&uarr; if not participating in DB pension
+* EA score $\SIM$ wealth, conditional on DB pension participation
+* EA score $\SIM$ wealth&uarr; if not participating in DB pension
 :::
@@
 
-* Previous studies: EA scores $\propto$ education
-* This study: EA scores $\propto$ wealth
-* EA scores $\propto$ probabilistic thinking &rArr; human capital accumulation heterogeneity  
+* Previous studies: EA scores $\SIM$ education
+* This study: EA scores $\SIM$ wealth
+* EA scores $\SIM$ probabilistic thinking &rArr; human capital accumulation heterogeneity  
    * Teaching probability may make human capital more homogenous?

### 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
Completed BPT.html for RG2024/10/15

@@
 :::{.fragment}
-Gradient: "ability" + school quality &larr; income measure error
+Gradient: "savvy" + school quality &larr; income meas errors
 :::
@@
 * Previous studies: EA scores $\SIM$ education
-* This study: EA scores $\SIM$ wealth
+* This study: EA scores $\SIM$ wealth|education, parents
 * EA scores $\SIM$ probabilistic thinking &rArr; human capital accumulation heterogeneity  

### 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
Completed BPT.html for RG2024/10/15

@@
 
-* Previous studies: EA scores $\SIM$ education
-* This study: EA scores $\SIM$ wealth|education, parents
+* Previous studies: $\corr$[EA scores, education]$>0$
+* This study: $\corr$[EA scores, wealth|education, parents]$>0$
 * EA scores $\SIM$ probabilistic thinking &rArr; human capital accumulation heterogeneity  
@@
 
-こうした学習の遅れが格差の源泉と認識できるようになった
+こうした学習の遅れが格差の源泉の可能性、と認識できるようになった
 
@@
 * 強い根拠はないにせよ、介入ポイントを議論したのは素晴らしい
-* Gene-wealth gradient = corr[PGS, wealth|(endogenous) variables]  
+* Gene-wealth gradient = $\corr$[PGS, wealth|(endogenous) variables]  
 * 偏相関が小さくなるポイント: 株式所有 &lArr; longer time horizon, optimism
@@
 * 確定給付年金をもっと広く手厚く(できるのか?)
+* 計量: 個人固定効果の中身をPGSとそれ以外に分解
 

### 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
Completed BPT.html for RG2024/10/15

@@
 
-1. Into genotyping: Older, more educated, more females, wealthier [(Table S1)](BarthPapgergeThome_SNPWealth_JPE2020Appendix.pdf#page=5) 
+1. Into genotyping: Older, more educated, more females, wealthier [(Table S1)](BarthPapageorgeThom_SNPWealth_JPE2020Appendix.pdf#page=5) 
    * Education $\SIM$ genotyping &rArr; individuals with low EA scores have higher than average education  &rArr; attenuates EA score $\SIM$ education gradient
@@
 
-2. Of using retired households: More precise wealth but smaller sample size/more selection [(Table 3)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=17) 
+2. Of using retired households: More precise wealth but smaller sample size/more selection [(Table 3)](BarthPapageorgeThom_SNPWealth_JPE2020.pdf#page=17) 
    * Selection results in more males, older in high EA scores &lArr; (Male) higher EA scorers live longer, but modest in size (Panel A)
@@
 Restricting our sample to retired households balances concerns about sample selection and measurement error.
-[(p.19)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=19)
+[(p.19)](BarthPapageorgeThom_SNPWealth_JPE2020.pdf#page=19)
 

### 2024-10-16 posts/BarthPapageorgeThom/BPT.qmd
Added gene-wealth gradient with education control in BPT.html

@@
     footer: <http://seiroito.github.io/QuartoFiles/posts/BarthPapageorgeThom/BPT.html>
+    css: ../seiro.css
     theme: [sky]
@@
 $$
-Y_{i}=\bfmu'\bfx_{i}+\beta_{j}SNP_{ij}+\epsilon_{ij}, \quad j=1,\dots,J.
+Y_{i}
+=
+\beta_{j}SNP_{ij}+\bfmu'_{j}\bfx_{i}+\epsilon_{ij}, \quad j=1,\dots,J.
 $$
+
+$\cov[SNP_{ij}, \epsilon_{ij}]> 0$: Parents see traits which may be correlated with $SNP_{ij}$ and decide on investments. Environment is endogenous to the trait in SNP regressions.
 
@@
    * Edu attainment SNPs $\SIM$ cognition SNPs [@Okbay2016]
+
+----
+
+Gene-wealth gradient $\gamma$ with education control
+
+$$
+\begin{aligned}
+W_{i}
+&=
+\bftheta'\bfx_{i}+\gamma PGS_{i}+\eta Y_{i}+e_{i},\\
+&=
+\bftheta'\bfx_{i}+\gamma PGS_{i}+\frac{\eta}{J}\left\{PGS_{i}+\sum_{j=1}^{J}\left(\bfmu'_{j}\bfx_{i}+\tilde{\epsilon}_{ij}\right)\right\}+e_{i},\\
+&=
+\bftheta'\bfx_{i}+\left(\gamma+\frac{\eta}{J}\right)PGS_{i}+\frac{\eta}{J}\sum_{j=1}^{J}\left(\bfmu'_{j}\bfx_{i}+\tilde{\epsilon}_{ij}\right)+e_{i},\\
+&\simeq
+\left(\bftheta'+\eta\bar{\bfmu}'\right)\bfx_{i}+\gamma PGS_{i}+\eta\bar{\tilde{\epsilon}}_{i}+e_{i}.
+\end{aligned}
+$$
+where
+$$
+JY_{i}
+=
+PGS_{i}+\sum_{j=1}^{J}\left(\bfmu'_{j}\bfx_{i}+\tilde{\epsilon}_{ij}\right).
+$$
+
+----
+
+Because $\cov[SNP_{ij}, \epsilon_{ij}]> 0$, in general, we have $\cov[PGS_{i}, \bar{\tilde{\epsilon}}_{i}]> 0$.
+
+. . .
+
+So, in general, $\plim \hat{\gamma}-\gamma>0$.
+
+. . .
+
+This is problematic: the paper tries to show persistence of positive $\hat{\gamma}$
+
+. . .
+
+Authors may have acknowledged it: Kept on cautioning that
+
+* Estimated relationship is not causal
+* Diverse mechanisms give rise to estimated relationship
+
+. . .
+
+A partial remedy: Use $\tilde{Y}_{i}=\tilde{\beta}_{j}SNP_{ij}+\bfmu'_{j}\bfx_{i}$ in place of $Y_{i}$ on RHS.
+
+This *partly* takes away the correlation between $GPS_{i}$ and the error term
+
+Because $\cov[SNP_{ij}, \epsilon_{ij}]> 0$, $\tilde{\beta}_{j}$ is overestimated, so deviation of $\hat{\tilde{Y}}_{i}$ from real $\tilde{Y}_{i}$ is larger for individual with large $SNP_{ij}$.
+Although partly removed, there still remain a positive correlation between PGS and the error term.
 

### 2025-01-07 posts/RustichiniIaconoLeeMcGue/RILM_PGSEdu_JPE2023.qmd
initial commit

---
title: "Educational attainment and intergenerational mobility: A polygenic score analysis"
subtitle: Aldo Rustichini, William G. Iacono, James J. Lee, and Matt McGue|
  Journal of Political Economy 2023 131:10, 2724-2779
author: "伊藤成朗"
engine: knitr
execute:
  echo: false
  freeze: auto
format:
  revealjs:
    standalone: true
    embed-resources: true
    slide-number: true
    incremental: true
    width: 1600
    height: 900
    preview-links: auto
    logo: GrootbergGiraffeHead.jpg
    footer: <http://seiroito.github.io/RILM_PGSEdu_JPE2023.html>
    css: ../../../../seiro.css
    #linkcolor: red#marked off because sky-add sets it to DarkGreen
    theme: [sky, sky-add]
    code-line-numbers: true
    code-copy: true
    code-block-border-left: "#31BAE9"
    toc: true
    toc-depth: 1
    template-partials: 
      - ../../../../toc-slide.html
    include-after-body: ../../../../toc-add.html
    toc-bg-image: "../../GrootbergGiraffe3.jpg"
lightbox: 
  match: auto
  desc-position: bottom
filters:
  - parse-latex
include-in-header:
  - text: |
      <style>
       .title {
        font-size: 2.0em;
      }
      </style>
#### From https://github.com/quarto-dev/quarto-cli/discussions/2951
#### I am happy with the left-indenting of the section line
#### Horizontal centering can be challenging (!) in quarto: stupid
#### https://github.com/quarto-dev/quarto-cli/issues/1231
#### It turns out an easy solution is :::{style="text-align: center;"}
  - text: |
      <style>
      .center-xy {
        margin: 0;
        position: absolute;
        top: 50%;
        left: 20%;
        -ms-transform: translateY(-50%), translateX(-50%);
        transform: translateY(-50%), translateX(-50%);
      }
      </style>
include-before-body:
#### latex math shorthand
  file: ../../../../MathShorthand.html 
bibliography: c:/seiro/docs/seiro.bib
#### ../../../../seiro.bib
resources:
  - RILM_PGSEdu_JPE2023.pdf
# quarto render RILM_PGSEdu_JPE2023.qmd
---
```{css, echo = F, results = "hide"}
TallerOutput {
  height: 800px;
}
.small-font-list {
  font-size: 70%;
  line-height: 1.2;
}
.small-font-list ol {
  list-style-type: none;
  padding-left: 0;
  margin-left: 0;
  counter-reset: item;
}
.small-font-list ol li {
  counter-increment: item;
  display: flex;
  align-items: flex-start;
  margin-bottom: 0.5em;
}
.small-font-list ol li::before {
  content: counter(item) ". ";
  flex: 0 0 2.5em;  /* Adjust this value to change the width of the number area */
  text-align: right;
  margin-right: 0.5em;
}
#### description list by Claude
/* Main container for the description list */
.description-list {
  font-size: 50% !important;  /* Reduces font size to 70% of the parent element */
  line-height: 1.2 !important;  /* Sets line height to 1.2 times the font size */
}
/* Styles for the description list element */
.description-list dl {
  margin-left: 0;  /* Removes default left margin */
  padding-left: 0;  /* Removes default left padding */
}
/* Styles for the term (header) in the description list */
.description-list dt {
  font-weight: bold;  /* Makes the term text bold */
  display: inline;  /* Displays the term inline. Uncomment for a line break after item header. */
  margin-right: 0.5em;  /* Adds some space after the term */
}
/* Styles for the description in the description list */
.description-list dd {
  display: inline;  /* Displays the description inline. Uncomment for a line break after item header. */
  margin-left: 0;  /* Removes default left margin */
}
/* Styles for paragraphs within the description */
.description-list dd p {
  display: inline;  /* Keeps paragraph content inline. Uncomment for a line break after item header. */
  margin: 0;  /* Removes default paragraph margins */
}
/* Creates a new line and adds space between list items */
.description-list dd::after {
  content: "";  /* Necessary for the pseudo-element to render */
  display: block;  /* Creates a line break */
  margin-bottom: 0.5em;  /* Adds space between items. Adjust as needed */
}

#### description with a font size control (not effective)
/* Main container for the description list */
.des-list-boxed {
    font-size: 60% !important;  /* Default font size */
    line-height: 1.2 !important;  /* Sets line height to 1.2 times the font size */
}
/* Specific font size override */
.des-list-boxed[data-font-size] {
    font-size: attr(data-font-size string, 60%) !important;
}
.des-list-boxed dl {
    margin-left: 0;  /* Removes default left margin */
    padding-left: 0;  /* Removes default left padding */
    width: 100%;  /* Ensures the container takes full width */
}
.des-list-boxed dt {
    font-weight: bold;  /* Makes the term text bold */
    text-align: right;  /* Right aligns the term text */
    padding-right: 1em;  /* Adds spacing after the term */
    float: left;  /* Allows description to wrap around header */
    white-space: nowrap;  /* Prevents item header from wrapping */
}
.des-list-boxed dd {
    margin-left: 0;  /* Removes default left margin */
    overflow: hidden;  /* Clears float */
}
.des-list-boxed dd p {
    margin: 0;  /* Removes default paragraph margins */
    word-wrap: break-word;  /* Allows description to wrap */
    overflow-wrap: break-word;  /* Modern alternative to word-wrap */
}
.des-list-boxed dd p:first-child {
    /* First line of description */
    /* No special styling needed, will naturally wrap around header */
}
.des-list-boxed dd p:not(:first-child) {
    /* Lines after the first line */
    display: block;  /* Ensures full width */
    margin-left: 1em;  /* Indentation for subsequent lines */
}
.des-list-boxed dd::after {
    content: "";  /* Necessary for the pseudo-element to render */
    display: block;  /* Creates a line break */
    margin-bottom: 0.5em;  /* Adds space between items */
    clear: both;  /* Clears the float */
}

#### description with a line break between item header and texts
.description-lb {
  font-size: 70%;
  line-height: 1.2;
}
.description-lb dl {
  margin-left: 0;
  padding-left: 0;
}
.description-lb dt {
  font-weight: bold;
  margin-bottom: 0.2em;
}
.description-lb dd {
  margin-left: 0;
  margin-bottom: 1em;  /* Adjust this value to control spacing between items */
}
.description-lb dd p {
  margin: 0;
}
gray {
  color: #aeb6bf;
}
```

## Summary of [this paper](https://doi.org/10.1086/724860)

* US Minnesota Twin Study Sample
* Build a genetic model (with better genetics foundation than **ad hoc** econ standard model)
* Intergenerational elaciticity of income is predicted and estimated to be larger in genetic models
* Empirical pathway: Through parental education and incomes, parental PGS has no direct effect
* Parental PGS &nrArr; educational attainment, after controling for parental education and incomes

----

:::{style="text-align: center;"}
# Background
:::

----

----

## SNP, polygenic score

::::: {layout="[10,-2,20]" layout-valign="bottom"}

::::{.fragment}
Genome

![](DNAMacrostructure.jpg){height="100%"}

::: {style="font-size: 70%;"}
[https://courses.lumenlearning.com/...](https://courses.lumenlearning.com/suny-ulster-ap1/chapter/the-nucleus-and-dna-replication)
:::
::::

::::{.fragment}
![](DNAStructure.jpg){height="80%"}

::: {style="font-size: 70%;"}
_[https://www.haibunda.com/...](https://www.haibunda.com/bundapedia/20230214003129-210-297249/deoxyribonucleic-acid--dna-)_
:::
::::

:::::

----

::::: {layout="[10,-2,20]" layout-valign="bottom"}

::::{.fragment}
![Double helix and rung](DoubleHelixRung.jpg)

::: {style="font-size: 70%;"}
[https://mammothmemory.net/...](https://mammothmemory.net/biology/dna-genetics-and-inheritance/dna-base-pairing/dna-double-helix.html)
:::
::::

::::{.fragment}
![SNP](SNP.png)

::: {style="font-size: 70%;"}
_[https://studylib.net/...](https://studylib.net/doc/5786293/snp-presentation)_
:::
::::

:::::

----

<!-- tree.nathanfriend.io -->

```
.
└── Chromosome
    └── DNA
        └── Gene
            └── Genome
```

:::::{.columns}

::::{.column width="40%"}
![](genome.jpg)

::: {style="font-size: 70%;"}
[_Source: L. C. Douglas_](https://historiesofthingstocome.blogspot.com/2010/08/age-of-genome-4.html){height=5cm}
:::
::::

::::{.column width="60%"}
* Chromosomes: 1/2 each from both parents 
* DNA: Rolled up and shapes like an X
* Gene: DNA segments
* Genome: 3 billion nucletiodes (letters)
   * Letter sequence = genome sequence
   * 1.5 billion base pairs (alelles, 対立遺伝子)
::::

:::::

* Base pairs differ between people in less than 1% (15 million) locations  
* Single-nucleotide polymorphism (SNP) = such location
   * 2 chromosomes per person &rarr; AT-AT, GC-GC, AT-GC or 0, 2, 1 (reference = GC)

----

Twin studies

* Estimate how much genetic factors *collectively* matter for **explaining variations** of a trait  
   * Collectively: Unit = group level aggregation
   * Explaining variations: Variance decomposition 
* Do not reveal which SNP is correlated

----

Genome wide association studies (GWASs)

* Collect $M$ observable (?) SNPs of (family $i$) individual $j$.

. . .

Regress outcome $y^{i}_{j}$ on $m$-th SNP for all $m=1, \dots, M$:

$$
y^{i}_{j}
=
\beta_{m}SNP^{i}_{jm}+\epsilon^{i}_{jm}, \quad m=1,\dots,M.
$$

Remember:

* Environment is endogenous to the trait in SNP regressions
   * $\cov[SNP^{i}_{jm}, \epsilon^{i}_{jm}]> 0$: Parents see traits which may be correlated with $SNP^{i}_{jm}$ and decide on investments 
   * Estimated $\beta_{m}$ should be upwardly biased
* No attempts to correct for this

----

Polygenic score = Predicted outcomes based on all relevant SNPs

$$
PGS^{i}_{j}=\sum_{m=1}^{M}\tilde{\beta}_{m} SNP^{i}_{jm}
$$

* Use Bayesian LDpred procedure to correct for correlations in $\tilde{\beta}_{m}$
  * @BarthPapageorgeThom2020 follow the convention and use all SNPs: Better out-of-sample results than using only SNPs with genome-wide significance $p$ value $< 5*10^{-8} =$ .0000005%

* PGS is considered to be a predictor of individual fixed effects
   * Edu attainment SNPs $\SIM$ biological process of brain development, cognition [@Okbay2016; @Lee2018]
   * Edu attainment SNPs $\SIM$ cognition SNPs [@Okbay2016]
   * HRS sample: cov[EA score, schooling years] / variance[schooling years] = 10.6%

----

## Intergenerational skill formation

* @BeckerTomes1979: AR(1)
* @Goldberger1989: Weighted average of all ancestors
* Both are incorrect, this paper shows there exists an invariant (stable) mapping father genotype distribution × mother genotype distribution $\to$ child genotype distribution, so skill formation depends on parental genotype distributions

----

```{mermaid}
%%| label: fig-mermaid
%%| fig-width: 6
%%| fig-cap: |
%%|   How parents affect child education outcomes
flowchart LR
  A1[father] --> A3(couple)
  A2[mother] --> A3(couple)
  A3[couple] --> B("child
   genotype")
  A3[couple] --"genetic nurture"--> C("child
   environment")
  B("child
   genotype") --> D("child
    education")
  C("child
   environment") <--"gene×environ"--> B("child
    genotype")
  C("child
   environment") --> D("child
    education")
```

----

:::{style="text-align: center;"}
# Estimation results 
:::

----

Controls used in estimation

* 10 principal components of PGS
* Parent-child age difference

----

::: {style="font-size: 70%;"}
**Data:** Minnesota Twin Study

* Cognitive measures to estimate IQ
   * Wechsler Intelligence Scale for Children-Revised (WISC-R)
   * Wechsler Adult Intelligence Scale-Revised (WAIS-R)
* Noncognitive measures 
   * MPQ (Mulitdimensional Personality Questionnaire)
      * PA (positive emotionality of affectivity; well-being, social potency, achievement, social closeness)
      * NA (negative emotionality of affectivity; stress reaction, alienation, aggression)
      * constraint (control, harm avoidance, traditionalism)
   * DSM-IV (Diagnostic and Statisitcal Manual of Mental Disorders, fourth edition)
      * Externalizing: defiant disorder, conduct disorder, adult antisocial behavior
      * Academic effort: 8 items answered by twin mothers
      * Academic problems: 3 items answered by twin mothers
* Family background
   * Hollingsworth scale on occupational status
   * Four year college degrees of parents
:::

----

:::::::::::{.columns}

::::{.column width="45%"}
![IGE](Tab1.jpg)
::::

::::{.column width="55%"}
* IGE = .13 is small relative to previous studies (.2 ~ .3)
   * Attenuation (measurement errors in family incomes)
   * European population has lower IGEs, Sweden = .125
   * Reduced by `r round((1-(.78/.134))*100, 0)`% when PGS and mediators are added
* Effects of PGS
   * = .078 in (2)
   * = 0 in (3) when mediators of PGS are added
::::

::::::::::

----

:::::::::::{.columns}

::::{.column width="45%"}
![recursive SEM](Tab2.jpg)
::::

::::{.column width="55%"}
Education years&uarr; with

* Cognitive ability, noncognitive ability
   * PGS of child, parents are not related when parental education, family income are added
* Parental income, education
   * Impacts are twice as large for parental education

Cognitive ability&uarr; with

* PGS

Noncognitive ability&uarr; with

* PGS, smaller magnitude
::::

::::::::::

----

----

:::::::::::{.columns}

::::{.column width="45%"}
![recursive SEM](Tab3.jpg)
::::

::::{.column width="55%"}
Dizygotic twins (fraternal twins)

Education years&uarr; with

* PGS of child and mother
   * PGS of father is unrelated
* Parental income, education
   * Magnitude is larger for parental education

Parental education years&uarr; with

* PGS of parents
   * Magnitude is larger for father PGS

Family income&uarr; with

* PGS of parents
   * Magnitude is larger for father PGS
::::

::::::::::

----

----

:::::::::::{.columns}

::::{.column width="25%"}
![](TabS6.jpg){width="50%"}

![](TabS7.jpg){width="50%"}

![](TabS8.jpg){width="50%"}

![](TabS9.jpg){width="50%"}
::::

::::{.column width="75%"}
* PGS of parents are not related to outcomes once IQ, soft skills index, parental education and family income are added
* Most parental impacts on education are through incomes and parental education
* No other indirect pathway remain from parental PGS, after parental education and incomes are addressed
   *  No genetic nurture
::::

::::::::::

----

感想

* Multiple testing気にせず

----

#### References

::: {#refs}
:::

### 2025-01-07 posts/RustichiniIaconoLeeMcGue/RILM_PGSEdu_JPE2023.qmd
initial commit

@@
 * US Minnesota Twin Study Sample
-* Build a genetic model (with better genetics foundation than **ad hoc** econ standard model)
-* Intergenerational elaciticity of income is predicted and estimated to be larger in genetic models
-* Empirical pathway: Through parental education and incomes, parental PGS has no direct effect
-* Parental PGS &nrArr; educational attainment, after controling for parental education and incomes
+* Build a genetic model of human capital investments with genotype transmission (with better genetics foundation than *ad hoc* standard econ models)
+* Intergenerational elaciticity of income is predicted to be larger in genetic models
+* Empirical pathway: Through parental education and incomes, parental PGS has no direct effect on child human capital
+* Child PGS is relevant
+* Genetic nurture (introverted genotype produces an environment to grow introverted child, independently of income and education) is not found
 
@@
 :::{style="text-align: center;"}
-# Estimation results 
+# Model 
 :::
@@
 ----
+
+::: {style="font-size: 80%;"}
+Family $i$ parents' problem with twins $j=1, 2$ of observable skills $\theta^{i}_{j}$
+$$
+\begin{aligned}
+\max_{\{E^{i}, I^{i}_{1}, I^{i}_{2}\}} \;\;\;&\E\left[(1-\delta)\ln E^{i}+\delta\left(y^{i}_{1}+y^{i}_{2}\right)\left|\theta^{i}_{1}, \theta^{i}_{2}\right.\right]\\
+\st \;\;\; & E^{i}+I^{i}_{1}+I^{i}_{2}=e^{y^{i}}
+\end{aligned}
+\tag{6}
+$$
+$$
+\begin{align}
+\scriptsize{\mbox{Human capital production function}} && 
+h^{i}_{j}&=\alpha_{I}\ln I^{i}_{j}+\alpha_{\theta}\theta^{i}_{j}+\epsilon^{\mathrm{h}i}_{j} && \phantom{mmmmmmmmmmmm} \tag{8}\label{eq8}\\
+\scriptsize{\mbox{Child income}} && 
+y^{i}_{j}&=\alpha_{\mathrm{h}}h^{i}_{j}+\epsilon^{\mathrm{y}i}_{j} &&\tag{9}\label{eq9}
+\end{align}
+$$
+Solution
+$$
+I^{i}=\frac{\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}{1-\delta+2\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}e^{y^{i}}
+\tag{11}
+$$
+Optimised child income
+$$
+y^{i}_{j}=\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}\ln 
+\frac{\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}{1-\delta+2\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}
++\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}y^{i}+\alpha_{\theta}\alpha_{\mathrm{h}}\theta^{i}_{j}+\alpha_{\mathrm{h}}\epsilon^{\mathrm{h}i}_{j}+\epsilon^{\mathrm{y}i}_{j}
+\tag{12}\label{eq12}
+$$ 
+:::
+
+----
+
+Matching of parents
+
+* Assume independence of genotype conditional on observable characteristics [incomes, skills $w(g)$]
+   * Matches are made on observables, not on genotype
+   * Matches are random within obserbable strata *"worth" classes*=$w_{\mathrm{y}}y+w(g)$]
+      * Genotype matching is random within a stratum
+   * Matches are nonrandom at the population due to population stratification based on observables
+
+. . .
+
+Genetic transmission: $\Delta$(father genotype) × $\Delta$(mother genotype) $\to$ $\Delta$(child genotype)
+
+The appendix D of paper shows an invariant mapping exists with properties shown in Theorem 3.1
+
+----
+
+Skills production function (genotype enters here, hence the parental genotypes)
+$$
+\theta^{i}_{j}=w\left(g^{i}_{j}\right)+\pi y^{i}+\bfPi\bfx^{i}_{j}+F^{i}+\epsilon^{\theta i}_{j}
+\tag{4}\label{eq4}
+$$
+Estimating equations: Substitute \eqref{4} into  \eqref{12} gives child income decomposition, and \eqref{11} into  \eqref{8} gives human capital decomposition
+$$
+\begin{align}
+y^{i}_{j}
+&=
+\alpha_{\theta}\alpha_{\mathrm{h}}w\left(g^{i}_{j}\right)
++\left(\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}+\alpha_{\theta}\alpha_{\mathrm{h}}\pi\right)y^{i}
++\alpha_{\theta}\alpha_{\mathrm{h}}F^{i}\\
+&\hspace{1em}
++\alpha_{\theta}\alpha_{\mathrm{h}}\bfPi\bfx^{i}_{j}
++\alpha_{\theta}\alpha_{\mathrm{h}}\epsilon^{\theta i}_{j}
++\alpha_{\mathrm{h}}\epsilon^{\mathrm{h} i}_{j}
++\epsilon^{\mathrm{y}i}_{j}\tag{21}\label{eq21}\\
+h^{i}_{j}
+&=
+\alpha_{\theta}w\left(g^{i}_{j}\right)
++\left(\alpha_{\mathrm{I}}+\alpha_{\theta}\pi\right)y^{i}
++\alpha_{\theta}F^{i}\\
+&\hspace{1em}
++\alpha_{\theta}\bfPi\bfx^{i}_{j}
++\alpha_{\theta}\epsilon^{\theta i}_{j}
++\epsilon^{\mathrm{y}i}_{j}\tag{23}\label{eq23}
+\end{align}
+$$
+
+----
+
+Can show IGE is smaller in "standard model" than genetic model
+
+::: {style="font-size: 80%;"}
+@BeckerTomes1979
+$$
+\begin{align}
+\theta_{t+1}
+&=
+\eta \theta_{t}+\epsilon^{\theta}_{t+1}\tag{26}\label{eq26}\\
+y^{i}_{t+1}
+&=
+\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}y^{i}_{t}+\alpha_{\theta}\alpha_{\mathrm{h}}\theta_{t+1}+\epsilon^{\mathrm{y}i}_{t+1}\tag{25}\label{eq25}
+\end{align}
+$$
+Genetic model of this paper assumes
+$$
+\begin{align}
+\theta
+&=
+w^{\theta}_{\mathrm{f}}\theta_{\mathrm{f}}+w^{\theta}_{\mathrm{m}}\theta_{\mathrm{m}}\tag{A4}\label{eqA4}\\
+y^{i}
+&=
+w^{\mathrm{y}}_{\mathrm{f}}y^{i}_{\mathrm{f}}+w^{\mathrm{y}}_{\mathrm{m}}y^{i}_{\mathrm{m}}\tag{A3}\label{eqA3}
+\end{align}
+$$
+Plugging these give
+$$
+\begin{align}
+\theta_{t+1}
+&=
+\eta \sum_{s=m,f}w^{\theta}_{s}\theta_{st}+\epsilon^{\theta}_{t+1}\tag{33}\label{eq33}\\
+y^{i}_{t+1}
+&=
+\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}\sum_{s=m,f}w^{\mathrm{y}}_{s}y^{i}_{st}+\alpha_{\theta}\alpha_{\mathrm{h}}\theta_{t+1}+\epsilon^{\mathrm{y}i}_{t+1}\tag{32}\label{eq32}
+\end{align}
+$$
+:::
+
+----
+
+:::{style="text-align: center;"}
+# Estimation strategy 
+:::
+
+----
+
+Structural equation model (SEM)
+
+* Structural equations
+* Measurement (error) equations
+
+. . .
+
+Identification conditions of structural parameters remain the same as in simultaneous equation models
+
+In this paper:
+
+* identification for only recursive model is discussed (Tab 2)
+* use of IVs is not discussed (Tab 3)
+
+. . .
 
@@
 * Parent-child age difference
+
+----
+
+Identification ideas
+
+1. Dizygotic twins [Table 1, 2]
+   * Share birth days, environment
+   * Differ in genotypes
+   * Ideal to see impacts of genotypes on outcomes
+
+----
+
+Identification ideas
+
+2. SEM
+   * Recursive system (37)-(42) are identified as usual [Table 2]
+   * Structural parameters of endogenous variables are recovered from \eqref{eq43}-\eqref{eq45} (how?) [Table 3]
+      * Given PGS of child, PGS of parents should be redundant, so long as they do not affect environment (gene-environment correlation, or "passive rGE")
+
+. . .
+
+::: {style="font-size: 80%;"}
+$$
+\begin{align}
+e^{i}_{\mathrm{h}}
+&=
+\alpha_{e_{\mathrm{h}}}+
+\gamma^{\mathrm{pGS}_{m}}_{e_{\mathrm{h}}}\mathrm{pGS}^{i}_{m}+\gamma^{\mathrm{pGS}_{f}}_{e_{\mathrm{h}}}\mathrm{pGS}^{i}_{f}+\xi_{e_{\mathrm{h}}}\tag{43}\label{eq43}\\
+y^{i}_{\mathrm{h}}
+&=
+\alpha_{y_{\mathrm{h}}}+\gamma^{\mathrm{pGS}_{m}}_{y_{\mathrm{h}}}\mathrm{pGS}^{i}_{m}+\gamma^{\mathrm{pGS}_{f}}_{y_{\mathrm{h}}}\mathrm{pGS}^{i}_{f}+\xi_{y_{\mathrm{h}}}\tag{44}\label{eq44}\\
+e^{i}_{j}
+&=
+\alpha_{e}+\gamma^{e_{\mathrm{h}}}_{\mathrm{e}}e_{\mathrm{h}}^{i}
++\gamma^{y_{\mathrm{h}}}_{\mathrm{e}}y_{\mathrm{h}}^{i}
++\gamma^{\mathrm{pGS}}_{\mathrm{e}}\mathrm{pGS}^{i}_{j}
++\gamma^{\mathrm{pGS}_{m}}_{\mathrm{e}}\mathrm{pGS}^{i}_{m}
++\gamma^{\mathrm{pGS}_{f}}_{\mathrm{e}}\mathrm{pGS}^{i}_{f}
++\xi_{\mathrm{e}}\tag{45}\label{eq45}
+\end{align}
+$$
+:::
 
@@
    * Four year college degrees of parents
+:::
+
+----
+
+:::{style="text-align: center;"}
+# Estimation results 
 :::
@@
 ::::{.column width="45%"}
-![recursive SEM](Tab3.jpg)
+![full SEM](Tab3.jpg)
 ::::
@@
 ::::{.column width="55%"}
-Dizygotic twins (fraternal twins)
+Dizygotic (fraternal) + monozygotic (identical) twins
 
@@
 * No other indirect pathway remain from parental PGS, after parental education and incomes are addressed
-   *  No genetic nurture
+   *  No passive rGE other than parental education and incomes (genetic nurture) is found
 ::::
@@
 
+* 咀嚼できていません
+* 人的資本投資を内生化したsteady state genetic transmission mappingの存在を証明
+* それと実証部分の結びつきが弱い(理解できていない)
+* 親の教育と所得を考慮すると、遺伝情報は子の教育に影響しないことを実証的に示したのは斬新
+* SEMはidentification assumptionが不明(recursiveモデル以外)、IVを使っていない?
 * Multiple testing気にせず

### 2025-01-08 posts/RustichiniIaconoLeeMcGue/RILM_PGSEdu_JPE2023.qmd
added RILM_PGSEdu_JPE2023.html

@@
     logo: GrootbergGiraffeHead.jpg
-    footer: <http://seiroito.github.io/RILM_PGSEdu_JPE2023.html>
-    css: ../../../../seiro.css
+    footer: <http://seiroito.github.io/QuartoFiles/posts/BarthPapageorgeThom/RILM_PGSEdu_JPE2023.html>
+    css: ../seiro.css
     #linkcolor: red#marked off because sky-add sets it to DarkGreen
-    theme: [sky, sky-add]
+    theme: [sky]
     code-line-numbers: true
@@
     template-partials: 
-      - ../../../../toc-slide.html
-    include-after-body: ../../../../toc-add.html
+      - ../toc-slide.html
+    includes:
+      in_header: ../sky_add_FilePointer.html
+      after-body: ../toc-add.html
     toc-bg-image: "../../GrootbergGiraffe3.jpg"
@@
 #### latex math shorthand
-  file: ../../../../MathShorthand.html 
-bibliography: c:/seiro/docs/seiro.bib
-#### ../../../../seiro.bib
+  file: ../MathShorthand.html 
+#### copy c:/seiro/docs/MathShorthand.html ../MathShorthand.html
+bibliography: ../seiro.bib
+#### ../../../../../seiro.bib
+#### copy c:/seiro/docs/seiro.bib ../seiro.bib
 resources:

### 2025-04-03 posts/AboutEditor/AboutEditor.qmd
added Horton2025_slides.html AboutEditor.html

---
title: "エディタ制(仮称)について"
author: "成果検討分科会 伊藤成朗"
date: last-modified
date-format: "MMMM DD, YYYY, HH:MM ZZ"
execute:
  echo: false
engine: knitr
pdf-engine: lualatex
format:
  html:
    toc: true
    toc-location: left
    grid:
      margin-width: 350px
  pdf: default
  docx: default
callout-appearance: simple
callout-icon: false
bibliography: c:/seiro/docs/seiro.bib
link-citations: true
reference-location: margin
citation-location: margin
#### # quarto render AboutEditor.qmd
#### c:\seiro\languages\R\R-4.4.1\bin\x64\Rscript.exe -e "path <- 'c:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/AboutEditor/'; rmarkdown::render(paste0(path, 'AboutEditor.Rmd'), output_format = 'all')"
---
```{css CSS setting, echo=FALSE, results = "hide"}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Noto S	ans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
/* Define a margin before hX (header level X) element */
h1  {
  margin-top: 3ex;
  margin-bottom: 3ex;
  /* background: #c2edff; */ /*背景色*/
  padding: 0.5em;/*文字まわり（上下左右）の余白*/
}
h2  {
  margin-top: 2ex;
  margin-bottom: 2ex;
  padding: 0.5em;/*文字周りの余白*/
  color: #010101;/*文字色*/
  /* background: #eaf3ff; */ /*背景色*/
  /* border-bottom: solid 3px #516ab6; */ /*下線*/
}
```

<style type="text/css">
  body, td{
  font-size: 16pt;
}
</style>

::: {.callout-important title='成果検討過程の役割' collapse='true'}
質の高い研究成果のみを年度内に出版すること
:::

# 役割

1. スケジュール管理
1. 対話の齟齬を減らす

# 目的

1. 年度内出版を実現すること
1. 成果の質や完成度を高めること
   a) **対話齟齬による時間損失を減らし、成果改善の時間を増やす**こと
   a) 校正担当者の負担を減らすこと

# 背景・動機

担当者不在の体制

* 座長が担当者
   * 多岐にわたる内容と多くの分量: 座長の能力を超える
   * 検討過程の生産性? 
   * 座長が替わると監理基準も(時間を通じた公平性)?
* 全員が全部を把握
   * 意識が散漫
   * 大過ないものは手分け
   * 課題ありを全員で考える

# エディタ(仮称)とは

## エディタがすべきこと

1. 検討者指摘と研究会対応の完・未了チェック
1. 研検間対話の円滑化、意見不一致のまとめ、検討者指摘と研究会対応の正当性監視、分科会への報告

::: {.callout-note title='対話を円滑にするためにエディタができること' collapse='true'}
* 明確化: 検討者の指摘事項、著者の修正事項、何をどのように変更すべきか/したか、なぜか
* 履歴の維持: 原則的に追加の指摘を認めない、(指摘対応以外も含め)論旨に関わる内容修正を明記させる
* 検討過程での表現の穏当さも「編集」
   * [事故によって自分の名前で公開されたときにも困らない表現や論調を堅持](https://macartan.github.io/teaching/how-to-critique#pointers)
:::
::: {.callout-note title='正当性監視' collapse='true'}
* 考察対象の尊重: 検討者は「XではなくYを問うべきでは」など、新たな問いを立てることを要求してはいけない、提示された考察に応えるべき
* 採否の論拠
   * 新奇性
   * 証拠や議論の質
   * 社会への貢献
   * 検討者の「学派」(意見・見方)との一致で決めない
:::

3. 事務局への検討完了日変更の相談
4. オプショナル: 検討者2名指摘の簡潔なまとめ(大きな項目提示のみ)

::: {.callout-note title='まとめ例' collapse="true"}
   * 例1
      1. 推計において効果多様性を検討すること 
      1. 母集団を明示し外的妥当性を議論すること
      1. 欠落変数を考慮したplacebo testを検討すること
      1. (以下略)
   * 例2
      1. 総論での枠組み提示が欠けていること 
      1. 一部章(3、5章)での完成度を高めること
      1. (以下略)
   * 「まとめ」ならば検討者に求める方が良いかも?&larr;今回は時すでに遅し^[新年度の成果検討体制は前年度2月中(第1稿提出時)に組む方が望ましい ]
   * 「まとめ」を積み重ねると内部出版物に求められる条件が明らかになっていくかも?
:::

## エディタがすべきではないこと

* 成果検討そのもの、出版可否の判断^[ジャーナル・エディタとの最大の相違点 ]
   * エディタは検討者を代替しない
* 研検間の意見不一致の強引な解消
   * 不一致は成果公開後に議論しても良い
   * 成果を歪めてしまう可能性、分科会に諮る方が安全

### 2025-04-03 posts/AboutEditor/AboutEditor.rmd
added Horton2025_slides.html AboutEditor.html

---
title: "エディタ制(仮称)について"
author: "成果検討分科会 伊藤成朗"
date: "`r format(Sys.time(), '%Y年%m月%d日 %R')`"
header-includes:
- \usepackage{seiro_minimal}
- usepackage{tikz}
- \usepackage{pgfplots}
- \pgfplotsset{width = 7cm, compat = 1.5}
- \pgfmathdeclarefunction{gauss}{2}{\pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}}
output: 
####  bookdown::tufte_html_book:
####    toc: true
  bookdown::html_document2:
    fig_caption: yes
  word_document: default
####  html_document
####    citation_package: natbib
#### tufte::tufte_handout: default
####   tufte_features: ["fonts", "background", "italics"]
bibliography: c:/seiro/docs/seiro.bib
link-citations: yes
#### c:\seiro\languages\R\R-4.4.1\bin\x64\Rscript.exe -e "path <- 'c:/seiro/docs/IDE/kikaku/seika_bunkakai/2024/'; rmarkdown::render(paste0(path, 'AboutEditor.Rmd'))"
#### c:\seiro\languages\R\R-4.4.1\bin\x64\Rscript.exe -e "path <- 'c:/seiro/docs/IDE/kikaku/seika_bunkakai/2024/'; rmarkdown::render(paste0(path, 'AboutEditor.Rmd'), output_format = 'all')"
---
```{css CSS setting, echo=FALSE}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Noto S	ans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
/* Define a margin before hX (header level X) element */
h1  {
  margin-top: 3ex;
  margin-bottom: 3ex;
  /* background: #c2edff; */ /*背景色*/
  padding: 0.5em;/*文字まわり（上下左右）の余白*/
}
h2  {
  margin-top: 2ex;
  margin-bottom: 2ex;
  padding: 0.5em;/*文字周りの余白*/
  color: #010101;/*文字色*/
  /* background: #eaf3ff; */ /*背景色*/
  /* border-bottom: solid 3px #516ab6; */ /*下線*/
}
```

<style type="text/css">
  body, td{
  font-size: 16pt;
}
</style>

# 役割

1. スケジュール管理
1. 対話の齟齬を減らす

# 目的

1. 年度内出版を実現すること
1. 成果の質や完成度を高めること
  1. 対話齟齬による時間損失を減らし、成果改善の時間を増やすこと
  1. 校正担当者の負担を減らすこと

# 背景・動機

 * 担当者不在の体制=座長が担当者&rarr;検討過程の生産性? 座長が替わると監理基準も(公平性)?
 * 全員が全部を把握=意識が散漫&larr;大過ないものは手分け、課題ありを全員で考える

# エディタとは

## エディタがすべきこと

* 検討者指摘と研究会対応の完・未了チェック
* 研検間の意見不一致のまとめ、検討者指摘の正当性監視、分科会への報告
   * とくに採否の論拠
      * 検討者の「学派」(意見・見方)との一致で決めない
      * 新奇性の有無、論理一貫性の有無、証拠や議論の質の高低、社会への貢献度合いなどで判断
* 事務局への検討完了日変更の相談
* オプショナル: 検討者指摘の簡潔なまとめ(大きな項目提示のみ)
   * 例: 「1. 推計における効果多様性 2. 母集団と外的妥当性の議論...」「1. 総論での枠組み提示 2. 一部章での完成度の未達...」など
   * 「まとめ」ならば検討者に求める方が良いかも?&larr;今回は時すでに遅し
   * 「まとめ」を積み重ねると内部出版物に求められる条件が明らかになっていくかも?

## エディタがすべきではないこと

* 成果検討そのもの、出版可否の判断
   * エディタは検討者を代替しない
* 研検間の意見不一致の強引な解消
   * 不一致は成果公開後に議論しても良い

### 2025-04-03 posts/Horton/Horton2024.qmd
added Horton2025_slides.html AboutEditor.html

--- 2024-07-31 posts/BarrEgglestonSmith/BarrEgglestonSmith2022.qmd
@@
 ---
-title: "Investing in infants: the lasting effects of cash transfers to new families"
-subtitle: "By Andrew Barr, Jonathan Eggleston, Alexander A. Smith"
+title: "Price floors and employer preferences: Evidence from a minimum wage experiment"
+subtitle: "By John J. Horton"
 author: "Seiro Ito"
@@
     logo: images/quarto.png
-    footer: <http://seiroito.github.io/QuartoFiles/BarrEgglestonSmith/BarrEgglestonSmith2022.html>
+    footer: <http://seiroito.github.io/QuartoFiles/Horton/Horton2025.html>
     theme: sky
@@
 #width: 1000
-bibliography: seiro.bib
+bibliography: ../seiro.bib
+# If it is 2 levels up, ../../seiro.bib
 resources:
-  - BarrEgglestonSmith2022.pdf
-# quarto render BarrEgglestonSmith2022.qmd
+  - Horton2025.pdf
+# quarto render Horton2025.qmd
 ---
@@
 
-## ECD Long-run effects
+## Firms' minimum wage adjustments
+
+MW&uarr;...,,,
+
+* Employment head counts&darr;
+* Employment hours&darr;
+* Nonmonetary compensation&darr;
+* Profits&darr;
+* Output prices&uarr;
+* Labour productivity &uarr; (labour-labour substitution)
+
+----
+
+## This paper
+
+* Online job platform: Post and bid
+* Random assignment of hourly MW: $2, $3, $4
+* Applicants: Not aware of MW, but bid < MW not accepted
+* Employers: Not aware of MW, see applicant profiles
+* Headcounts&darr; little
+* Hours&darr by 27% for low wage jobs
+* Employment substitution towards more productive workers
+
+----
 
@@
 :::
+
+----
+
+```{r tikz pic, engine = 'tikz'}
+\begin{tikzpicture}
+\draw[ultra thick] (-3, 0) -- (5, 0) node[right]{Birthday};
+\draw[ultra thick] (-3, 0) -- (-3, 5) node[above]{Impacts};
+\draw[blue] (2, 0) node[below]{Jan 1} -- (2, 4.0) node (a){};
+\draw[thick] (a)  -- (-2, 3.5);
+\draw[thick] (2, 1.5) node (b){}  -- (4, 1.75);
+\draw[red, thick, xshift = .1] (a) edge[<->] node[right]{difference} (b);
+\end{tikzpicture}
+```
 

### 2025-04-03 posts/Horton/Horton2024_slides.qmd
added Horton2025_slides.html AboutEditor.html

--- 2025-04-03 posts/Horton/Horton2024.qmd
@@
     logo: images/quarto.png
-    footer: <http://seiroito.github.io/QuartoFiles/Horton/Horton2025.html>
+    footer: <http://seiroito.github.io/QuartoFiles/Horton/Horton2025_slides.html>
     theme: sky
@@
 resources:
-  - Horton2025.pdf
-# quarto render Horton2025.qmd
+  - Horton2025_slides.pdf
+# quarto render Horton2025_slides.qmd
 ---

### 2025-04-03 posts/Horton/Horton2025_slides.qmd
added Horton2025_slides.html AboutEditor.html

---
title: |
  | John J. Horton
  | Price floors and employer preferences: 
  | Evidence from a minimum wage experiment
subtitle: |
  | *American Economic Review*, 2025, vol. 115, No. 1
author: "Seiro Ito"
engine: knitr
format:
  revealjs: 
    standalone: true
    embed-resources: true
    slide-number: true
    width: 1600
    height: 900
    preview-links: auto
    logo: images/quarto.png
    footer: <http://seiroito.github.io/QuartoFiles/Horton/Horton2025_slides.html>
    theme: sky
lightbox: auto
include-in-header:
  - text: |
      <style>
      #title-slide
       .title {
        font-size: 2.0em;
      }
      </style>
bibliography: ../../../../../seiro.bib
#### If it is 2 levels up, ../../seiro.bib
resources:
  - Horton2025_slides.pdf
#### quarto render Horton2025_slides.qmd
---
<style>
:root{--r-main-font-size: 36px;}
.reveal h1, .reveal h2, .reveal h3, .reveal h4, .reveal h5 {
              text-transform: none;
      }
.spacer {
    margin-top: 20px;
}
br {
  /* change <br> space length*/
  display: block; /* makes it have a width */
  content: ""; /* clears default height */
  margin-top: 10px; /* change this to whatever height you want it */
}
</style>

```{css, echo=F, results = "hide"}
DBlue {
  color: dodgerblue;
}
Red {
  color: red;
}
```

## TL; DR

:::{.incremental}
* Im many studies: MW&uarr; &rArr; $\Delta$ employment $\simeq 0$
* One reason: Oligopsony
* This paper: Labor-labor substitution
   * Unskilled employment&darr;
   * Skilled employment&uarr;
   * Net effect $\simeq 0$
   * Hours&darr; &larr; Skilled requires fewer hours
* MW experiment on an online labor market: $2, $3, $4 per hour
   * Skill/productivity proxy = past wage rates, past hours
:::

----

## Why are disemployment impacts of minimum wage modest?

<br>

. . .

Firms can adjust in other dimensions (even under non-oligopsony)

<br>

. . .

MW&uarr;

* Employee head counts&darr;  
* Employment hours&darr;
* Nonmonetary compensation&darr;
* Profits&darr;
* Output prices&uarr;
* Labour productivity &uarr; (labour-labour substitution)

. . .

A problem: Not all adjustments are observable in firm surveys...

----

## Previous literature

Insufficient data

:::{.incremental}
* Do not collect (prices, profits) &rArr; Markup? Prices?
* Not granular (rate, hour, nonmonetary benefits) &rArr; Small employment/hiring impacts 
[@Card1992a; @CardKrueger1995; @KatzKrueger1992; @AllegrettoDubeReich2011; @DubeLesterReich2010; @NSW2014a; @ClemensWither2019; @Powell2022; @MeerWest2015]
   * Workers are considered heterogenous in "types" [@KatzMurphy1992], not in productivity (this paper)
* Unobservable &rArr; Labor-labor substitution is examined only along demographic characteristics [@Giuliano2013; @FairissBujanda2008]
:::

. . .

Even within the "low wage" category/type, more productive workers get jobs while less productive workers lose out

* Net impact $\simeq 0$: Averaging gives muted disemployment impacts
* Previous studies did not show this &larr; Lack of productivity proxy

----

## Experiment design

::: {.incremental}
* Online job platform: Post and bid
* Random assignment of hourly MW: $2, $3, $4
* Applicants: Not aware of MW, but a bid < MW not accepted
* Employers: Not aware of MW, see applicant profiles

1. Firms post hourly job opening &larr; immediate treatment arm assignment
   * Arms: $2 (8%), $3 (8%), $4 (8%), control (75%) assigned to firms, not jobs
   * Job description, skills required, expected duration, bid closing time
1. Workers bid the hourly wage&larr; bid < MW not accepted
1. Workers with the lowest bid receive a contract
:::

. . .

No reference to the name of the platform

----

## Data

* N=156,656 (employers)
* No info on dates, experiment duration, DID uses Jan 2014 - Jan 2015
* No info on employers except job category
* Reshaped data: Category monthly panel, individual ($\pm$ 2 week window) panel 
   * DID: Announcement/imposition impacts on wage, employment, etc.

. . .

I wonder if nondisclosure agreement has to do with...

. . .

No descriptive statistics shown, difficult to understand: all posted jobs $\subset$ all jobs

. . .

What are the differences:

:::{.incremental}
* Sample vs. population of all jobs
* Outsourced jobs vs. internalized jobs
:::

. . .

In DID, are the trends of higher paid jobs appropriate as the control group?

----

## Headline results

::: {.incremental}
   * Headcounts&darr; small
   * Employment probability&darr; by 12% for workers < MW4
   * Hours&darr; by 27% (for low wage jobs, MW4)
   * Employment substitution towards more productive workers
:::

----

## Threats to internal validity ... unlikely

::: {.incremental}
1. Failed randomization
1. Applicant attrition before wage bidding
1. Applicant sorting across arms (MW levels)
   * \# of applicants balance between arms
   * Practically difficult to know the employer's treatment assignment
1. Employer sorting through time (reposts the job later)
   * \# of filled posts: before < after &larr; data show oppositte 
1. Employer sorting to other platform
   * \# of posts: before > after&larr; not supported by data 
:::

. . .

I would worry more about the common trend violation for DID/ES estimates

----

## Where to look for the impacts

What are the "low paying jobs"?

1. Administrative support jobs: Lowest mean wage in all job categories
1. Predictive model (LASSO) of wages using pre-experiment data: Pay less than $5

----

## Mechanism behind the post-bid

:::: {.columns}
::: {.column width="55%"}
![](Fig3.jpg){height="550"}
:::

::: {.column width="45%"}

:::{.fragment}
An example case:

A $\succ$ B $\succ$ C $\succ$ No hire
:::

:::{.fragment}
wage$_{A}$ < wage$_{r}$ < wage$_{B}$ < MW < wage$_{C}$ 
:::

:::{.fragment}
* wage$_{A}$ is not considered
:::
:::{.fragment}
* B'(MW, $y_{B})$ $\prec$ C(wage_{C}, $y_{C})$
:::

:::{.fragment}
Why do applicants bid different $\frac{w}{y}$?

* Different reservation wages
* Not take own characteristics in bids
:::
:::
::::

----

## Estimation

**Experiment** on employer $j$

$$
y_{j}=\beta_{0}+\beta_{2}MW2_{j}+\beta_{3}MW3_{j}+\beta_{4}MW4_{j}+\epsilon_{j}
\tag{1}
$$

:::{.incremental}
* MWs+control are MECE: <Red>@GoldsmithPinkhamHullKolesar2024 bias applies</Red>
* Hereafter, I will assume all the estimates are legit...
:::

. . .

**Confirmatory event study** (platform-wide MW) of treated category, category $i$

$$
y_{it}=\sum_{t=s}^{S}\beta_{t} A\small{DMIN}+\alpha_{i}+\delta_{t}+\gamma_{i}t+\epsilon_{it}
\tag{2}
$$

. . .

**Applicant impact DID** (platform-wide MW), applicant $i$, posted job $j$, wage band $k$

$$
y_{ij}=\sum_{k\in K}\beta_{k}\left(P\small{OST}_{ij}\times \normalsize{P}\small{RE}\normalsize{W}\small{AGE}\normalsize{B}\small{AND}^{k}_{i}\right)+c_{i}+\epsilon_{ij}
\tag{3}
$$

----

## Results

:::::::: {.columns}
::: {.column width="50%"}
![](Fig4.jpg){height="700"}
:::

::: {.column width="50%"}

::::{.fragment}
Employment&darr;

Hours&darr;&darr;

Earnings $\simeq 0$

Sample size is large^[But "All" jobs should not be smaller than "Admin" jobs... ]
::::

<br>

::::{.fragment}
More for

* MW4
* Low predicted wage jobs
::::
:::
::::::::

----

## Results

:::::::: {.columns}
::: {.column width="50%"}
![](Fig5.jpg){height="800"}
:::

::: {.column width="50%"}

::::{.fragment}
Hired workers' past wage&uarr;, cumulative past earnings&uarr;

* Not so much for Admin jobs

Hired workers' profile rate&uarr;

Hired workers' markup rate of bids&uarr;
::::

<br>

::::{.fragment}
More for

* US based workers
* Low predicted wage jobs
   * Not so much for Admin jobs
* Effects on Non-US workers: Because of many less productive workers or country effect?
::::
:::
::::::::

----

## Results

Wage&uarr; sources <DBlue>...All three need to be accounted for</DBlue>

. . .

1. Selection (by firms slashing less productive, low wage jobs)

. . .

* This canot be the only reason
* If so, closure of lowest 10% of all jobs &rArr; 10%&uarr; in wages &larr; control group elasticity = 1
* MW3 Admin: $\Delta$\# of jobs = -2%, mean wage&uarr; $\simeq$ +8% (All jobs is -2%, +25%...), more than selection happens here

. . .

2. Substitution (Increase in worker productivity and pay higher wages)

. . .

* Employ higher paid applicants (Fig 5, past wage +15% for MW4, past earnings +27% for MW4)

. . .

3. Markup (No change in worker productivity but pay higher wages)

. . .

* Markup rate&uarr; (Fig 5, +10% or higher for all MWs)
* A windfall gain for hired workers

----

## Confirmatory DID (Event study)

Online, partial, secret RCT

. . .

<br>

Leaves questions

::: {.incremental}
a) Is L-L substitution **useful margin of adjustments** in equilibrium? <DBlue>yes</DBlue>
a) Does **employer knowledge** on MW matter? <DBlue>no</DBlue>
a) Do **quantity and contents of jobs** change after MW imposition? <DBlue>yes</DBlue>
a) Do the **worker impacts** remain the same in equilibrium? <DBlue>yes</DBlue>
:::

----

## Confirmatory DID

Platform wide imposition of MW $3/hour

* Announcement
* Imposition (2.5 months later)

. . .

1. Single diff (Before-after): Compare market outcomes
1. Double diff (DID): Use <DBlue>the categories that show little difference in wages as the control</DBlue>

. . .

Reconstruct (reshape) data: Category panel data

$$
y_{it}=\sum_{t=s}^{S}\beta_{t} ADMIN+\alpha_{i}+\delta_{t}+\gamma_{i}t+\epsilon_{it}
$$

. . .

Confirms most of experiment results

Jobs posted&darr;, hours&darr;, past earnings &uarr;, past hours&uarr;

. . .

Labor-labor substitution

----

## Confirmatory DID

:::::::: {.columns}
::: {.column width="40%"}
![](Fig7.jpg){height="700"}
:::

::: {.column width="60%"}

::::{.fragment}
* Jobs posted&darr; (mainly hourly jobs&darr;)
::::
::::{.fragment}
* $\Delta$ Fraction filled $\simeq 0$ or &darr; (-.02 vs. -.01 in experiment)
::::
::::{.fragment}
* Mean hours &darr; (-.44 vs. -.08 in experiment)...why larger is not discussed
::::
::::{.fragment}
* $\Delta$ Log mean earnings $\simeq 0$; (same in experiment)
::::
::::{.fragment}
* Log past mean wage of hired worker&uarr; (+.07 vs. +.09 in experiment MW3)^[Author notes "-.02 log points in the experiment" which is for Admin and notes this is of "similar magnitude"'. Perhaps confused with +.09 for all jobs. Or is Fig 7 Panel C for Admin? ]
::::
::::{.fragment}
* Log past mean earnings of hired worker&uarr; (+.2 vs. +.05 in experiment)
::::
::::{.fragment}
* Log past hours of hired worker&uarr;
::::
:::
::::::::

----

## Applicant impact DID

Applicant $i$, posted job $j$, wage band $k$
$$
y_{ij}=\sum_{k\in K}\beta_{k}\left(P\small{OST}_{ij}\times \normalsize{P}\small{RE}\normalsize{W}\small{AGE}\normalsize{B}\small{AND}^{k}_{i}\right)+c_{i}+\epsilon_{ij}
\tag{3}
$$

:::::::: {.columns}
::: {.column width="40%"}
![](Fig8.jpg){height="600"}

Wrong x-axis labels:

(0, 1], (1, 2], (2, 3], ...
:::

::: {.column width="60%"}
* No spillover of MW to higher wages
* Pr[hiring]&darr; for pre-period wage < MW and higher
* \# of hires&darr; for pre-period wage < MW and higher
* \# of bids&darr; for pre-period wage < MW ... they gave up?

::::: {style="font-size: 80%;"}
Reshape (= estimate DID to control typical new applicant wage growth: bid low &rarr; bid high) to

::: {.column width="48%"}
*long* applicant panel

[-14 days, +14 days] ×

{-1 year, imposition year} ×

{all jobs applied} ×

{wage, results}
:::

::: {.column width="48%"}
*short* applicant panel

[-14 days, +14 days] ×

{-1 year, imposition year} ×

{\# of jobs applied, \# of hires}
:::
:::::
:::
::::::::

----

## Conclusions

Experimental findings

1. Wages of hired workers increase
1. Prob of hiring decreases at a sufficiently high MW
1. Hours decrease
1. Reduction in hours can be explained in part by labor-labor substitution

. . .

Observational findings

1. Wages of hired workers increase
1. Employers posted fewer jobs "that would have likely paid lower wages" 
   * They showed reduction in all jobs combined, not only the low paid jobs
1. Hours decrease
1. Firms substitute toward more productive workers
1. Prob of hiring for low wage workers reduced

----

## Conclusions

Labor-labor substitution is an important margin of adjustment for firms

* Less productive gets less jobs, more productive gets more jobs
   * Heterogeneity in disemployment impacts even within a cateory/type
   * Averaging labor-labor substitution gives muted impacts

. . .

> higher minimum wages should be paired with more targeted exemptions for lower-productivity
workers likely to be displaced.

. . .

<br>
<br>
<br>

Workers of different productivity being paid the marginal product in a competitive market?

* Possible, firms can choose the hour-productivity pair that suits them best
* If the labor market is thin, some jobs may not get filled, foreseeing this, firms may not post jobs

----

## 感想

:::{.incremental}
* 最低賃金&uarr; &rArr; 雇用労働の熟練&uarr;, 雇用時間量&darr; を実証的に示したのは画期的
   * 南アの農場主が対応策として同じことを話していた
   * Oligopsonyとの関係? オンライン労働市場M Turkは買い手寡占的[@DubeJacobsNaiduSuri2020]
* 低技能労働が代替されて失職、全労働の純効果はゼロに近いことを示したのも画期的
   * @Cengiz2019: \# of jobs lost below MW $\simeq$ \# of jobs gained above MW
   * Data: Pooled cross section &rArr; Estimates are counts, not individuals moving from below to above MW, laid off workers can be out of labor force
:::

----

## 感想

:::{.incremental}
* 記述統計すらないので結果がどの母集団に一般化できるのか不明
   * Single taskのアウトソース作業 vs. フルタイム雇用: 後者では生産性の差はそこまでないかも
      * 植樹・幡種、剪定、収穫などの単作業の仕事は労働の高熟練化に帰結しやすい?
      * 論文での労働の高熟練化は資本装備率の差も含まれていないか
   * 統御群=高賃金仕事はDID推計で適切か
      * Common trend violationへの頑健性[@RambachanRoth2023]を示してほしい
* 最低賃金規制は変な政策
   * Single taskのアウトソース作業は他の生産工程から分離
   * 労働需要+最低賃金規制=部品需要+最低時間単価規制
   * 最低時間単価を規制する政策=他にはない変わった政策
   * 低効率な部品供給者を閉め出す(=労働時間を過少申告させる)効果あり
* 著者が雇用されていたプラットフォームのデータ、情報非開示合意書NDAでexternal validityを示す手段を失ったのでは?
* プラットフォームが利潤のために実施した実験だからIRB不要、は斬新
:::

----

#### References

::: {#refs}
:::

### 2025-04-03 posts/AboutEditor/AboutEditor.qmd
correceted bib file location in AboutEditor.qmd

@@
 callout-icon: false
-bibliography: c:/seiro/docs/seiro.bib
+bibliography: ../seiro.bib
 link-citations: true

### 2025-04-03 posts/AboutEditor/AboutEditor.qmd
drop pdf from output format in AboutEditor.qmd

@@
       margin-width: 350px
-  pdf: default
+####  pdf: default
   docx: default

### 2025-04-03 posts/Horton/Horton2025_slides.qmd
correceted bib file location in Horton2025_slides.qmd

@@
       </style>
-bibliography: ../../../../../seiro.bib
+bibliography: ../seiro.bib
 #### If it is 2 levels up, ../../seiro.bib

### 2025-04-04 posts/AboutEditor/AboutEditor.qmd
updated deploy to Github Pages in publish.yml

@@
 :::
-::: {.callout-note title='正当性監視' collapse='true'}
+::: {.callout-note title='正当性監視' collapse='false'}
 * 考察対象の尊重: 検討者は「XではなくYを問うべきでは」など、新たな問いを立てることを要求してはいけない、提示された考察に応えるべき
@@
    * 検討者の「学派」(意見・見方)との一致で決めない
+* 相互に矛盾する結果の消去
 :::
@@
 * 成果検討そのもの、出版可否の判断^[ジャーナル・エディタとの最大の相違点 ]
-   * エディタは検討者を代替しない
+   * エディタは検討者を代替しない^[成果検討分科会委員(研究企画委員)としてコメント・質問をすることは可 ]
+   * それとも、第3査読者が必要になったときのみ許容するか=エディタが第3査読者になる^[分科会委員として伊藤が第3査読者になった経験あり(福西座長、企画課長からの依頼) ]
 * 研検間の意見不一致の強引な解消

### 2025-04-04 posts/AboutEditor/AboutEditor.qmd
updated deploy to Github Pages in publish.yml

@@
    * エディタは検討者を代替しない^[成果検討分科会委員(研究企画委員)としてコメント・質問をすることは可 ]
-   * それとも、第3査読者が必要になったときのみ許容するか=エディタが第3査読者になる^[分科会委員として伊藤が第3査読者になった経験あり(福西座長、企画課長からの依頼) ]
+   * それとも、第3査読者が必要になったときのみ許容するか=エディタが第3査読者になる^[分科会委員として伊藤が第3査読者になった経験あり(座長、企画課長からの依頼) ]
 * 研検間の意見不一致の強引な解消

### 2025-04-04 posts/AboutEditor/AboutEditor.qmd
Minor modification in callout block option collapse in AboutEditor.html

@@
   html:
+    standalone: true
+    embed-resources: true
     toc: true
@@
 :::
-::: {.callout-note title='正当性監視' collapse='false'}
+::: {.callout-note title='正当性監視' collapse='true'}
 * 考察対象の尊重: 検討者は「XではなくYを問うべきでは」など、新たな問いを立てることを要求してはいけない、提示された考察に応えるべき

### 2025-04-22 posts/AboutEditor/AboutEditor.qmd
updated versions in actions of publish.yml

--- 2025-04-04 posts/AboutEditor/AboutEditor.qmd
@@
 ```{css CSS setting, echo=FALSE, results = "hide"}
+DBlue {
+  color: dodgerblue;
+}
+Blue {
+  color: blue;
+}
+Red {
+  color: red;
+}
 .SeiroBenign {
@@
 
-1. 検討者指摘と研究会対応の完・未了チェック
-1. 研検間対話の円滑化、意見不一致のまとめ、検討者指摘と研究会対応の正当性監視、分科会への報告
+1. 検討者指摘と研究会対応の<Red>完・未了</Red>チェック
+1. 研検間対話の<Red>円滑化</Red>、意見不一致のまとめ、検討者指摘と研究会対応の<Red>正当性監視</Red>、分科会への報告
 
@@
 * 履歴の維持: 原則的に追加の指摘を認めない、(指摘対応以外も含め)論旨に関わる内容修正を明記させる
-* 検討過程での表現の穏当さも「編集」
-   * [事故によって自分の名前で公開されたときにも困らない表現や論調を堅持](https://macartan.github.io/teaching/how-to-critique#pointers)
+* 検討過程で表現の穏当さの堅持
+   * [事故によって自分の名前で公開されたときにも困らない表現や論調](https://macartan.github.io/teaching/how-to-critique#pointers)
 :::
-::: {.callout-note title='正当性監視' collapse='false'}
+::: {.callout-note title='正当性監視' collapse='true'}
 * 考察対象の尊重: 検討者は「XではなくYを問うべきでは」など、新たな問いを立てることを要求してはいけない、提示された考察に応えるべき
@@
    * 検討者の「学派」(意見・見方)との一致で決めない
-* 相互に矛盾する結果の消去
+* (主に著者による)相互に矛盾する結果の消去
 :::
 
-3. 事務局への検討完了日変更の相談
+3. \sout{事務局への}検討完了日変更\sout{の相談}への関与
 4. オプショナル: 検討者2名指摘の簡潔なまとめ(大きな項目提示のみ)

### 2025-04-22 posts/AboutEditor/AboutEditorQA.qmd
updated versions in actions of publish.yml

---
title: "エディタ制についてQ&A1"
author: "成果検討分科会 伊藤成朗"
date: last-modified
date-format: "MMMM DD, YYYY, HH:MM ZZ"
execute:
  echo: false
engine: knitr
pdf-engine: lualatex
format:
  html:
    standalone: true
    embed-resources: true
    toc: true
    toc-location: left
    grid:
      margin-width: 350px
####  pdf: default
  docx: default
  epub: default
theme: 
  light: sky
  dark: darkly
title-block-banner: "#f0f3f5"
title-block-banner-color: "black"
callout-appearance: simple
callout-icon: false
bibliography: ../seiro.bib
link-citations: true
reference-location: margin
citation-location: margin
#### # quarto render AboutEditorQA.qmd
#### c:\seiro\languages\R\R-4.4.1\bin\x64\Rscript.exe -e "path <- 'c:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/AboutEditor/'; rmarkdown::render(paste0(path, 'AboutEditor.Rmd'), output_format = 'all')"
---
```{css CSS setting, echo=FALSE, results = "hide"}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Noto S	ans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
/* Define a margin before hX (header level X) element */
h1  {
  margin-top: 3ex;
  margin-bottom: 3ex;
  /* background: #c2edff; */ /*背景色*/
  padding: 0.5em;/*文字まわり（上下左右）の余白*/
}
h2  {
  margin-top: 2ex;
  margin-bottom: 2ex;
  padding: 0.5em;/*文字周りの余白*/
  color: #010101;/*文字色*/
  /* background: #eaf3ff; */ /*背景色*/
  /* border-bottom: solid 3px #516ab6; */ /*下線*/
}
```

<style type="text/css">
  body, td{
  font-size: 16pt;
}
</style>

## 箭内さんからの質問と暫定的回答

1. 「相互に矛盾する結果の消去」の意味は？
   * これは原稿についての記述で、検討票についての記述ではありません。主に推計結果について想定しています。推計によっては、相互に矛盾する結果が出ることがあり、安易な著者だと「都合の良い結果」だけを残すことがあります。(定性的な分析でもこうした事例はあるのかもしれないですが)
   * 検討者間に矛盾がある場合(「終章を付けるべき」「終章は不要」)は、出てきたら考えましょう。
1. 「検討完了日変更の相談」については、引き続き事務局で対応していただけないか^[分科会委員が研究会側の場合、実質的にブラインドできないだろうというのが前回の認識でした。検討者名は分科会委員にブラインドされています。よって、分科会委員は研究会側としか直接連絡できません。 ]
   * 長期の延長や幾度の延長でない限り、事務局+座長による了承&rarr;エディタおよび分科会に報告、了承を得る
1. 「指摘内容のまとめ」を審査者に求めるのは負荷が多すぎないか？
   * 検討票の講評部分に下記を加える

   > できる限り、個別指摘点の要約ともなるように記してください。

1. 特定の指摘事項を「まとめ」として抽出すると、研究会の対応がそれだけにとどまったり、指摘内容の意図を正確に伝えきれなかったりする虞れが生じる。
   * その恐れが確かにあります。よって、まとめは分科会限りにします。用途は主な指摘事項をクリアしているかどうかの確認です。
1. 既に提出された原稿の仕上がり具合が悪い場合、どうしたらよいか？
   * 別のエディタからも同じ懸念が表明されました。
1. フォーム
   * フォームのお試し版を事務局で用意します。次項参照。

## エディタ報告フォーム

第2回分科会で必要性について提起されました。記録のためにもエディタ報告には報告フォームを用意するようお願いいたします。

フォームは特に準備いたしませんが、以下の項目を含めてください。

1. 研究会概要
   a) 課題ID
   a) 最終成果の発表形態・備考(一般書、学術書)^[課題DBの成果提出表示ページに記載 ]
1. 成果概要
   a) 章題-執筆者名の表^[誰がどの章を執筆したかを知ることが目的です。 内部/外部、専門分野などの知識は、検討過程効率化の材料にも成りえます。 ] ^[成果検討分科会に対してブラインドすることも考えられますが、手間に見合う便益がないと思い、執筆者名は明示したいと思います。研究会主査に表提出を求めます。 ]
   a) 単行書のトピック
1. 指摘概要^[検討者別にしても、前検討者をまとめても、どちらでも良いと思います。 ]
   a) 内容
   a) 表現、構成、体裁
1. エディタ判断
   a) 検討票・原稿への修正要請有無
      * 指摘内容の正当性
      * 指摘内容の一意性(対象・疑問・意見が具体的・特定的であること)
      * 表現の穏当さ
   a) 検討者・研究会への参考意見有無
      * 分科会からの参考意見付与には下記の条件を要する
        1. 分科会委員全員の同意を得ること
        1. 任意対応の意見とすること
        1. 発議するエディタが論拠を示すこと

## 原稿の仕上がり具合が悪い場合

### 検討者はどう対応できるか

* 今まで: 原稿ファイルに直接書き込み
* 段落単位、もしくは、節単位で内容を明確化するように尋ねる

### 検討制度を変更

* 検討回数: 2回&rarr;1回^[検討者とエディタ(成果検討分科会委員)の業務負担を減らせる ] ^[ジャーナルではたいてい1回 ]

## その他

第2回成果分科会後に、事務局内で以下を話し合いました

### 検討回数

検討回数が2回になっていることの長短

* 長所
   * 複数回読んでもらうことで着実に内容が改善する
   * 却下となる研究会成果が減るので発行点数(単行本の本数)を減らさずに済む
* 短所
   * 時間と労力を要する
   * 初稿の完成度が低いなどの甘えを誘発する

その他の意見

* 初稿提出期限を延期すれば1回で完成度を高められるか? 2月中旬&rarr;3月中旬
* 研究企画委員会だけで決められる問題ではない

### 検討日程のリスケジューリング

* リスケジューリングを断っても立ち行かない場合がほとんど&larr;了承せざるを得ない
* 長期の延長や幾度の延長でない限り、事務局+座長による了承&rarr;エディタおよび分科会に報告、了承を得る

### 2025-04-22 posts/AboutEditor/Ito2024Apr.qmd
updated versions in actions of publish.yml

---
title: |
  | 日本・バングラデシュ関係50年: 現状と課題
  | Japan-Bangladesh relationship at 50: Achievements and challenges
subtitle: 研究課題ID319
author: "成果検討分科会 伊藤成朗"
date: last-modified
date-format: "MMMM DD, YYYY, HH:MM ZZ"
execute:
  echo: false
engine: knitr
pdf-engine: lualatex
format:
  html:
    standalone: true
    embed-resources: true
    toc: true
    toc-location: left
    grid:
      margin-width: 350px
####  pdf: default
  docx: default
theme: 
  light: sky
  dark: darkly
title-block-banner: "#f0f3f5"
title-block-banner-color: "black"
callout-appearance: simple
callout-icon: false
bibliography: ../seiro.bib
link-citations: true
reference-location: margin
citation-location: margin
#### # quarto render Ito2024Apr.qmd
#### c:\seiro\languages\R\R-4.4.1\bin\x64\Rscript.exe -e "path <- 'c:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/Ito2024Apr/'; rmarkdown::render(paste0(path, 'Ito2024Apr.Rmd'), output_format = 'all')"
---
```{css CSS setting, echo=FALSE, results = "hide"}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Noto S	ans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
/* Define a margin before hX (header level X) element */
h1  {
  margin-top: 3ex;
  margin-bottom: 3ex;
  /* background: #c2edff; */ /*背景色*/
  padding: 0.5em;/*文字まわり（上下左右）の余白*/
}
h2  {
  margin-top: 2ex;
  margin-bottom: 2ex;
  padding: 0.5em;/*文字周りの余白*/
  color: #010101;/*文字色*/
  /* background: #eaf3ff; */ /*背景色*/
  /* border-bottom: solid 3px #516ab6; */ /*下線*/
}
li > ul {
    padding-left:25px;
}
li > ol {
    padding-left:25px;
}
```

<style type="text/css">
  body, td{
  font-size: 16pt;
}
</style>

# 成果概要

1. 日バングラ関係史^[こういう経緯でした、というカタログ的内容。なので、Research questionはなくても良いのでは? ]: 序.要約、独立(1.日, 2.バ)、3.経済関係、4.ODA、5.NGO、人材育成(6.日, 7.バ)、8.新たな国際情勢、9.和文バングラデシュ研究目録
1. (おそらく)非研究者を含む執筆陣

# 検討者指摘概要

## 各項目の評価

:::{style="line-height: 92%;caption-side: top;"}
```{r result table, warning = F}
library(tinytable)
library(data.table)
ktb <- fread("kentohyo_tableformat.prn", header = F, sep = "\t")
ktb0 <- unlist(ktb[grepl("内容|適切", V1), ])
ktb1 <- unlist(ktb[grepl("総論|先行|論旨|知見|長さ|表現", V1), ])
ktb2 <- ktb[!grepl("内容|適切|総論|先行|論旨|知見|長さ|表現", V1), ]
ktb2[, c("J1", "J2", "E1", "E2") := ""]
setnames(ktb2, "V1", "項目")
ktb2[c(2, 8, 13, 17, 22, 27, 34), J1 := "○"]
ktb2[c(2, 7, 12, 16, 22, 28, 33), J2 := "○"]
ktb2[c(seq(1, 25, 5), 26, 32), E1 := "○"]
ktb2[c(seq(3, 20, 5), 22, 29, 33), E2 := "○"]

kt <- tt(ktb2, caption = "各項目の評価")
rowlist <- list()
#### Need to split row group label 1 to two lines as it is too long
rowlist[[1]] <- 1
rowlist[[2]] <- 1
rg1 <- c("「総論」（として位置づけられる章）において" ,
  "問題意識または仮説が明確に提示されているか。")
for (i in 2:length(ktb1)) rowlist[[i+1]] <- 1+5*(i-1)
names(rowlist)[1:2] <- rg1
names(rowlist)[-(1:2)] <- ktb1[-1]
kt <- theme_tt(kt, theme = "void")
kt <- group_tt(kt, i = rowlist)
kt <- style_tt(kt, 
  i = kt@group_index_i,
  align = "l",
  color = "black",
  background = "e4ebfa",
  bold = F,
  line = NULL
)

ktb3 = copy(ktb2)
ktb3[, question := ""]
rg <- c(rg1, ktb1[-1])
ktb3[1, question := rg[1]]
ktb3[2, question := rg[2]]
for (i in 3:length(ktb1)) ktb3[1+5*(i-2), question := rg[i]]
for (i in 1:length(ktb1)) ktb3[1+5*(i-1), question := ktb1[i]]
setcolorder(ktb3, c("question", "項目", "J1", "J2", "E1", "E2"))
setnames(ktb3, c("項目", "評価段階", "日1", "日2", "英1", "英2"))
kt <- tt(ktb3, width = c(.4, .4, rep(.1, 4)),
  caption = "各項目の評価")
kt <- theme_tt(kt, theme = "void")
for (i in 1:length(ktb1))
  kt <- style_tt(kt, 
    i = 1 + (i-1)*5, j = 1,
    color = "black",
    background = "e4ebfa",
    rowspan = 3
  )
kt <- style_tt(kt, bootstrap_css_rule = "table{caption-side: top;}")
kt
```
:::

## 講評、個別指摘事項

1. 類書はない、存在価値あり^[全検討者]
1. Research questionへの疑問
   a. 前提とする認識は正しいか?^[英2、日1 ]

::: {.callout-note title='good relationship?' collapse='true'}
The book also assumes that the relationship between the two countries has always been "good." While this is likely true, interestingly, reports (Chapter 8) suggest that Japan's ranking as a country with a favorable impression is not as high as that of countries with one or more periods of conflict (such as the USA and China) and is instead similar to that of India. 
:::

   b. 扱っているのはWhyではなくHowでは?^[英2 ]

::: {.callout-note title='research question, intro (p.1)' collapse='true'}
> 50 years of diplomatic relations...Japan and Bangladesh have always maintained good relations...it is worth considering **why this has been possible,** given the fact that Japan is the only major diplomatic partner with which Bangladesh has consistently maintained good relations...
:::

   c. 本書の回答・解釈は何か^[英2 ]

::: {.callout-note title='overall answer?' collapse='true'}
序章

please clearly discuss the answers to the overall findings. 
:::

3. 内容として全体にまとまりがない^[日2 ]
   * ただし、全体として何かを示すという方向性の指摘はない
   * 各章を改稿すべきという指摘のみ
   * 削除も視野に検討: 9章
1. 一部の章では分析対象の延伸が望ましい^[英2、日2 ]
   a) 最近10年ではなく、可能な限り遡るべき: 3章
   a) 経済関係は他国も含めるべき: 3章
1. 章間で統一性がない
   a) 各章: 質・量ともに不均一^[英2、日2 ]
      * 2倍量: 2、4章&larr; 他章との差を縮めるべき
      * 明確に目的/結論がない: 序、3、4、5、7、8章
      * 問題意識や導入がない章がある: 「序章および各章」^[日1 ]
      * 第4章: 図表合計=37
   a) 体裁(参考文献の記載・引用方法、固有名詞)の不統一^[英2、日1、日2 ]
1. それ以前の推敲不足^[英2、日1、日2 ]
1. 引用が明示的ではない
   a) 無断引用による著作権侵害の疑い^[日1「図表を個人で作成するようにお願いしたい」、英2 ]: 5、8章
   a) 原典が機関名のみ^[日1、日2 ]: 3、4章

# 所感

## 判断

このまま研究会に送付

* 検討票の誤字脱字は検討者に修正依頼、修正完了

1. 検討者指摘の評価
   * 内容は正当
   * 対象・疑問は具体的
   * 表現は適切
1. 章を加えるべきという指摘はない
1. 先行文献との違いに関する指摘はない
   * 読者は先行文献を知りたいはず&rarr; 「序章で言及があると良い」と分科会から伝えるか?

## 感想

* 意図は良いが実行が悪い研究成果
* 専門書として表現方法、内容ともに完成度の低い成果
   * これを検討者に検討させるのか
   * これを日英両言語で出版するのか
   * カタログ的内容なのに専門書としての和文単行書で良いのか 
* 同一内容を4人に検討させるのは無駄
* 検討者の努力も濃淡がある
   * 特濃: 英検討者2(大幅加筆修正)
   * 濃: 日検討者2^[指摘の内容や対象が特定されていないこともあるが、すべて「B.任意」対応指定なので問題なし ](大幅加筆修正)
   * 中: 日検討者1(大幅加筆修正)
   * 特淡: 英検討者1(了承)
      * どこまで真剣に検討したか疑問

### 2025-06-18 posts/DelabastitaRubens/DelabastitaRubens.qmd
added DelabastitaRubens repo and ethics repo

---
title: |
  Colluding against workers\
  \
  *Journal of Political Economy* \
  2025 (Volume 133), June (No.6)
author: "Vincent Delabastita and Michael Rubens, heavy lifting of drafting the memo is done by claude 4 sonnet, overseen by me"
date: last-modified
date-format: "MMMM DD, YYYY, HH:MM ZZ"
execute:
  echo: false
engine: knitr
pdf-engine: lualatex
format:
  html:
    standalone: true
    embed-resources: true
    toc: true
    toc-location: left
    grid:
      margin-width: 350px
    linkcolor: "green"
    urlcolor: "#009900"
####  pdf: default
  docx: default
  epub: default
lightbox: auto
theme: sky
  ####light: [sky, ./sky-add]
  ####dark: darkly
backgroundcolor: "#fffcf2"
title-block-banner: "#f0f3f5"
title-block-banner-color: "black"
doi-title: "doi.org/10.1086/734780"
callout-appearance: simple
callout-icon: false
bibliography: ../seiro.bib
link-citations: true
reference-location: margin
citation-location: margin
#### # quarto render DelabastitaRubens.qmd
#### c:\seiro\languages\R\R-4.4.1\bin\x64\Rscript.exe -e "path <- 'c:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/DelabastitaRubens/'; rmarkdown::render(paste0(path, 'DelabastitaRubens.Rmd'), output_format = 'all')"
---
```{css CSS setting, echo=FALSE, results = "hide"}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightWreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Noto S	ans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
/* Define a margin before hX (header level X) element */
h1  {
  margin-top: 3ex;
  margin-bottom: 3ex;
  /* background: #c2edff; */ /*背景色*/
  padding: 0.5em;/*文字まわり（上下左右）の余白*/
}
h2  {
  margin-top: 2ex;
  margin-bottom: 2ex;
  padding: 0.5em;/*文字周りの余白*/
  ####color: #010101;/*文字色*/
  color: #87a5f2;
  /* background: #eaf3ff; */ /*背景色*/
  /* border-bottom: solid 3px #516ab6; */ /*下線*/
}
DBlue {
  color: dodgerblue;
}
Blue {
  color: blue;
}
Red {
  color: red;
}
```

<style type="text/css">
  body, td{
  font-size: 16pt;
}
</style>

::: {.flushright}

[doi.org/10.1086/734780](https://doi.org/10.1086/734780)

:::

## Summary

::: {.column-margin}
![Wage markdown](./Fig3.jpg)

![coal prices](./Fig2.jpg)
:::

* Workers are paid `r round((1/1.5)*100, 0)`% - `r round((1/2.2)*100, 0)`% of MRPL at the turn of 20th century Belgium coal mines
* Descriptive stat: Cartel &rArr; $P$&uarr;  &rArr; wage markdown $\frac{MRPL}{W}$&uarr;
* Regime change: Employer Association&rarr;coal price cartel
* Observe: Wage markdown estimates vs. "collusion" regime change 
* Claim: Collusion&uarr; &rArr; markdown&uarr;
* Literature: Friction, factor market power &rArr; wage markdown
* New: Downstream/employer collusion &rArr; upstream wage markdown

## Why the wage markdown matters

* Intrinsic interest: Is labour exploited?
* Analytical interest: 
   * How big is the welfare (deadweight) loss? 
   * Markdown&uarr;,  markups&uarr; &rArr; TFP&darr;, inflation&uarr;^[How much is TFP reduced/inflation hightened by the resource misallocation in the economy due to wage markdown, not just price markups? ]
* Need to look at both ends: An exploited (from steel plants, railways) needs to be an exploiter (coal mine)
* Measurement issue: What does the markup formula give under the upstream and downstream market power?

An extreme case: Monopsonistic-monopolistic profit maximisation [@Syverson2024]
$$
\max_{\{X\}} \;\;\; \pi = R\{f(X)\} - W(X)X
$$

FOC^[$\varepsilon^{-1}_{x}=\frac{dW}{dX}\frac{X}{W}$ is factor supply elasticity of price $\varepsilon_{x}\equiv\frac{dX}{dW}\frac{W}{X}$ inverted, or inverse price elasticity of factor suply. ]
$$
\begin{aligned}
R'(Q)f'(X) 
&= 
W'(X)X + W(X)\\
&=
W(X)\bigg(1+\underbrace{W'(X)\frac{X}{W(X)}}_{=\epsilon^{-1}_{x}}\bigg)\\
&=
W(X)\left(1+\varepsilon^{-1}_{x}\right)
\end{aligned}
$$
Markdown
$$
\mu^{x}\equiv \frac{MRPX}{W}=\frac{R'(Q)f'(X)}{W(X)}=1+\varepsilon^{-1}_{x}
$$
So^[ 1st equality uses
$$
\theta_{x}\equiv f'(X)\frac{X}{Q}
$$
or
$$
f'(X)=\theta_{x}\frac{Q}{X}
$$ ] ^[ 2nd equality uses FOC of factor $X$ from monopolist profit maximisation $PF(X)-C\{F(X)\}$, denoting $R'(X)=R'(Q)F'(X)$:
$$
\begin{aligned}
P'(Q)F'(X)Q+P(Q)F'(X)
&=
C'(Q)F'(X)\\
&\Leftrightarrow\\
\{\underbrace{P'(Q)Q+P(Q)}_{=R'(Q)}\}F'(X)
&=
C'(Q)F'(X)\\
&\Leftrightarrow\\
R'(Q)=P(Q)\left(1+P'(Q)\frac{Q}{P}\right)
&=
MC\\
&\Leftrightarrow\\
P(1+\eta^{-1})
&=
MC
\end{aligned}
$$
where $\eta^{-1}=\frac{dP}{dQ}\frac{Q}{P}$ is demand elasticity of price $\eta\equiv\frac{dQ}{dP}\frac{P}{Q}$ inverted, or inverse price elasticity of demand, so they say.
$$
\mu^{-1}=\frac{MC}{P}=1+\eta^{-1}=1-\left|\eta^{-1}\right|
$$
 ]
$$
\begin{aligned}
\mu^{x}
&=
\frac{R'(Q)}{W(X)}\frac{\theta_{x}Q}{X}\\
&=
\frac{P\left(1+\eta^{-1}\right)}{W(X)}\frac{\theta_{x}Q}{X}\\
&=
\left(1-\left|\eta^{-1}\right|\right)\frac{\theta_{x}}{\alpha^{x}}
\end{aligned}
$$
Note $\mu^{-1}=1-\left|\eta^{-1}\right|$, so
$$
\mu^{x}\mu=\frac{\theta_{x}}{\alpha^{x}}
$$
The popular markup formula $\frac{\theta_{x}}{\alpha^{x}}$ gives a product of markdown and markup.

$\frac{\theta_{x}}{\alpha^{x}}$ gives

* Markup, only when the firm is non-oligopsonistic
* Markdown, only when the firm is non-oligopolistic^[Same problem applies to this paper. It cancels off markup by $\frac{\theta_{\ell}}{\alpha^{\ell}}/\frac{\theta_{m}}{\alpha^{m}}=\mu^{\ell}\mu/\mu^{m}\mu=\mu^{\ell}$ under the assumption of material market competition $\mu^{m}=1$. See \eqref{eq14} in @sec-identification ]
* Overestimation of markup/markdown when markup>1 and markdown>1, if attributing RHS to only one of the two&rarr;...@DeLoeckerEeckhoutUnger2020?

## Model

### Primitives

:::: {.callout-important title='Variable definitions' collapse='true'}

| Variable | Definition |
|----------|------------|
| Markdown | $\mu^{\ell}_{ft} \equiv \frac{MRPL_{ft}}{W^{\ell}_{ft}}$ |
| Percentage markdown | $d^{\ell}_{ft} \equiv \frac{MRPL_{ft} - W^{\ell}_{ft}}{MRPL_{ft}} = \frac{\mu^{\ell}_{ft} - 1}{\mu^{\ell}_{ft}}$ |
| Markup | $\mu_{ft} \equiv \frac{P_{ft}}{MC_{ft}}$ |
| Labor market share | $s^{\ell}_{ft} \equiv \frac{L_{ft}}{L_{it}}$ |
| Revenue share of labor | $\alpha^{\ell}_{ft} \equiv \frac{W^{\ell}_{ft} L_{ft}}{P_{ft} Q_{ft}}$ |
| Revenue share of materials | $\alpha^{m}_{ft} \equiv \frac{W^{m}_{ft} M_{ft}}{P_{ft} Q_{ft}}$ |
| Inverse labor supply elasticity (firm-level) | $\theta^{l}_{ft} \equiv \frac{\partial W^{\ell}_{ft}}{\partial L_{ft}} \frac{L_{ft}}{W^{\ell}_{ft}}$ |
| Inverse materials supply elasticity (firm-level) | $\theta^{m}_{ft} \equiv \frac{\partial W^{m}_{ft}}{\partial M_{ft}} \frac{M_{ft}}{W^{m}_{ft}}$ |
::::

Cost minimization 
\begin{align}
\min_{\left\{L_{ft}, M_{ft}\right\}} \;\;\;
& \sum_{g \in F_{i(f)t}} \lambda_{fgt} \left( L_{gt} W^{\ell}_{gt} + M_{gt} W^{m}_{gt} \right)\\
&\hspace{1em} - MC_{ft} \left\{ Q(L_{ft}, M_{ft}, K_{ft}, \Omega_{ft}; \beta) - Q_{ft} \right\}  \tag{4}
\end{align}

Cobb-Douglas production function^[Results also follow with output elasticity + Hicks neutrality
$$
Q_{ft} = \Omega_{ft}F(L_{ft}, M_{ft}, K_{ft}).
$$
Why CD? ]
$$
q_{ft} = \beta_{l} l_{ft} + \beta_{m} m_{ft} + \beta_{k} k_{ft} + \omega_{ft} \tag{1}
$$

TFP AR(1): Blundell and Bond way^[Justification given in the paper: Input demand is affected by distributions of markdowns and output markups, otherwise one cannot invert. To avoid assumptions on the markdown and markup distributions, we use AR(1). ]
$$
\omega_{ft} = \rho \omega_{ft-1} + u_{ft} \tag{2}
$$

Labor supply^[Estimated using demand shocks as IVs: 1987 coal demand shock, cartel membership ]
$$
W^{\ell}_{it} = L_{it}^{\Psi^{l}} \nu_{it} \tag{3}
$$

FOC of labour
$$
\begin{alignat}{3}
MC_{ft}\frac{\partial Q_{ft}}{\partial L_{ft}} 
&= 
W^{\ell}_{ft} 
&&+ \frac{\partial W^{\ell}_{ft}}{\partial L_{ft}} L_{ft} 
&&+\sum_{g\neq f}\lambda_{fgt} \frac{\partial W^{\ell}_{gt}}{\partial L_{ft}}L_{gt}\\
&\Leftrightarrow &&\\
\frac{P_{ft}}{\mu_{ft}}\frac{\partial Q_{ft}}{\partial L_{ft}} 
&= 
W^{\ell}_{ft}
&&+\frac{W^{\ell}_{ft}}{1}\underbrace{\frac{\partial W^{\ell}_{ft}}{\partial L_{ft}}\frac{L_{ft}}{W^{\ell}_{ft}}}_{=\psi^{\ell}_{ft}}
&&+ W^{\ell}_{ft}\sum_{g\neq f}\lambda_{fgt} \underbrace{\frac{\partial W^{\ell}_{gt}}{\partial L_{ft}}\frac{L_{ft}}{W^{\ell}_{gt}}}_{=\psi^{\ell}_{fgt}}\frac{W^{\ell}_{gt}}{L_{ft}}\frac{L_{gt}}{W^{\ell}_{ft}}
\end{alignat}
$$

Imposing profit maximisation: marginal revenue = marginal cost:

$$
\frac{\partial R_{ft}}{\partial Q_{ft}}
=
\frac{\partial C_{ft}}{\partial Q_{ft}}
=
\frac{P_{ft}}{\frac{P_{ft}}{\frac{\partial C_{ft}}{\partial Q_{ft}}}}
=
\frac{P_{ft}}{\mu_{ft}}
$$

Then, MRPL is
$$
MRPL_{ft}
\equiv 
\frac{\partial R_{ft}}{\partial Q_{ft}}\frac{\partial Q_{ft}}{\partial L_{ft}}
=
\frac{P_{ft}}{\mu_{ft}}\frac{\partial Q_{ft}}{\partial L_{ft}}
$$

So^[Can derive a similar expression under collusion $\mu^{x}\mu=\frac{\theta_{x}}{\alpha^{x}}$ as in Syverson (2022) by redefining elasticities $\varepsilon_{\ell}, \eta$. ]

\begin{aligned}
MRPL_{ft}
&= 
W^{\ell}_{ft}\bigg(1+\psi^{\ell}_{ft}+\sum_{g\neq f}\lambda_{fgt}\psi^{\ell}_{fgt}\frac{W^{\ell}_{gt}L_{gt}}{W^{\ell}_{ft}L_{ft}}\bigg)\\
&\Leftrightarrow \\
\mu^{\ell}_{ft}
&= 
1+\psi^{\ell}_{ft}+\sum_{g\neq f}\lambda_{fgt}\psi^{\ell}_{fgt}\frac{W^{\ell}_{gt}L_{gt}}{W^{\ell}_{ft}L_{ft}}
\end{aligned}

Markup from variable material inputs^[Competitive cost minimization $W^{m}M-\lambda[Q(M)\geqslant \bar{Q}]$ FOC gives
$$ 
W^{m}= \lambda Q'(M)=MCQ'(M)
$$
\begin{align}
\frac{W^{m}}{MC}\frac{PM}{PQ}
&=
Q'(M)\frac{M}{Q}\\
\frac{W^{m}M}{PQ}\frac{P}{MC}
&=
\theta_{m}\\
\alpha^{m}\mu
&=
\theta_{m}\\
\mu
&=
\frac{\theta_{m}}{\alpha^{m}}
\end{align}
 ]
$$
\mu_{ft} = \frac{\beta_{m}}{\alpha^{m}_{ft}} \tag{13}
$$

### No collusion $\lambda_{fgt}=0$

\begin{align}
\min_{\left\{L_{ft}, M_{ft}\right\}} \;\;\;
& L_{ft} W^{\ell}_{it} + M_{ft} W^{m}_{it}\\
&\hspace{1em} - MC_{ft} \left\{ Q(L_{ft}, M_{ft}, K_{ft}, \Omega_{ft}; \beta) - Q_{ft} \right\}  \tag{5}
\end{align}

FOC
\begin{align}
W^{\ell}_{it} + \frac{\partial W^{\ell}_{it}}{\partial L_{it}} L_{ft} 
&= 
\frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} \tag{6}\\
\end{align}
Rewrite $L_{ft}=s^{\ell}_{ft}L_{it}$
\begin{align}
MRPL_{ft}
&= 
s^{\ell}_{ft} \frac{\partial W^{\ell}_{it}}{\partial L_{it}}\frac{L_{it}}{W^{\ell}_{it}}W^{\ell}_{it} + W^{\ell}_{it} \\
&=
W^{\ell}_{it}\left(1+s^{\ell}_{ft}\psi^{\ell}_{ft}\right)
\end{align}

::: {.column-margin}
![stata code](./dofile.jpg)
:::

Markdown
<!--
^[Where did $\mu_{ft}$ go? ] ^[Looking at [2. Markups and markdowns of stata code](https://dataverse.harvard.edu/file.xhtml?fileId=10480221&version=1.0):
In the loop, 1st line is markup, 2nd line is markdown computed by dividing with markup: $\frac{1}{\mu}\frac{\beta_{\ell}}{\alpha^{\ell}}$
 ] ^[Yeh et al. (2024) used markup from materials to compute markdown
$$
\mu^{\ell}=\mu\left(1+s^{\ell}_{ft}\psi^{\ell}_{ft}\right)=
$$
 ]
 -->
 ^[From data, it is shown that $\mu_{ft}$ is empirically 1... ] 
$$
\underline{\mu}^{\ell}_{ft} = 1 + s^{\ell}_{ft} \Psi^{\ell} \tag{7}
$$

* Markdown is greater for larger firms

### Perfect collusion $\lambda_{fgt}=1$

\begin{align}
\min_{\left\{L_{ft}, M_{ft}\right\}} \;\;\;
& \sum_{g \in F_{i(f)t}} \left( L_{gt} W^{\ell}_{gt} + M_{gt} W^{m}_{gt} \right)\\
&\hspace{1em} - MC_{ft} \left\{ Q(L_{ft}, M_{ft}, K_{ft}, \Omega_{ft}; \beta) - Q_{ft} \right\}  \tag{8}
\end{align}

FOC
$$
W^{\ell}_{it} + \frac{\partial W^{\ell}_{it}}{\partial L_{it}}L_{it}  = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} \tag{9}
$$

Markdown
$$
\overline{\mu}^{\ell}_{ft} = 1 + \Psi^{\ell} \tag{10}
$$

* Markdown is uniform across firms

### Intermediate case $\lambda_{fgt}\in (0,1)$

Introduce conduct parameter $\widetilde{\lambda}_{ft}$ by rewriting FOC to nest no collusion $\widetilde{\lambda}_{ft}=s_{ft}$ and collusion $\widetilde{\lambda}_{ft}=1$ (so $\widetilde{\lambda}_{ft}\in[s_{ft}, 1]$):
\begin{equation}
W^{\ell}_{it} + \widetilde{\lambda}_{ft} \frac{\partial W^{\ell}_{it}}{\partial L_{it}} L_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} \tag{11}
\label{eq11}
\end{equation}

Markdown 
$$
\mu^{\ell}_{ft} = 1 + \widetilde{\lambda}_{ft} \Psi^{\ell} \tag{12}
$$

* Markdown is heterogenous across firms

Redefine conduct parameter $\widehat{\lambda}_{ft}\in[0, 1]$^[
$$
\widehat{\lambda}_{ft} =
\left\{
\begin{array}{cc}
0\\
1
\end{array}
\right. \quad \mbox{if} \quad
\mu^{\ell}_{ft}=
\left\{
\begin{array}{c}
\underline{\mu}^{\ell}_{ft}\\
\bar{\mu}^{\ell}_{ft}
\end{array}
\right.
$$ ]
$$
\widehat{\lambda}_{ft} = \frac{\mu^{\ell}_{ft} - \underline{\mu}^{\ell}_{ft}}{\overline{\mu}^{\ell}_{ft} - \underline{\mu}^{\ell}_{ft}} \tag{15}
$$

:::: {.callout-important title='Summary of theoretical predictions' collapse='true'}

::: {style="font-size: 100%;"}

$$
\mbox{When market conduct is} 
\left\{
\begin{array}{l}
\mbox{Cournot}\\
\mbox{intermediate}\\
\mbox{collusive}
\end{array}
\right.
\quad \mu^{\ell} \ \mbox{is} 
\left\{
\begin{array}{l}
\mbox{size proportional }\underline{\mu}^{\ell}_{ft}\\
\mbox{heterogenous } \mu^{\ell}_{ft}\\
\mbox{uniform (size orthogonal) } \bar{\mu}^{\ell}_{ft}
\end{array}
\right.
$$

<!--^[Fundamentally weird. Given $W^{\ell}_{ft}, L_{ft}, P_{ft}, Q_{ft}$, how can there be multiple $\mu^{\ell}_{ft}$ estimates (lower-,  upper-bound, intermediate) depending on the market conduct? $Q_{ft}, L_{ft}$ determine MRPL, together with $W^{\ell}_{ft}$ uniquely determine $\mu^{\ell}_{ft}$. Is bounding a right idea? When I have three wage markdowns, "it is 1.2 if there is no collusion, 2 when there is perfect collusion, 1.5 in the intermediate case. No, we have only one set of $W^{\ell}_{ft}, L_{ft}, P_{ft}, Q_{ft}$ and are using the same production function." WTF? ] -->
$$
\underline{\mu}^{\ell}_{ft} \leqslant \mu^{\ell}_{ft} \leqslant \bar{\mu}^{\ell}_{ft}
$$
:::
::::

## Identification of collusion {#sec-identification}

Markup from labour = markup from materials

\begin{align}
\frac{\beta_{\ell}}{\alpha^{\ell}_{ft} \cdot (1 + \tilde{\lambda}_{ft} \cdot \Psi^{\ell})} &= \frac{\beta_m}{\alpha^m_{ft}}\\
&\Leftrightarrow \\
\underbrace{1 + \widetilde{\lambda}_{ft} \Psi^{\ell}}_{=\mu^{\ell}_{ft}} &= \frac{\beta_{\ell} \alpha^{m}_{ft}}{\beta_{m} \alpha^{\ell}_{ft}} \tag{14}
\label{eq14}
\end{align}

* LHS sans $\widetilde{\lambda}_{ft}$ is observed from wage elasticity $\Psi^{\ell}$
* RHS is observed from CD production function $\beta_{\ell}, \beta_{m}$
* $\Psi^{\ell} + \beta_{\ell}, \beta_{m}$ &rArr; $\widetilde{\lambda}_{ft}$
* In general, @Mertens2022 showed:^[This is exactly the same result as Syverson (2024). If both markets are oligopsonistic, expenditure shares and elasticities give a ratio of markdowns $\frac{\mu^{\ell}}{\mu^{M}}$, which is not an interesting object... ]

::: {style="font-size: 80%;"}
$$
\left.
\begin{array}{rrr}
\mbox{Expenditure shares $\alpha_{x}$, materials and labour}\\
\mbox{Output elasticities $\theta^{x}$, materials and labour}\\
\mbox{Material market is non-oligopsonistic}\\
\mbox{Profit maximisation}
\end{array}
\right\}
\; \mbox{gives} \;
\mbox{wage markdown}
$$ 
:::

* Adding a factor supply elasticity gives factor market collusion index.

:::: {.callout-important title='Other approaches to wage markdown' collapse='true'}
 * Revenue based markdown: revenue elasticity of labour + labour cost share + specific form of output demand function &rarr; wage markdown [@Treuren2022]^[$$
\begin{aligned}
\gamma^L_{it} 
&= 
\frac{MRPL_{it}}{W_{it}}\\
&=
\frac{\partial R_{it}}{\partial L_{it}}\frac{1}{W_{it}}\\
&=
\frac{\partial R_{it}}{\partial L_{it}}\frac{L_{it}}{R_{it}}\frac{R_{it}}{W_{it}L_{it}}\\
&=
\frac{\theta^{L}_{R,it}}{\alpha^{R}_{L}}
\end{aligned}
$$
 ] ^[The challenge is in estimating revenue elasticity $\theta^{L}_{R,it}$. Treuren (2022) assumes the inverse output demand to be multiplicably separable in quantity and shocks to use Olley-Pakes inversion. ]
<!--
 * Simultaneous input and output market wedges [@DobbelaereWiersma2024]
 * Simultaneous input and output market wedges under non-Hicks neutrality with log-linear approximation [@AzzamJaumandreuLopez2025]
-->
 * Job search theory based markdown [@HirshJahnManningOberfichtner2023]
::::
<!--* This is neat: If only labour market is oligopsonistic and if there is at least one flexible input purchased from a competitive market, markdown with collusion parameter appears only on LHS-->
* Can test $\widetilde{\lambda}_{ft}=1, s_{ft}$ or just heterogenous within market $i$
* Not an equilibrium condition, so comparative statics is not tenable^[Need to model output response to exogenous wage or coal price shocks. Wage&darr; &rArr; $\alpha^{\ell}_{ft}$&darr;, $\Psi^{\ell}$&darr; &rArr; $\mu^{\ell}_{ft}$&uarr;, $\widetilde{\lambda}_{ft}$&uarr;, but $Q$ (hence $P$) also changes. Also want to know $P_{ft}$&uarr; &rArr; $\alpha^{\ell}_{ft}$&darr;, $\alpha^{m}_{ft}$&darr; with $\frac{\alpha^{m}_{ft}}{\alpha^{\ell}_{ft}}$ unchanged &rArr; $\mu^{\ell}_{ft}$ and $\widetilde{\lambda}_{ft}$ unchanged, but this is wrong...MRPL *must* increase so must $\mu^{\ell}_{ft}$.  Possible: $P_{ft}$&uarr; &rArr; $Q_{ft}$&uarr; &rArr; $\Psi^{\ell}$&uarr; &rArr; $\mu^{\ell}_{ft}$&uarr;. ]
<!--
      * If profit maximised, supply function &rarr; increase outputs hence labour, so operating in larger $\Psi^{\ell}$ segment of labour supply curve &rarr; $\widetilde{\lambda}_{ft}$&darr; if RHS remains unchanged, but cartel does not allow for a large increase in production
      * Note %$P$&uarr; $\simeq$ %$\mu^{\ell}$&uarr; $\simeq$ 35% &lArr; $\Psi^{\ell}$&uarr;, $\widetilde{\lambda}_{ft}$ unchanged (LHS) but on RHS, $\frac{\alpha^{m}_{ft}}{\alpha^{\ell}_{ft}}$ must increase...not tenable. 
-->

To get LHS of \eqref{eq14}, rearrange FOC in \eqref{eq11}.^[Taking the derivative with respect to labor on $Q_{ft} = e^{\beta_l l_{ft} + \beta_m m_{ft} + \beta_k k_{ft} + \omega_{ft}}$ gives $$\frac{\partial Q_{ft}}{\partial L_{ft}} = \frac{\partial Q_{ft}}{\partial l_{ft}} \cdot \frac{\partial l_{ft}}{\partial L_{ft}} = \beta_l \cdot Q_{ft} \cdot \frac{1}{L_{ft}} = \beta_l \frac{Q_{ft}}{L_{ft}}$$ ]
^[From equation (3): $W^l_{it} = L_{it}^{\Psi^l} \eta_{it}$, taking the derivative: $$\frac{\partial W^l_{it}}{\partial L_{it}} = \Psi^l L_{it}^{\Psi^l - 1} \eta_{it} = \Psi^l \frac{W^l_{it}}{L_{it}}$$ ]

\begin{alignat}{2}
W^l_{it} &+ \tilde{\lambda}_{ft} & \frac{\partial W^l_{it}}{\partial L_{it}} L_{it}
&= 
\frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}}\\
&&&\Leftrightarrow\\
W^{\ell}_{it} &+ \tilde{\lambda}_{ft} & \Psi^{\ell} W^{\ell}_{it}
&=
\beta_{\ell} \frac{P_{ft} Q_{ft}}{\mu_{ft} L_{ft}}\\
&&&\Leftrightarrow\\
W^{\ell}_{it}L_{ft} &+ \tilde{\lambda}_{ft} &\Psi^{\ell} W^{\ell}_{it}L_{ft}
&=
\beta_{\ell} \frac{P_{ft} Q_{ft}}{\mu_{ft}}
\end{alignat}

Markup^[Using revenue share of labor: $W^{\ell}_{ft} L_{ft} = \alpha^{\ell}_{ft} P_{ft} Q_{ft}$ ]
\begin{alignat}{2}
\mu_{ft}
&= 
\frac{\beta_{\ell} P_{ft} Q_{ft}}{W^{\ell}_{it} L_{ft} (1 + \tilde{\lambda}_{ft} \Psi^{\ell})}\\
&= 
\frac{\beta_{\ell}}{\alpha^{\ell}_{ft} (1 + \tilde{\lambda}_{ft} \Psi^{\ell})}
\end{alignat}

<!--Materials market is competitive, so the usual formula.^[FOC
$$
\begin{aligned}
W^m_{ft} 
&= 
\frac{\partial Q_{ft}}{\partial M_{ft}} \frac{P_{ft}}{\mu_{ft}}\\
&=
\beta_m \frac{P_{ft} Q_{ft}}{\mu_{ft} M_{ft}}
\end{aligned}
$$
Solve for the markup
$$
\mu_{ft}
= \frac{\beta_m P_{ft} Q_{ft}}{W^m_{ft} M_{ft}}
= \frac{\beta_m}{\alpha^m_{ft}}
$$
 ]-->

## Estimation

Moment conditions for production function (IVs: standard set + lagged regional wage)
$$
E[u_{ft} | (l_{fr-1}, m_{fr-1}, k_{fr}, w^{agri}_{r-1})_{r \in \{2,...,t\}}] = 0 \tag{16}
$$
GMM moment conditions for production function estimation
\begin{align}
E[
&q_{ft} - \rho q_{ft-1} - \beta_{0}(1-\rho) - \beta_{l}(l_{ft} - \rho l_{ft-1})\\
&- \beta_{m}(m_{ft} - \rho m_{ft-1}) - \beta_{k}(k_{ft} - \rho k_{ft-1}) |\\
& l_{ft-1}, m_{ft-1}, k_{ft}, k_{ft-1}, w^{agri}_{t-1}] = 0 \tag{17}
\end{align}

## Data

1. Administration des Mines (annual reports on mines)
1. Employer Association membership
    * monthly Bulletin of the Union des Charbonnages, Mines et Usines Métallurgiques de la Province de Liège
    * Association Charbonnière et l’Industrie Houillière des Bassins de Charleroi et de la Basse-Sambre
1. Cartel membership (De Leener data) 
1. Other: Opening dates of railroad and tramway stations, agricultural wages

## Research design

* $\mu^{\ell}_{ft}$ using the production approach under material market competition
* Uses financial statement data from historical records when cartel was legal
* Only after using material FOC, we identify $\mu^{\ell}_{ft}$ and collusion index $\widehat{\lambda}_{ft}$
* Under oligopsony, $\mu^{\ell}_{ft} \in[\underline{\mu}^{\ell}_{ft}, \bar{\mu}^{\ell}_{ft}]$, depending on the extent of collusion
* The difference from lower- and upper-bound = degree of collusion
* Hypothesis: Degree of collusion vs. collusion regimes: collusion&uarr; (cartel) &rArr; markdown&uarr;^[The paper lacks a theoretical analysis of exogenous price increase when we are talking about the coal price cartel. The price increase could have been due to exogenous shocks/shifts. What happens to $\mu^{\ell}_{ft}$ and implied $\widehat{\lambda}_{ft}$ when the price increase exogenously? I.e., when the market conduct remains the same in the degree of collusion, does a coal price increase leave the estimated $\mu^{\ell}_{ft}$ unchanged? ] ^[%$P$&uarr; $\simeq$ %$\mu^{\ell}$&uarr;, roughly 35% on eyeballs, early 1900's. A possible exogenous impact: Price&uarr;exogenously &rArr; $\frac{MRPL_{ft}\uparrow}{W^{\ell}_{ft}}=\mu^{\ell}_{ft}$&uarr; A possible endogenous response: $P_{ft}$&uarr; &rArr; $Q_{ft}$&uarr; &rArr; $\Psi^{\ell}$&uarr; &rArr; $\mu^{\ell}_{ft}$&uarr;. The former does not aggravate misallocation, but the latter does. ]

## Results

::: {.column-margin}
![Table 3](./Tab3.jpg)

![Figure 4A](./Fig4A.jpg)

![Figure 4B](./Fig4B.jpg)
:::

:::: {.callout-important title='Wage markdown is pervasive' collapse='true'}
Proportions of MPRL workers receive are:

* `r round((1/1.5)*100, 0)`% (40's - 70's) &rarr; `r round((1/1.75)*100, 0)`% (80's - 90's) &rarr; `r round((1/2.2)*100, 0)`% (1897 - )^[I can see: %$P$&uarr; $\simeq$ %$\mu^{\ell}$&uarr;, roughly 35% on eyeballs, early 1900's. Firms did not pass coal price windfall to wages. Why did not they have to? Collusive oligopsony.  ] 
::::

:::: {.callout-important title='Markdown $\propto$ Employer Association membership (Table 2, B)' collapse='true'}
* (pre cartel) $+$ &rarr; zero (post cartel)
* Cartel took over the role of collusion device
::::

:::: {.callout-important title='Markdown $\propto$ size (Table 3, TWFE)' collapse='true'}
* $+$ only for non-cartel firms &larr; Cournot non-collusion
* Zero for cartel firms &larr; (almost) Perfect collusion = size insensitive
::::

:::: {.callout-important title='"Decomposition": Collusive vs. noncollusive origin of markdowns (Figure 4A)' collapse='true'}
* Median $\mu^{\ell}_{ft}$&uarr; after 1897 cartel formation, surpassing the upperbound (perfect collusion $\bar{\mu}^{\ell}$)
* Noncollusive $\underline{\mu}^{\ell}_{ft}$&uarr; only up to 1987, then plateaued
* Concludes: Wage collusion &rArr; post cartel $\mu^{\ell}_{ft}$&uarr;^["Given that the noncollusive markdown does not grow after 1900, the vast increase in markdowns after the introduction of the coal cartel appears to have been entirely driven by wage collusion." ]
* But can we say collusion is weaker under Employer Associations? Coal price levels increased by less in pre-cartel era.
* Strange features:^[No mention on these features. ]
  * $\bar{\mu}^{\ell}_{ft}$ is constant at 2
  * $\bar{\mu}^{\ell}_{ft}=2$ &rArr; $\Psi^{\ell}=1$
  * $Med\left(\mu^{\ell}_{ft}\right)>\bar{\mu}^{\ell}_{ft}$ &rArr; $Med\left(\tilde{\lambda}_{ft}\right)>1$?
* Perfectly collusive $\bar{\mu}^{\ell}_{ft}$ estimates <Red>may not be reliable</Red>
::::

:::: {.callout-important title='Collusion index $\hat{\lambda}_{it}$&uarr; detects cartel based markdown collusion (Figure 4B)' collapse='true'}
* Questionable $\bar{\mu}^{\ell}_{ft}=2$ enters as a constant in the denominator, so the relative changes in the index are not affected (but the level is)

   > From C.2. Testing for employer collusion: *the collusion estimates seem to be able to detect collusion due to the cartel, without requiring any a priori information about the cartel.* 
::::

:::: {.callout-important title='Sensitivity analysis' collapse='true'}
* Input adjustment costs (frictions): Immediate response of labour costs to 1871 demand shock
* Factor biased tech change:^[Size sensitivity of $\mu^{\ell}=1+s\Psi^{\ell}$ can be due to tech change, not necessarily to collusion. ] After increased investments &larr; No change in labour share,  low correation between $\mu^{\ell}_{ft}$ and horsepower of machines, capital*labour interaction in prod function = 0
* Unionisation and socialist party popularity: Region's strike participation is low, no difference in markdown between socialist and non-socialist districts
* Labour supply: Time variant coefficients, linear time trend, different labour market definitions, incorporating expansion of railway, changes in IVs
::::

:::: {.callout-important title='Markup is smaller than 1 (Table 1)' collapse='true'}

* .714 - .816, but statistically indistinguishable from 1
* Despite cartel formation, point estimates suggest oligpsony in coal market, effectively handing over the wage rents &larr; competitiion with imported coal
* $\mu^{\mathrm{tot}}_{ft}\equiv 1+(\mu_{ft}-1)+(\mu^{\ell}_{ft}-1)$^[ $\frac{\pi_{\ell}}{P}=\frac{MR_{\ell}}{W}\frac{W}{P}-\frac{MC_{\ell}}{P}$, or $\mu^{\mathrm{tot}}=\frac{W}{P}\mu^{\ell}-\frac{1}{\mu}$...? ]

::::

:::: {.callout-important title='Simulation if there was no cartel (with/without coal market Cournot competition)' collapse='true'}

Cournot vs. 2 scenarios: Exogenous output price, endogenous output price (output market full collusion)

* Wage: -10.3%, -25.9% (coal market collusion)
* Employment: -10.2%, -24.9% (coal market collusion)

Employer Assoc. vs. 2 scenarios: Exogenous output price, endogenous output price (output market full collusion)

* Wage: -5.9%, -16.7% (coal market collusion)
* Employment: -5.9%, -16.6% (coal market collusion)

When under full collusion, $\mu^{\ell}_{ft}$&uarr;, $W^{\ell}_{ft}$&darr;, $L_{ft}$&darr;, $Q_{ft}$&darr;, relative to no collusion/intermediate case
::::

<!--
1. Wage markdown is pervasive, proportions of MPRL workers receive are: 
   * `r round((1/1.5)*100, 0)`% (40's - 70's) &rarr; `r round((1/1.75)*100, 0)`% (80's - 90's) &rarr; `r round((1/2.2)*100, 0)`% (1897 - ) 
1. Markdown $\propto$ Employer Association membership (Table 2, B)
   * (pre cartel) $+$ &rarr; zero (post cartel)
   * Cartel took over the role of collusion device
1. Markdown $\propto$ size (Table 3, TWFE)
   * $+$ only for non-cartel firms &larr; Cournot non-collusion
   * Zero for cartel firms &larr; (almost) Perfect collusion = size insensitive
1. "Decomposition": Collusive vs. noncollusive origin of markdowns (Figure 4A)
   * Median $\mu^{\ell}_{ft}$&uarr; after 1897 cartel formation, surpassing the upperbound (perfect collusion $\bar{\mu}^{\ell}$)
   * Noncollusive $\underline{\mu}^{\ell}_{ft}$&uarr; only up to 1987, then plateaued
   * Concludes: Post cartel $\mu^{\ell}_{ft}$&uarr; &larr; wage collusion^["Given that the noncollusive markdown does not grow after 1900, the vast increase in markdowns after the introduction of the coal cartel appears to have been entirely driven by wage collusion." ]
   * Strange features:^[No mention on these features. ]
      * $\bar{\mu}^{\ell}_{ft}$ is constant at 2
      * $\bar{\mu}^{\ell}_{ft}=2$ &rArr; $\Psi^{\ell}=1$
      * $Med\left(\mu^{\ell}_{ft}\right)>\bar{\mu}^{\ell}_{ft}$ &rArr; $Med\left(\tilde{\lambda}_{ft}\right)>1$?
   * Perfectly collusive $\bar{\mu}^{\ell}_{ft}$ estimates <Red>may not be reliable</Red>
1. Collusion index $\hat{\lambda}_{it}$&uarr; after cartel formation (Figure 4B)
   * Questionable $\bar{\mu}^{\ell}_{ft}=2$ enters as a constant in the denominator, so the relative changes in the index are not affected (but the level is)

   > the collusion estimates seem to be able to detect collusion due to the cartel, without requiring any a priori information about the cartel. (C.2. Testing for employer collusion)
1. Sensitivity analysis
   * Input adjustment costs (frictions): Immediate response of labour costs to 1871 demand shock
   * Factor biased tech change:^[Size sensitivity of $\mu^{\ell}=1+s\Psi^{\ell}$ can be due to tech change, not necessarily to collusion. ] After increased investments &larr; No change in labour share,  low correation between $\mu^{\ell}_{ft}$ and horsepower of machines, capital*labour interaction in prod function = 0
   * Unionisation and socialist party popularity: Region's strike participation is low, no difference in markdown between socialist and non-socialist districts
1. Simulation (w/o coal market Cournot competition) if there was no cartel: 
   * Wage: +10.3%, +25.9% (coal market Cournot)
   * Employment: +10.2%, +24.9% (coal market Cournot)
-->

## 感想

* Rubensは研究材料の幅が広い: 中国たばこ農家のデータを使って下流の企業統合がmarkdownに与えた影響を推計、ベルギーの古文書をデジタル化して共謀とmarkdownの関係を推計...IO専門家?
* 価格カルテルが合法だったときの古いデータを使うことで、素のカルテル効果を鮮明に示しているのは素晴らしい
* 下流の市場慣行&rarr;上流の市場慣行、しかも、下流の寡占買い手が鉱山労働者から得たレントを吸い上げている可能性を示す
* wage markdown=oligopsonyの識別そのものは目新しくないが、Cournot仮定下で共謀度の識別という付加価値がある
* 雇用者協会よりも価格カルテルの方が共謀の程度が強いと当然のように書いているが、なぜ?
* wage markdownはproduction approachが適用の範囲の広さで群を抜いているが、Hicks neutralityとmaterial market competition以外あまりに制約をかけないので、推計精度が低い(技術の同質的な大標本が必要)かも

### 2025-06-18 posts/DelabastitaRubens/DelabastitaRubens2.qmd
added DelabastitaRubens repo and ethics repo

---
title: "Delabastita and Rubens 2025"
author: "claude 4 sonnet, instructed by me"
date: last-modified
date-format: "MMMM DD, YYYY, HH:MM ZZ"
execute:
  echo: false
engine: knitr
pdf-engine: lualatex
format:
  html:
    standalone: true
    embed-resources: true
    toc: true
    toc-location: left
    grid:
      margin-width: 350px
####  pdf: default
  docx: default
  epub: default
theme: 
  light: sky
  dark: darkly
title-block-banner: "#f0f3f5"
title-block-banner-color: "black"
callout-appearance: simple
callout-icon: false
bibliography: ../seiro.bib
link-citations: true
reference-location: margin
citation-location: margin
#### # quarto render DelabastitaRubens2.qmd
#### c:\seiro\languages\R\R-4.4.1\bin\x64\Rscript.exe -e "path <- 'c:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/DelabastitaRubens/'; rmarkdown::render(paste0(path, 'DelabastitaRubens.Rmd'), output_format = 'all')"
---
```{css CSS setting, echo=FALSE, results = "hide"}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Noto S	ans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
/* Define a margin before hX (header level X) element */
h1  {
  margin-top: 3ex;
  margin-bottom: 3ex;
  /* background: #c2edff; */ /*背景色*/
  padding: 0.5em;/*文字まわり（上下左右）の余白*/
}
h2  {
  margin-top: 2ex;
  margin-bottom: 2ex;
  padding: 0.5em;/*文字周りの余白*/
  ####color: #010101;/*文字色*/
  color: #87a5f2;
  /* background: #eaf3ff; */ /*背景色*/
  /* border-bottom: solid 3px #516ab6; */ /*下線*/
}
```

<style type="text/css">
  body, td{
  font-size: 16pt;
}
</style>

# Equations and Variable Definitions from Coal Mining Study

## Main Equations

### Equation (1) - Cobb-Douglas Production Function
$$
q_{ft} = \beta_{l} l_{ft} + \beta_{m} m_{ft} + \beta_{k} k_{ft} + \omega_{ft} \tag{1}
$$

### Equation (2) - TFP Transition Process
$$
\omega_{ft} = \rho \omega_{ft-1} + u_{ft} \tag{2}
$$

### Labor Supply Function - Equation (3)
$$
W^{l}_{it} = L_{it}^{\Psi^{l}} \eta_{it} \tag{3}
$$

### Cost Minimization Objective - Equation (4)
$$
\min_{L_{ft}, M_{ft}} \left\{ \sum_{g \in F_{i(f)t}} \left( \lambda_{fgt} (L_{gt} W^{l}_{gt} + M_{gt} W^{m}_{gt}) \right) - MC_{ft} \left( Q(L_{ft}, M_{ft}, K_{ft}, \Omega_{ft}; \beta) - Q_{ft} \right) \right\} \tag{4}
$$

### No Collusion Cost Minimization - Equation (5)
$$
\min_{L_{ft}, M_{ft}} \left( L_{ft} W^{l}_{ft} + M_{ft} W^{m}_{ft} \right) - MC_{ft} \left( Q_{ft} - Q(L_{ft}, M_{ft}, K_{ft}, \Omega_{ft}; \beta) \right) \tag{5}
$$

### First-Order Condition for Labor (No Collusion) - Equation (6)
$$
L_{ft} \frac{\partial W^{l}_{it}}{\partial L_{it}} + W^{l}_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} \tag{6}
$$

### Markdown Under No Collusion - Equation (7)
$$
\widetilde{m}^{l}_{ft} = 1 + s^{l}_{ft} \Psi^{l} \tag{7}
$$

### Perfect Collusion Cost Minimization - Equation (8)
$$
\min_{L_{ft}, M_{ft}} \left\{ \sum_{g \in F_{i(f)t}} \left( L_{gt} W^{l}_{gt} + M_{gt} W^{m}_{gt} \right) - MC_{ft} \left( Q_{ft} - Q(L_{ft}, M_{ft}, K_{ft}, \Omega_{ft}; \beta) \right) \right\} \tag{8}
$$

### First-Order Condition for Labor (Perfect Collusion) - Equation (9)
$$
L_{it} \frac{\partial W^{l}_{it}}{\partial L_{it}} + W^{l}_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} \tag{9}
$$

### Markdown Under Perfect Collusion - Equation (10)
$$
\overline{m}^{l}_{ft} = 1 + \Psi^{l} \tag{10}
$$

### General First-Order Condition - Equation (11)
$$
W^{l}_{it} + \widetilde{\lambda}_{ft} \frac{\partial W^{l}_{it}}{\partial L_{it}} L_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} \tag{11}
$$

### General Markdown Expression - Equation (12)
$$
m^{l}_{ft} = 1 + \widetilde{\lambda}_{ft} \Psi^{l} \tag{12}
$$

### Markup from Materials - Equation (13)
$$
\mu_{ft} = \frac{\beta_{m}}{\alpha^{m}_{ft}} \tag{13}
$$

### Core Identification Equation - Equation (14)
$$
m^{l}_{ft} = 1 + \widetilde{\lambda}_{ft} \Psi^{l} = \frac{\beta_{l} \alpha^{m}_{ft}}{\beta_{m} \alpha^{l}_{ft}} \tag{14}
$$

### Conduct Parameter Definition - Equation (15)
$$
\widehat{\lambda}_{ft} = \frac{m^{l}_{ft} - \widetilde{m}^{l}_{ft}}{\overline{m}^{l}_{ft} - \widetilde{m}^{l}_{ft}} \tag{15}
$$

### Moment Conditions for Production Function - Equation (16)
$$
E[u_{ft} | (l_{fr-1}, m_{fr-1}, k_{fr}, w^{agri}_{r-1})_{r \in \{2,...,t\}}] = 0 \tag{16}
$$

### GMM Moment Conditions - Equation (17)
$$
E[q_{ft} - \rho q_{ft-1} - \beta_{0}(1-\rho) - \beta_{l}(l_{ft} - \rho l_{ft-1}) - \beta_{m}(m_{ft} - \rho m_{ft-1}) - \beta_{k}(k_{ft} - \rho k_{ft-1}) | (l_{ft-1}, m_{ft-1}, k_{ft}, k_{ft-1}, w^{agri}_{t-1})] = 0 \tag{17}
$$

## Additional Key Definitions

### Markdown Definition
$$
m^{l}_{ft} \equiv \frac{MRPL_{ft}}{W^{l}_{ft}}
$$

### Percentage Markdown Definition
$$
d^{l}_{ft} \equiv \frac{MRPL_{ft} - W^{l}_{ft}}{MRPL_{ft}} = \frac{m^{l}_{ft} - 1}{m^{l}_{ft}}
$$

### Markup Definition
$$
\mu_{ft} \equiv \frac{P_{ft}}{MC_{ft}}
$$

### Labor Market Share
$$
s^{l}_{ft} \equiv \frac{L_{ft}}{L_{it}}
$$

### Revenue Share of Labor
$$
\alpha^{l}_{ft} \equiv \frac{W^{l}_{ft} L_{ft}}{P_{ft} Q_{ft}}
$$

### Revenue Share of Materials
$$
\alpha^{m}_{ft} \equiv \frac{W^{m}_{ft} M_{ft}}{P_{ft} Q_{ft}}
$$

### Inverse Labor Supply Elasticity (Firm-level)
$$
\theta^{l}_{ft} \equiv \frac{\partial W^{l}_{ft}}{\partial L_{ft}} \frac{L_{ft}}{W^{l}_{ft}}
$$

### Inverse Materials Supply Elasticity (Firm-level)
$$
\theta^{m}_{ft} \equiv \frac{\partial W^{m}_{ft}}{\partial M_{ft}} \frac{M_{ft}}{W^{m}_{ft}}
$$

## Variable Definitions

| Variable | Definition |
|----------|------------|
| $Q_{ft}$ | Output (tonnage of coal) extracted by firm $f$ in year $t$ |
| $L_{ft}$ | Amount of effective labor throughout the year for firm $f$ in year $t$ |
| $M_{ft}$ | Amount of intermediate inputs purchased by firm $f$ in year $t$ |
| $K_{ft}$ | Capital stock (steam engines) used by firm $f$ in year $t$ |
| $q_{ft}$ | Logarithm of output $Q_{ft}$ |
| $l_{ft}$ | Logarithm of labor $L_{ft}$ |
| $m_{ft}$ | Logarithm of materials $M_{ft}$ |
| $k_{ft}$ | Logarithm of capital $K_{ft}$ |
| $\beta_{l}$ | Output elasticity of labor |
| $\beta_{m}$ | Output elasticity of materials |
| $\beta_{k}$ | Output elasticity of capital |
| $\omega_{ft}$ | Log total factor productivity |
| $u_{ft}$ | Unexpected productivity shock |
| $\rho$ | Serial correlation parameter in productivity process |
| $W^{l}_{it}$ | Wage in labor market $i$ in year $t$ |
| $W^{l}_{ft}$ | Wage for firm $f$ in year $t$ |
| $W^{m}_{ft}$ | Price of materials for firm $f$ in year $t$ |
| $L_{it}$ | Market-level employment in market $i$ in year $t$ |
| $\Psi^{l}$ | Inverse market-level labor supply elasticity |
| $\eta_{it}$ | Market-specific residual in labor supply |
| $\lambda_{fgt}$ | Collusion weight that firm $f$ puts on firm $g$'s costs |
| $\widetilde{\lambda}_{ft}$ | Conduct parameter (firm-level aggregate of bilateral conduct parameters) |
| $\widehat{\lambda}_{ft}$ | Normalized conduct parameter ranging from 0 to 1 |
| $MC_{ft}$ | Marginal cost for firm $f$ in year $t$ |
| $P_{ft}$ | Coal price for firm $f$ in year $t$ |
| $MRPL_{ft}$ | Marginal revenue product of labor for firm $f$ in year $t$ |
| $m^{l}_{ft}$ | Wage markdown (ratio of MRPL to wage) |
| $\widetilde{m}^{l}_{ft}$ | Wage markdown under no collusion |
| $\overline{m}^{l}_{ft}$ | Wage markdown under perfect collusion |
| $d^{l}_{ft}$ | Percentage wage markdown |
| $\mu_{ft}$ | Price markup (ratio of price to marginal cost) |
| $s^{l}_{ft}$ | Labor market share of firm $f$ |
| $\alpha^{l}_{ft}$ | Revenue share of labor for firm $f$ in year $t$ |
| $\alpha^{m}_{ft}$ | Revenue share of materials for firm $f$ in year $t$ |
| $\theta^{l}_{ft}$ | Inverse firm-level labor supply elasticity |
| $\theta^{m}_{ft}$ | Inverse firm-level materials supply elasticity |
| $F_{i(f)t}$ | Set of firms in market $i$ (where firm $f$ operates) in year $t$ |
| $w^{agri}_{t-1}$ | Agricultural wages in Belgium in year $t-1$ (instrument) |

## Tab-Separated Variable Definitions

```
Variable	Definition
Q_ft	Output (tonnage of coal) extracted by firm f in year t
L_ft	Amount of effective labor throughout the year for firm f in year t
M_ft	Amount of intermediate inputs purchased by firm f in year t
K_ft	Capital stock (steam engines) used by firm f in year t
q_ft	Logarithm of output Q_ft
l_ft	Logarithm of labor L_ft
m_ft	Logarithm of materials M_ft
k_ft	Logarithm of capital K_ft
beta_l	Output elasticity of labor
beta_m	Output elasticity of materials
beta_k	Output elasticity of capital
omega_ft	Log total factor productivity
u_ft	Unexpected productivity shock
rho	Serial correlation parameter in productivity process
W_l_it	Wage in labor market i in year t
W_l_ft	Wage for firm f in year t
W_m_ft	Price of materials for firm f in year t
L_it	Market-level employment in market i in year t
Psi_l	Inverse market-level labor supply elasticity
eta_it	Market-specific residual in labor supply
lambda_fgt	Collusion weight that firm f puts on firm g's costs
lambda_tilde_ft	Conduct parameter (firm-level aggregate of bilateral conduct parameters)
lambda_hat_ft	Normalized conduct parameter ranging from 0 to 1
MC_ft	Marginal cost for firm f in year t
P_ft	Coal price for firm f in year t
MRPL_ft	Marginal revenue product of labor for firm f in year t
m_l_ft	Wage markdown (ratio of MRPL to wage)
m_l_tilde_ft	Wage markdown under no collusion
m_l_bar_ft	Wage markdown under perfect collusion
d_l_ft	Percentage wage markdown
mu_ft	Price markup (ratio of price to marginal cost)
s_l_ft	Labor market share of firm f
alpha_l_ft	Revenue share of labor for firm f in year t
alpha_m_ft	Revenue share of materials for firm f in year t
theta_l_ft	Inverse firm-level labor supply elasticity
theta_m_ft	Inverse firm-level materials supply elasticity
F_i_f_t	Set of firms in market i (where firm f operates) in year t
w_agri_t_minus_1	Agricultural wages in Belgium in year t-1 (instrument)
```

# Table 1

::: {#tbl-table1 html-table-processing="true"}

```{=html}
{{< include "./Tab1.html" >}}
```

:::

## Derivation of (14)

Looking at equation (14) and the context provided, I can explain how the wage markdown is derived by combining the two markup estimates from the production approach.

### Cost minimization

From the cost minimization problem, the first-order condition for labor (equation 11) is:
$$W^l_{it} + \tilde{\lambda}_{ft} \cdot \frac{\partial W^l_{it}}{\partial L_{it}} \cdot L_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \cdot \frac{P_{ft}}{\mu_{ft}}$$

Where the right side is the marginal revenue product of labor (MRPL).

Rearranging and using:
- $\beta_l$ = output elasticity of labor
- $\alpha^l_{ft} = \frac{W^l_{ft} \cdot L_{ft}}{P_{ft} \cdot Q_{ft}}$ = revenue share of labor

The markup derived from the labor first-order condition becomes:
$$\mu_{ft} = \frac{\beta_l}{\alpha^l_{ft} \cdot (1 + \tilde{\lambda}_{ft} \cdot \Psi^l)}$$

From (11)

$$W^l_{it} + \tilde{\lambda}_{ft} \frac{\partial W^l_{it}}{\partial L_{it}} L_{it} = \frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}}$$

From the Cobb-Douglas production function: $Q_{ft} = e^{\beta_l l_{ft} + \beta_m m_{ft} + \beta_k k_{ft} + \omega_{ft}}$

Taking the derivative with respect to labor:
$$\frac{\partial Q_{ft}}{\partial L_{ft}} = \frac{\partial Q_{ft}}{\partial l_{ft}} \cdot \frac{\partial l_{ft}}{\partial L_{ft}} = \beta_l \cdot Q_{ft} \cdot \frac{1}{L_{ft}} = \beta_l \frac{Q_{ft}}{L_{ft}}$$

Substitute this into the RHS of equation (11)
$$\frac{\partial Q_{ft}}{\partial L_{ft}} \frac{P_{ft}}{\mu_{ft}} = \beta_l \frac{Q_{ft}}{L_{ft}} \frac{P_{ft}}{\mu_{ft}} = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft} L_{ft}}$$

Multiply both sides by $L_{ft}$
$$L_{ft} W^l_{it} + \tilde{\lambda}_{ft} \frac{\partial W^l_{it}}{\partial L_{it}} L_{it}^2 = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

From equation (3): $W^l_{it} = L_{it}^{\Psi^l} \eta_{it}$

Taking the derivative:
$$\frac{\partial W^l_{it}}{\partial L_{it}} = \Psi^l L_{it}^{\Psi^l - 1} \eta_{it} = \Psi^l \frac{W^l_{it}}{L_{it}}$$

Substitute this back to the above
$$L_{ft} W^l_{it} + \tilde{\lambda}_{ft} \Psi^l \frac{W^l_{it}}{L_{it}} L_{it}^2 = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

$$L_{ft} W^l_{it} + \tilde{\lambda}_{ft} \Psi^l W^l_{it} L_{it} = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

$$W^l_{it} L_{ft} (1 + \tilde{\lambda}_{ft} \Psi^l) = \beta_l \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

Markup
$$\mu_{ft} = \frac{\beta_l P_{ft} Q_{ft}}{W^l_{it} L_{ft} (1 + \tilde{\lambda}_{ft} \Psi^l)}$$

Define the revenue share of labor: $\alpha^l_{ft} = \frac{W^l_{ft} L_{ft}}{P_{ft} Q_{ft}}$, therefore: $W^l_{ft} L_{ft} = \alpha^l_{ft} P_{ft} Q_{ft}$

$$\mu_{ft} = \frac{\beta_l P_{ft} Q_{ft}}{\alpha^l_{ft} P_{ft} Q_{ft} (1 + \tilde{\lambda}_{ft} \Psi^l)} = \frac{\beta_l}{\alpha^l_{ft} (1 + \tilde{\lambda}_{ft} \Psi^l)}$$

This gives us the markup expression derived from the labor first-order condition, which incorporates the wage markdown term $(1 + \tilde{\lambda}_{ft} \Psi^l)$ in the denominator.

### Markup from materials (following De Loecker & Warzynski 2012)

For materials, since firms are price-takers (competitive input market), the standard markup formula applies:
$$\mu_{ft} = \frac{\beta_m}{\alpha^m_{ft}}$$

Where $\alpha^m_{ft}$ is the revenue share of materials.

From the cost minimization problem, the first-order condition for materials is:
$$W^m_{ft} = \frac{\partial Q_{ft}}{\partial M_{ft}} \frac{P_{ft}}{\mu_{ft}}$$

From the Cobb-Douglas production function: $Q_{ft} = e^{\beta_l l_{ft} + \beta_m m_{ft} + \beta_k k_{ft} + \omega_{ft}}$

Taking the derivative with respect to materials:
$$\frac{\partial Q_{ft}}{\partial M_{ft}} = \frac{\partial Q_{ft}}{\partial m_{ft}} \cdot \frac{\partial m_{ft}}{\partial M_{ft}} = \beta_m \cdot Q_{ft} \cdot \frac{1}{M_{ft}} = \beta_m \frac{Q_{ft}}{M_{ft}}$$

Substitute this into the RHS

$$W^m_{ft} = \beta_m \frac{Q_{ft}}{M_{ft}} \frac{P_{ft}}{\mu_{ft}} = \beta_m \frac{P_{ft} Q_{ft}}{\mu_{ft} M_{ft}}$$

$$W^m_{ft} M_{ft} = \beta_m \frac{P_{ft} Q_{ft}}{\mu_{ft}}$$

Solve for the markup
$$\mu_{ft} = \frac{\beta_m P_{ft} Q_{ft}}{W^m_{ft} M_{ft}}$$

Define the revenue share of materials: $\alpha^m_{ft} = \frac{W^m_{ft} M_{ft}}{P_{ft} Q_{ft}}$

Therefore: $W^m_{ft} M_{ft} = \alpha^m_{ft} P_{ft} Q_{ft}$

$$\mu_{ft} = \frac{\beta_m P_{ft} Q_{ft}}{\alpha^m_{ft} P_{ft} Q_{ft}} = \frac{\beta_m}{\alpha^m_{ft}}$$

This gives us the standard De Loecker & Warzynski (2012) markup formula:
$$\mu_{ft} = \frac{\beta_m}{\alpha^m_{ft}}$$

### Equate the two markup expressions

Since both expressions equal $\mu_{ft}$:
$$\frac{\beta_l}{\alpha^l_{ft} \cdot (1 + \tilde{\lambda}_{ft} \cdot \Psi^l)} = \frac{\beta_m}{\alpha^m_{ft}}$$

Rearranging:
$$1 + \tilde{\lambda}_{ft} \cdot \Psi^l = \frac{\beta_l \cdot \alpha^m_{ft}}{\beta_m \cdot \alpha^l_{ft}}$$

From equation (12), we know that:
$$m^l_{ft} = 1 + \tilde{\lambda}_{ft} \cdot \Psi^l$$

Therefore:
$$m^l_{ft} = \frac{\beta_l \cdot \alpha^m_{ft}}{\beta_m \cdot \alpha^l_{ft}}$$

This is equation (14) in the paper.

The wage markdown $m^l_{ft}$ can be calculated in two equivalent ways:

1. **Labor supply approach**: $1 + \tilde{\lambda}_{ft} \cdot \Psi^l$ (depends on conduct parameter and labor supply elasticity)
2. **Production approach**: $\frac{\beta_l \cdot \alpha^m_{ft}}{\beta_m \cdot \alpha^l_{ft}}$ (depends only on production function parameters and cost shares)

### 2025-06-18 posts/DelabastitaRubens/SyversonMath.qmd
added DelabastitaRubens repo and ethics repo

---
title: "Syverson (2024): Markups and markdowns"
author: "claude 4 sonnet, instructed by me"
date: last-modified
date-format: "MMMM DD, YYYY, HH:MM ZZ"
execute:
  echo: false
engine: knitr
pdf-engine: lualatex
format:
  html:
    standalone: true
    embed-resources: true
    toc: true
    toc-location: left
    grid:
      margin-width: 350px
####  pdf: default
  docx: default
  epub: default
theme: 
  light: sky
  dark: darkly
title-block-banner: "#f0f3f5"
title-block-banner-color: "black"
callout-appearance: simple
callout-icon: false
bibliography: ../seiro.bib
link-citations: true
reference-location: margin
citation-location: margin
#### # quarto render SyversonMath.qmd
#### c:\seiro\languages\R\R-4.4.1\bin\x64\Rscript.exe -e "path <- 'c:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/DelabastitaRubens/'; rmarkdown::render(paste0(path, 'SyversonMath.Rmd'), output_format = 'all')"
---
```{css CSS setting, echo=FALSE, results = "hide"}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Noto S	ans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
/* Define a margin before hX (header level X) element */
h1  {
  margin-top: 3ex;
  margin-bottom: 3ex;
  /* background: #c2edff; */ /*背景色*/
  padding: 0.5em;/*文字まわり（上下左右）の余白*/
}
h2  {
  margin-top: 2ex;
  margin-bottom: 2ex;
  padding: 0.5em;/*文字周りの余白*/
  ####color: #010101;/*文字色*/
  color: #87a5f2;
  /* background: #eaf3ff; */ /*背景色*/
  /* border-bottom: solid 3px #516ab6; */ /*下線*/
}
```

<style type="text/css">
  body, td{
  font-size: 16pt;
}
</style>

This is all done by claude sonnet v4

# Mathematical Derivations from Syverson (2024): Markups and Markdowns

## Section 1.1: Markups, Costs, and Inflation

### Starting Point
The basic relationship:
$$P = \mu \cdot C$$

### Derivation of Growth Rate Relationship

To convert to growth rates, we take the natural logarithm of both sides:
$$\ln(P) = \ln(\mu) + \ln(C)$$

Taking the total differential:
$$d\ln(P) = d\ln(\mu) + d\ln(C)$$

Since $d\ln(x) = \frac{dx}{x}$ (which approximates the growth rate for small changes):
$$\frac{dP}{P} = \frac{d\mu}{\mu} + \frac{dC}{C}$$

Therefore:
$$\text{Growth in } P \approx \text{Growth in } \mu + \text{Growth in } C$$

**Explanation**: This follows from the logarithmic differentiation property. When we have a product $P = \mu \cdot C$, the growth rate of the product equals the sum of the growth rates of its components.

## Section 1.2: Markups, Profit Shares, and Scale Elasticities

### Step 1: Basic Markup Definition and Manipulation

Starting with the markup definition:
$$\mu \equiv \frac{P}{MC}$$

Multiply and divide by average cost (AC):
$$\mu = \frac{P}{MC} \cdot \frac{AC}{AC} = \frac{P}{AC} \cdot \frac{AC}{MC}$$

**Explanation**: This is valid algebraic manipulation - multiplying by $\frac{AC}{AC} = 1$ doesn't change the value.

### Step 2: Rewrite Price-to-Average-Cost Ratio

The ratio $\frac{P}{AC}$ can be rewritten by multiplying numerator and denominator by quantity $Q$:
$$\frac{P}{AC} = \frac{P \cdot Q}{AC \cdot Q} = \frac{Revenue}{Total Cost} = \frac{R}{TC}$$

**Explanation**: Since $Revenue = P \times Q$ and $Total Cost = AC \times Q$, this transformation is straightforward.

### Step 3: Define Scale Elasticity

The scale elasticity $\nu$ is defined as:
$$\nu \equiv \frac{AC}{MC}$$

This is equivalent to the inverse of the elasticity of costs with respect to quantity. To see why:

For any cost function $C(Q)$, the cost elasticity is:
$$\frac{d\ln C}{d\ln Q} = \frac{dC/C}{dQ/Q} = \frac{dC}{dQ} \cdot \frac{Q}{C} = MC \cdot \frac{Q}{TC} = \frac{MC}{AC}$$

Therefore: $\nu = \frac{AC}{MC} = \frac{1}{\text{cost elasticity}}$

**Explanation**: When $MC < AC$, we have economies of scale ($\nu > 1$). When $MC > AC$, we have diseconomies of scale ($\nu < 1$).

### Step 4: Combine Results

Substituting our results:
$$\mu = \frac{R}{TC} \cdot \nu$$

### Step 5: Introduce Profit Share

Define profit share as:
$$s_\pi \equiv \frac{R - TC}{R}$$

This can be rearranged to express the revenue-to-total-cost ratio:
$$\frac{R - TC}{R} = s_\pi$$
$$\frac{R}{R} - \frac{TC}{R} = s_\pi$$
$$1 - \frac{TC}{R} = s_\pi$$
$$\frac{TC}{R} = 1 - s_\pi$$
$$\frac{R}{TC} = \frac{1}{1 - s_\pi}$$

### Step 6: Final Result

Substituting back:
$$\mu = \frac{R}{TC} \cdot \nu = \frac{1}{1 - s_\pi} \cdot \nu$$

Therefore:
$$\boxed{\mu = \frac{\nu}{1 - s_\pi}}$$

## Section 1.3: Markups and Elasticity of Demand

### Step 1: Profit Maximization First-Order Condition

A profit-maximizing firm facing residual demand $q(p)$ maximizes:
$$\pi = p \cdot q(p) - C(q(p))$$

Taking the derivative with respect to price $p$ and setting equal to zero:
$$\frac{d\pi}{dp} = q + p \frac{dq}{dp} - \frac{dC}{dq} \cdot \frac{dq}{dp} = 0$$

**Explanation**: We use the product rule for $p \cdot q(p)$ and the chain rule for $C(q(p))$.

### Step 2: Rearrange the First-Order Condition

$$q + p \frac{dq}{dp} = \frac{dC}{dq} \cdot \frac{dq}{dp}$$

Since $\frac{dC}{dq} = MC$ (marginal cost):
$$q + p \frac{dq}{dp} = MC \cdot \frac{dq}{dp}$$

### Step 3: Solve for Price

Divide both sides by $\frac{dq}{dp}$ (assuming $\frac{dq}{dp} \neq 0$):
$$\frac{q}{\frac{dq}{dp}} + p = MC$$

$$p = MC - \frac{q}{\frac{dq}{dp}}$$

### Step 4: Express in Terms of Elasticity

The price elasticity of demand is:
$$\eta = \frac{dq}{dp} \cdot \frac{p}{q}$$

Therefore: $\frac{dq}{dp} = \eta \cdot \frac{q}{p}$

Substituting:
$$p = MC - \frac{q}{\eta \cdot \frac{q}{p}} = MC - \frac{p}{\eta}$$

### Step 5: Solve for Markup

Rearranging:
$$p = MC - \frac{p}{\eta}$$
$$p + \frac{p}{\eta} = MC$$
$$p\left(1 + \frac{1}{\eta}\right) = MC$$
$$p = \frac{MC}{1 + \frac{1}{\eta}}$$

Therefore, the markup is:
$$\mu = \frac{p}{MC} = \frac{1}{1 + \frac{1}{\eta}}$$

### Step 6: Final Form

Since demand is typically downward sloping, $\eta < 0$. Using $|\eta|$ for the absolute value:
$$\mu = \frac{1}{1 - \frac{1}{|\eta|}} = \frac{|\eta|}{|\eta| - 1}$$

Therefore:
$$\boxed{\mu = \frac{|\eta|}{|\eta| - 1}}$$

**Note**: Profit maximization requires $|\eta| > 1$ (elastic portion of demand curve).

## Section 1.4: Markups, Factor Cost Shares, and Output Elasticities

### Step 1: Cost Minimization Condition

For a flexible factor $X$, cost minimization yields:
$$p_X = \lambda f_X(\cdot)$$

where $p_X$ is the factor price, $\lambda$ is the Lagrange multiplier (marginal cost), and $f_X$ is the marginal product of $X$.

### Step 2: Manipulation to Get Markup

Rearrange the cost minimization condition:
$$\frac{p_X}{\lambda} = f_X$$

Since $\lambda = MC$ and we want the markup $\mu = \frac{P}{MC}$, we need to relate this to observable quantities.

### Step 3: Multiply by Convenient Terms

Multiply both sides by $\frac{X}{Q}$ where $Q$ is output:
$$\frac{p_X}{\lambda} \cdot \frac{X}{Q} = f_X \cdot \frac{X}{Q}$$

### Step 4: Recognize Economic Meanings

The right side is the output elasticity of factor $X$:
$\theta_X = f_X \cdot \frac{X}{Q} = \frac{\partial Q}{\partial X} \cdot \frac{X}{Q}$

For the left side, we need to relate it to the markup. Starting from:
$\frac{p_X}{\lambda} \cdot \frac{X}{Q} = \theta_X$

Since $\lambda = MC$ (marginal cost), we can write:
$\frac{p_X X}{MC \cdot Q} = \theta_X$

Now multiply both numerator and denominator of the left side by $P$ (price):
$\frac{p_X X \cdot P}{MC \cdot Q \cdot P} = \frac{p_X X}{P \cdot Q} \cdot \frac{P}{MC} = \frac{p_X X}{R} \cdot \mu$

where $R = P \cdot Q$ is revenue and $\mu = \frac{P}{MC}$ is the markup.

### Step 5: Final Result

From our manipulation:
$\frac{p_X X}{R} \cdot \mu = \theta_X$

Since $s_X = \frac{p_X X}{R}$ (factor's revenue share), we have:
$s_X \cdot \mu = \theta_X$

### Step 6: Final Result

Therefore:
$$s_X \cdot \mu = \theta_X$$

Solving for markup:
$$\boxed{\mu = \frac{\theta_X}{s_X}}$$

**Explanation**: The markup equals the ratio of the factor's output elasticity to its revenue share.

## Section 1.4.1: Why Revenue Elasticity Doesn't Work

### Step 1: Define Revenue Elasticity

Revenue elasticity of factor $X$:
$$\rho_X = \frac{d[p(q)q]}{dX} \cdot \frac{X}{pq}$$

### Step 2: Apply Chain Rule

$$\rho_X = \frac{d[p(q)q]}{dq} \cdot \frac{dq}{dX} \cdot \frac{X}{pq}$$

### Step 3: Compute Revenue Derivative

$$\frac{d[p(q)q]}{dq} = p + q\frac{dp}{dq}$$

Using the demand elasticity $\eta = \frac{dq}{dp} \cdot \frac{p}{q}$, we get $\frac{dp}{dq} = \frac{1}{\eta} \cdot \frac{p}{q}$:

$$\frac{d[p(q)q]}{dq} = p + q \cdot \frac{1}{\eta} \cdot \frac{p}{q} = p + \frac{p}{\eta} = p\left(1 + \frac{1}{\eta}\right)$$

### Step 4: Substitute Back

$$\rho_X = p\left(1 + \frac{1}{\eta}\right) \cdot \frac{dq}{dX} \cdot \frac{X}{pq} = \left(1 + \frac{1}{\eta}\right) \cdot \frac{dq}{dX} \cdot \frac{X}{q}$$

Since $\theta_X = \frac{dq}{dX} \cdot \frac{X}{q}$:
$$\rho_X = \left(1 + \frac{1}{\eta}\right) \theta_X$$

### Step 5: The Problem

From the markup-elasticity relationship: $\mu = \frac{|\eta|}{|\eta| - 1}$

This means: $1 + \frac{1}{\eta} = \frac{1}{\mu}$ (for $\eta < 0$)

Therefore: $\rho_X = \frac{\theta_X}{\mu}$

### Step 6: Revenue Elasticity Ratio

$$\frac{\rho_X}{s_X} = \frac{\theta_X/\mu}{s_X} = \frac{\theta_X}{\mu \cdot s_X}$$

But from our markup formula: $\mu = \frac{\theta_X}{s_X}$, so $\theta_X = \mu \cdot s_X$

Therefore:
$$\frac{\rho_X}{s_X} = \frac{\mu \cdot s_X}{\mu \cdot s_X} = 1$$

**Conclusion**: The revenue elasticity approach always gives a ratio of 1, regardless of the true markup value, making it uninformative.

## Section 1.5: Markdowns

### Step 1: Monopsony Profit Maximization

A firm with monopsony power faces upward-sloping factor supply $G(X)$ and maximizes:
$$\pi = R(q) - G(X) \cdot X$$

subject to $q = f(X)$.

### Step 2: First-Order Condition

$$\frac{d\pi}{dX} = R'(q) \cdot f'(X) - G'(X) \cdot X - G(X) = 0$$

**Explanation**: We use the chain rule for $R(q)$ and the product rule for $G(X) \cdot X$.

Rearranging:
$$R'(q) \cdot f'(X) = G'(X) \cdot X + G(X)$$

### Step 3: Factor Out and Express as Ratio

$$\frac{R'(q) \cdot f'(X)}{G(X)} = \frac{G'(X) \cdot X}{G(X)} + 1$$

### Step 4: Recognize Supply Elasticity

The elasticity of factor supply is:
$$\varepsilon_X = \frac{G'(X) \cdot X}{G(X)}$$

Therefore:
$$\boxed{\frac{R'(q) \cdot f'(X)}{G(X)} = \varepsilon_X + 1}$$

**Interpretation**: The ratio of marginal revenue product to factor price equals one plus the inverse of supply elasticity. When $\varepsilon_X > 0$ (upward-sloping supply), the factor is paid less than its marginal revenue product.

### Step 5: Combined Product and Factor Market Power

When the firm has both product market power (markup $\mu$) and factor market power, we can show:
$$(\varepsilon_X^{-1} + 1) \cdot \mu = \frac{\theta_X}{s_X}$$

**Explanation**: This combines the effects of both types of market power. The left side captures the factor market power effect (markdown) and product market power effect (markup), while the right side is the ratio of output elasticity to revenue share.

# Section 1.5 Markdowns - Step-by-Step Derivations

## Setup and Assumptions

We have a producer with monopsony power in the market for input $X$. The key elements are:

- **Inverse residual input supply curve**: $G(X)$, where $G$ is the price paid per unit of $X$
- **Total expenditures on input**: $G(X) \cdot X$  
- **Revenue function**: $R(q)$
- **Production function**: $q = f(X)$
- **Elasticity of residual factor supply**: $\varepsilon_X$

## Derivation 1: Basic Markdown Relationship

### Step 1: Set up the profit maximization problem
The producer maximizes profit = Revenue - Input costs:
$$\pi = R(f(X)) - G(X) \cdot X$$

### Step 2: Take the first-order condition
Taking the derivative with respect to $X$ and setting equal to zero:
$$\frac{d\pi}{dX} = R'(q) \cdot f'(X) - [G'(X) \cdot X + G(X)] = 0$$

**Explanation**: 
- $R'(q) \cdot f'(X)$ is the marginal revenue product of input $X$
- $G'(X) \cdot X + G(X)$ is the marginal cost of hiring input $X$ (not just the wage $G(X)$, but includes the effect of higher wages on all existing workers)

### Step 3: Rearrange to get the basic relationship
$$R'(q) \cdot f'(X) = G'(X) \cdot X + G(X)$$

### Step 4: Factor out $G(X)$ on the right side
$$R'(q) \cdot f'(X) = G(X) \cdot \left[\frac{G'(X) \cdot X}{G(X)} + 1\right]$$

### Step 5: Recognize the elasticity definition
The elasticity of residual factor supply is:
$$\varepsilon_X = \frac{dG/dX \cdot X}{G} = \frac{G'(X) \cdot X}{G(X)}$$

However, since the final expression in the document shows $\varepsilon_X^{-1}$, we interpret this as:
$$\frac{G'(X) \cdot X}{G(X)} = \varepsilon_X^{-1}$$

### Step 6: Substitute the inverse elasticity
$$R'(q) \cdot f'(X) = G(X) \cdot [\varepsilon_X^{-1} + 1]$$

### Step 7: Divide both sides by $G(X)$
$$\frac{R'(q) \cdot f'(X)}{G(X)} = \varepsilon_X^{-1} + 1$$

**Final Result 1**: The ratio of marginal revenue product to input price equals one plus the inverse elasticity of factor supply:
$$\frac{R'(q) \cdot f'(X)}{G(X)} = \varepsilon_X^{-1} + 1$$

## Derivation 2: Combined Product and Factor Market Power

### Step 1: Start with the basic markdown relationship
$$R'(q) \cdot f'(X) = G(X) \cdot (\varepsilon_X^{-1} + 1)$$

### Step 2: Define key variables
- **Output elasticity**: $\theta_X = \frac{\partial f}{\partial X} \cdot \frac{X}{f} = \frac{f'(X) \cdot X}{q}$
- **Revenue share**: $s_X = \frac{G(X) \cdot X}{R(q)}$

### Step 3: Express $f'(X)$ in terms of $\theta_X$
From $\theta_X = \frac{f'(X) \cdot X}{q}$, we get:
$$f'(X) = \frac{\theta_X \cdot q}{X}$$

### Step 4: Substitute into the markdown relationship
$$R'(q) \cdot \frac{\theta_X \cdot q}{X} = G(X) \cdot (\varepsilon_X^{-1} + 1)$$

### Step 5: Multiply both sides by $X$ and divide by $R(q)$
$$\frac{R'(q) \cdot \theta_X \cdot q}{R(q)} = \frac{G(X) \cdot X \cdot (\varepsilon_X^{-1} + 1)}{R(q)}$$

### Step 6: Simplify using definitions
Since $R(q) = P \cdot q$ and $s_X = \frac{G(X) \cdot X}{R(q)}$:
$$\frac{R'(q) \cdot \theta_X}{P} = s_X \cdot (\varepsilon_X^{-1} + 1)$$

### Step 7: Introduce the product market relationship
From the product market markup derivation, we know:
$$\frac{R'(q)}{P} = 1 - \frac{1}{|\eta|}$$
where $\eta$ is the price elasticity of demand.

We need to establish what $\frac{R'(q)}{P}$ equals. Let's derive this step by step.

* Sub-step 7a: Start with the revenue function
$R(q) = P(q) \cdot q$

* Sub-step 7b: Take the derivative with respect to quantity
$R'(q) = \frac{dR}{dq} = \frac{d[P(q) \cdot q]}{dq} = P'(q) \cdot q + P(q)$

* Sub-step 7c: Factor out $P(q)$
$R'(q) = P(q) \cdot \left[\frac{P'(q) \cdot q}{P(q)} + 1\right]$

* Sub-step 7d: Recognize the price elasticity of demand
The price elasticity of demand is:
$\eta = \frac{dq}{dP} \cdot \frac{P}{q}$
By the inverse function theorem:
$\frac{dP}{dq} = \frac{1}{\frac{dq}{dP}} = \frac{1}{\eta \cdot \frac{q}{P}} = \frac{P}{\eta \cdot q}$
Therefore: $P'(q) = \frac{P}{\eta \cdot q}$

* Sub-step 7e: Substitute back into the expression
$\frac{P'(q) \cdot q}{P(q)} = \frac{\frac{P}{\eta \cdot q} \cdot q}{P} = \frac{1}{\eta}$

* Sub-step 7f: Complete the marginal revenue expression
$R'(q) = P(q) \cdot \left[\frac{1}{\eta} + 1\right] = P(q) \cdot \left[\frac{1 + \eta}{\eta}\right]$

* Sub-step 7g: Divide by $P$ to get the ratio
$\frac{R'(q)}{P} = \frac{1 + \eta}{\eta} = 1 + \frac{1}{\eta}$

* Sub-step 7h: Account for the sign of elasticity
Since demand curves slope downward, $\eta < 0$. Using $|\eta|$ for the absolute value:
$\frac{R'(q)}{P} = 1 + \frac{1}{\eta} = 1 - \frac{1}{|\eta|}$

Final result for Step 7:
$\frac{R'(q)}{P} = 1 - \frac{1}{|\eta|}$
where $\eta$ is the price elasticity of demand.

### Step 8: Substitute the product market relationship
$$\left(1 - \frac{1}{|\eta|}\right) \cdot \theta_X = s_X \cdot (\varepsilon_X^{-1} + 1)$$

### Step 9: Rearrange to isolate the markup-markdown relationship
$$\frac{\theta_X}{s_X} \cdot \left(1 - \frac{1}{|\eta|}\right) = \varepsilon_X^{-1} + 1$$

### Step 10: Express in terms of markup
Since the markup $\mu = \frac{|\eta|}{|\eta| - 1}$, we have:
$$1 - \frac{1}{|\eta|} = \frac{|\eta| - 1}{|\eta|} = \frac{1}{\mu}$$

Therefore:
$$\frac{\theta_X}{s_X} \cdot \frac{1}{\mu} = \varepsilon_X^{-1} + 1$$

**Final Result 2**: 
$$(\varepsilon_X^{-1} + 1) \cdot \mu = \frac{\theta_X}{s_X}$$

## Economic Interpretation

- **When $\varepsilon_X^{-1} = 0$** (perfect competition in factor market): The firm pays factors their marginal product, and we get the standard markup relationship $\mu = \frac{\theta_X}{s_X}$

- **When $\varepsilon_X^{-1} > 0$** (monopsony power): The firm pays factors less than their marginal product. The markdown creates an additional wedge between what inputs are paid and their contribution to production

- **The combination**: Both product market power (markup $\mu$) and factor market power (markdown via $\varepsilon_X^{-1}$) together determine the relationship between input productivity and input compensation

The key insight is that when firms have both product and factor market power, the simple markup formula $\mu = \frac{\theta_X}{s_X}$ no longer holds. Instead, the product of the markup and the markdown factor $(1 + \varepsilon_X^{-1})$ equals the ratio of output elasticity to cost share.

### 2025-06-18 posts/ethics/KenkyuRinri2025_slides.qmd
added DelabastitaRubens repo and ethics repo

---
title: |
  | 研究倫理
subtitle: |
  | 2025年6月19日(木)
  | 開発研究センター
  | ミクロ経済分析研究グループ
author: "伊藤成朗"
engine: knitr
format:
  revealjs: 
    standalone: true
    embed-resources: true
    slide-number: true
    width: 1600
    height: 900
    preview-links: auto
    logo: images/quarto.png
    footer: <http://seiroito.github.io/QuartoFiles/Horton/Horton2025_slides.html>
    theme: sky
lightbox: auto
include-in-header:
  - text: |
      <style>
      #title-slide
       .title {
        font-size: 2.0em;
      }
      </style>
bibliography: ../seiro.bib
#### If it is 2 levels up, ../../seiro.bib
resources:
  - KenkyuRinri2025_slides.pdf
#### quarto render KenkyuRinri2025_slides.qmd
---
```{r xaringan-panelset, echo=FALSE, warning = F}
library(xaringanExtra)
use_panelset()
library(RefManageR)
```
```{r setting, eval = T, cache = T, warning = F, message = F, echo = F}
library(RefManageR)
myBib <- ReadBib("c:/seiro/settings/TeX/seiro.bib")
BibOptions(max.names = 3)
```
```{r flair_color, eval = F, echo=FALSE}
library(flair)
red <- "#FF0000"
green <- "#00FF00"
blue <- "#0000FF"
```
```{css echo=FALSE}
.red {color: #FF0000;}
Green {color: #00FF00;}
.blue {color: #0000FF;}
White {
  color: white;
}
LightBlue {
  color: #33E7FF;
}
DarkBlue {
  color: #010638;
}
DBlue {
  color: dodgerblue;
}
Red {
  color: red;
}
Orange {
  color: #ff8442;
}
Yellow {
  color: yellow;
}
#### @font-face {
####   font-family: "hiragino";
####   #### src: url(../../../../../settings/fonts/HiraginoMinchoProW3.otf");
####   src: url("c:/seiro/settings/fonts/HiraginoMinchoProW3.otf");
#### }
@font-face {
  font-family: "Roboto";
  src: url("c:/seiro/settings/fonts/Roboto-Regular.ttf");
}
#### h1 {
####   font-size: 42pt;
####   font-family: 'hiragino';
####   color: black;
####   border-bottom: 1px;
####   border-color: rgb(248, 250, 252);
####   padding-bottom: 0pt;
#### }
p {
  #### font-size: 32pt;
  font-family: 'Roboto';
  color:white
}

.highlight-last-item > ul > li,
.highlight-last-item > ol > li {
  opacity: 0.5;
}
.highlight-last-item > ul > li:last-of-type,
.highlight-last-item > ol > li:last-of-type {
  opacity: 1;
}
.inverse {
  background-color: #00003c;
  color: #d6d6d6;
  text-shadow: 0 0 20px #333;
}
mark.red {
    color:#ff0000;
    background: none;
}
mark.blue {
    color:#0000A0;
    background: none;
}
.my-style {
  font-weight: bold;
  font-style: italic;
  font-size: 1.5em;
  color: red;
}
.large {
  font-size: 2em;
}
blockquote {
  border-left: .2px solid #275d38;
  margin: -5px 80px -5px 20px;
  padding-top: -0.5px;
  padding-bottom: -0.5px;
  line-height: 1.35em;
}
.pull-left60 {
  float: left;
  width: 57%;
  padding-right: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-right40 {
  float: right;
  width: 37%;
  padding-left: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-left70 {
  float: left;
  width: 67%;
  padding-right: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-right30 {
  float: right;
  width: 29%;
  padding-left: -4px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-left75 {
  float: left;
  width: 73%;
  padding-right: 2px
  margin-top: -5em;
  margin-bottom: 0em;
}
.pull-right25 {
  float: right;
  width: 22%;
  padding-left: 2px
  margin-top: -5em;
  margin-bottom: 0em;
}
.pull-left50 {
  float: left;
  width: 50%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: 0em;
}
.pull-right50 {
  float: right;
  width: 50%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: 0em;
}
.pull-right50jaxa {
  float: right;
  width: 50%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: -5em;
}
.pull-left20 {
  float: left;
  width: 18%;
  padding-right: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-right80 {
  float: right;
  width: 79%;
  padding-left: -4px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-left100 {
  float: left;
  width: 100%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: 0em;
}
.pull-right {
  float: right;
  width: 100%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: 0em;
}
.center2 {
  margin: 0;
  position: absolute;
  top: 50%;
  left: 50%;
  -ms-transform: translate(-50%, -50%);
  transform: translate(-50%, -50%);
}
```

::: {.center2}
[https://seiroito.github.io/Miscellaneous/ethics/KenkyuRinri2025_slides.html](https://seiroito.github.io/Miscellaneous/ethics/KenkyuRinri2025_slides.html)
:::

----

# 研究倫理{background-color="#010638"}

----

## {background-color="#010638"}

::: {style="font-size: 140%;"}

:::: {.fragment}
仕事は社会・業界の行動規範の範囲内で実施します
::::

<br>

:::: {.fragment}
法令、組織の規則、社会通念、業界通念
::::

<br>

:::: {.fragment}
研究業界(学界+その周辺)にも特有の規範=研究倫理があります
::::

<br>

:::: {.fragment}
本講義の目的: 研究コミュニティに身を置く者が研究倫理逸脱で失敗しないようにすること
::::

<br>

:::: {.fragment}
崇高な研究哲学論ではなく、実際的な自衛の話
::::
:::

----

# 研究倫理 = {狭義の研究倫理,  研究公正} {background-color="#010638"}

## {background-color="#010638"}

::: {style="font-size: 120%;"}
極言すると

<br>

::::{.fragment}
研究倫理とは  
::::

<br>

::: {style="font-size: 150%;"}
::::{.fragment}
研究で対象社会に<Red>迷惑をかけない</Red>という原則  
::::
:::

<br>

::::{.fragment}
 研究公正とは  
::::

<br>

::: {style="font-size: 150%;"}
::::{.fragment}
研究で<Red>ごまかしをしない</Red>という原則  
::::
:::

<br>

::::{.fragment}
対象社会は日本や学界も含みます
::::
:::

----

## {background-color="#010638"}

::: {style="font-size: 140%;"}
研究は社会と契約を結んで実施しています

<br>

:::: {.fragment}
よって、これらは最低限の原則です  
::::

<br>

::::{.fragment}
それでも足をすくわれる研究者はいます
::::

<br>

::::{.fragment}
どう判断すれば分からなくなったとき、大原則に戻ると考えやすいです  
::::
:::

<br>

::: {style="font-size: 180%;"}
::::{.fragment}
<Red>社会から託された信任を裏切らない</Red>
::::
:::

----

## {background-color="#010638"}

<!--::: {style="text-align: center;"}
![](REAndRI.jpg){height="90%"}
:::-->

:::::::::: {.columns}
::::: {.column width="20%"}
::::{.fragment}
* Do good for the society (not just for oneself/clients).  
::::
:::::

::::: {.column width="60%"}
::: {style="text-align: center;"}
![](REAndRI.jpg){width="85%"}
:::
:::::

::::: {.column width="20%"}
::::{.fragment}
* Be honest to the scientific standard (not just to oneself).  
::::
:::::

::::::::::

::::{.fragment}
ごまかしのある研究は社会に迷惑をかけるように、研究公正と研究倫理はきれいに分かれず、相互に関連する部分があります
::::

::::{.fragment}
[研究公正の鳥瞰図研究](https://ukrio.org/news/publication-of-research-integrity-a-landscape-study/) が示す研究公正のキーワード: honest, rigorous, transparent, open, respect and stewardship of research, care for the people of research.
::::

----

## {background-color="#010638"  background-image="./JAXASumimasen.jpg" background-size="100%"}

::: {.center2}
::: {style="font-size:160%; color:#33E7FF;"}
<span style="color:LightBlue;">研究公正 research integrity</span> 
:::
:::

----

## 研究公正 research integrity {background-color="DarkBlue"  background-image="./JAXASumimasen.jpg" background-size="100%"}

----

# 研究公正 research integrity {background-color="DarkBlue"  background-image="./JAXASumimasen.jpg" background-size="100%"}

----

## 研究公正 research integrity {background-color="black" background-image="./JAXASumimasen.jpg" background-size="100px" background-repeat="repeat"}

----

## {background-color="#010638"}

:::: {.absolute top="45%" left="0%" width="1600"}
::: {style="text-align: center;"}
<font color="white">
写真はJAXA記者会見映像からキャプチャしました
</font>
:::
::::

# 科学的根拠をもとに研究し、得た知識を歪めずに発表すること。{background-color="#010638"}

# 研究公正は、やってはいけないことで定義されます。{background-color="#010638"}

## <Red>一発退場 Research misconducts</Red> (<LightBlue>FFP</LightBlue>, [厚労省研究不正ガイドライン「特定不正行為」](https://www.mhlw.go.jp/file/06-Seisakujouhou-10600000-Daijinkanboukouseikagakuka/0000152685.pdf#page=8)){background-color="#010638"}

:::::::::: {.columns}
::::: {.column width = "65%"}
::: {.fragment}
::: {style="font-size:130%"}
<White> <LightBlue>F</LightBlue><Orange>abrication ねつ造</Orange>「存在しないデータや研究成果を作成すること。」  
 <LightBlue>F</LightBlue><Orange>alsification 改ざん</Orange>「操作」によって「データ」や「成果などを真正でないものに加工すること」。  
 <LightBlue>P</LightBlue><Orange>lagerism 剽窃、盗用</Orange> 他の研究者の「アイディア」「方法」「データ」「結果」を「了解または適切な表示なく流用すること」。</White>
:::
:::
:::::

::::: {.column width = "35%"}
::: {style="font-size:90%"}
::: {.incremental}
* やると研究者人生が終わります。  
* 理由: 研究者社会は相互信頼で成立しています。ウソをつかないことを前提に評価がされて、知識が共有されます。裏切ると、研究者社会全体が資源を浪費するので、厳罰処分されます。不正が露見するまで、他の研究者は存在しない現象を追求して苦しみます。  
   * STAP論文: 追試資源や継続研究資源の浪費、周辺研究まで疑いの目
:::
:::
:::::
::::::::::

----

## <Yellow>信用失墜</Yellow> <Orange>Questionable research practices (QRPs)</Orange>{background-color="#010638"}

:::::::::: {.columns}
::::: {.column width = "55%"}
::: {.incremental}
::: {style="font-size:130%"}
* データの未保存や再現性の欠如  
* 不適切な統計手法  
* ギフト・オーサーシップ ([CRediT](https://www.elsevier.com/authors/policies-and-guidelines/credit-author-statement))  
* 論文の多重投稿  
* 利益相反によるバイアス  
:::
:::
:::::

::::: {.column width = "45%"}
::: {.incremental}
* 識者によって内容が異なるので幾つか列記  
* $P$-hacking: E.g.,  望んだ結果が出るまで推計を試行し、その結果のみ提示すること  
* GAは贈賄双方に便益あり、[なくなりにくい、でも、著者責任も付いてくる](https://www.bmj.com/content/309/6967/1456#:~:text=Perhaps%20this%20scandal%20will%20finally%20undermine%20gift%20authorship.%20At%20the%20very%20least%20it%20should%20make%20researchers%20think%20hard%20about%20the%20responsibilities%20that%20come%20with%20putting%20their%20names%20on%20papers.)   
* 刊行された論文を翻訳して投稿してはいけません ([International Committee of Medical Journal Editors](https://www.icmje.org/recommendations/browse/publishing-and-editorial-issues/overlapping-publications.html#:~:text=Authors%20should%20not%20submit%20the%20same%20manuscript%2C%20in%20the%20same%20or%20different%20languages%2C%20simultaneously%20to%20more%20than%20one%20journal.))、翻訳であることを明示しても危ういので、翻訳論文はワーキング・ペーパーに留めましょう  
:::
:::::
::::::::::

See a list of QRPs used in Dutch national survey (DNSRI) summarised [here](https://htmlpreview.github.io/?https://github.com/SeiroIto/PaperNotes/blob/main/QRPsInDutchNationalSurvey_Tufte.html)

----

## {background-color="#010638"}

### 蔓延度は不明ですが、結構深刻では?

<White>
<Orange>例1:</Orange> Dutch National Survey on Research Integrity [`r Citep(myBib, "Gopalakrishnaetal2022")`](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0263023&fbclid=IwAR17MW5qz2wpZXQz6crAQVrsDbGsx0DKeE5ZIuvIywF137rYlf_5hXz1Alw):

::: {style="font-size:80%"}
* 15大学+7医学研究所、大学院生以上、`r formatC(63778-1913, format = "f", digits = 0)`人メール送達、6813人回答(`r round(100*6813/(63778-1913), 0)`%)、過去3年  
* 回答者内訳: 人文科学(全体の`r round(100*636/6813, 0)`%)、社会・行動科学(同`r round(100*1965/6813, 0)`%)、生命・医学科学(同`r round(100*2747/6813, 0)`%)、自然・工学科学(同`r round(100*1465/6813, 0)`%)  
* 大学院生(全体の`r round(100*2013/6813, 0)`%)でも準教授以上(同`r round(100*2066/6813, 0)`%)でも回答内容の傾向はほぼ同じ  
</White>
:::

:::::::::: {.columns}
::::: {.column width = "55%"}
<White>
FF(P) 8.3% engaged in any FF(P).  
</White>

<White>
QRP 51.3% engaged frequently in at least 1 QRP.  
</White>
:::::

::::: {.column width = "45%"}
<White>
Beh&Soc.Sci 5.7%  
</White>

<White>
Beh&Soc.Sci 50.2%
</White>
:::::
::::::::::

----

## {background-color="#010638"}

### 蔓延度は不明ですが、結構深刻では?{background-color="#010638"}

<White>
<Orange>例2:</Orange> The International Research Integrity Survey `r Citep(myBib, "Allumetal2022")`:

::: {style="font-size:80%"}
* Web of Scienceのジャーナル21894誌・本、2016-2020年、816万 &rarr; 392万 &rarr; 375万 &rarr; 307万2372 (対象国: 米欧)  
* Stratified random sampling (strata=natural, medical, social sciences and humanities)  
* (おそらく)307万にemail送付  &rarr; 73757が回答 (ウェイト後回答率=7.2%)  &rarr; 64074 (分析標本サイズ)  
* 過去3年のQRP([IRISSupplementaryTable.xlsx](https://osf.io/s5w9q), [download](https://osf.io/download/s5w9q/))  
</White>
:::

:::::::::: {.columns}
::::: {.column width = "70%"}
:::: {.fragment}
::: {style="font-size:100%"}
<White>
ギフト・オーサーシップ: 69.0%(欧)、55.1%(米)  
</White>

<White>
予想と矛盾する結果の秘匿: 25.8%(欧)、20.3%(米)  
</White>

<White>
信念と矛盾する先行研究の無視: 20.5%(欧)、15.3%(米)  
</White>

<White>
不完全な査読をする: 53.8%(欧)、49.7%(米)  
</White>

<White>
剽窃: 6.4%(欧)、6.5%(米)  
</White>
:::
::::
:::::

```{r IRIS data}
#### "C:/seiro/docs/personal/Miscelleneous/QuartoFiles/posts/ethics"
library(data.table)
ir <- fread("./IRIS_QRP.prn", sep = "\t")
irL <- melt(ir, idvars = 1:2, measure.vars = 3:ncol(ir))
irLW <- reshape(irL, direction = "wide", idvar = c("region", "variable"), 
  timevar = "unit", 
  v.names = "value")
irLW[, group := "field"]
irLW[grepl("Aca|Oth", variable), group := "institution"]
irLW[grepl("	Per|Tem", variable), group := "tenure"]
irLW[grepl("Ea|Mid|La", variable), group := "career period"]
irLW[grepl("Ma|Fe", variable), group := "gender"]
irLW[, group := factor(group, levels = c("field", "institution", "tenure", "career period", "gender"))]
library(ggplot2)
ThisTheme <- theme(
    strip.text.x = element_text(color = "blue", size = 10, 
      margin = margin(0, 1.25, 0, 1.25, "cm")), 
    strip.text.y = element_text(color = "blue", size = 6, 
      margin = margin(t=1.65, r=.1, b=1.65, l=.1, "cm")),
    axis.text.x = element_text(size = 10, angle = 25, vjust = 1, hjust = 1), 
    axis.text.y = element_text(size = 6), 
  legend.position = "bottom",
  legend.title = element_blank()
)
g <- ggplot(data = irLW, aes(x=variable, y=value.mean, 
    group = region, color = region, fill = region, shape = region)) +
  geom_pointrange(aes(ymin = value.lb, ymax = value.ub), 
    position = position_dodge(width=.1))+ 
  facet_grid(cols = vars(group), scales = "free", space = "free") +
  xlab("")+
  ylab("% of respondents\ncommiting at least 1 QRP\nin last 3 years (asked 2021)")+
  ThisTheme
ggsave(
    "./IRIS_QRP.jpg"
  , g, 
  width = 10*2, height = 5*2, units = "cm", dpi = 300)
```

::::: {.column width = "30%"}
:::: {.fragment}
::: {#fig-QRP layout-ncol=1}
![QRP率: グループ別の内訳](./IRIS_QRP.jpg){width="60%"}
:::
::::
:::::
::::::::::

----

## {background-color="#010638"}

### 蔓延度は不明ですが、結構深刻では?

<Orange>例3:</Orange> [Data Collada (psychologist blog)](https://datacolada.org/) vs. [Francesca Gino (HBS prof)](https://www.francesca-v-harvard.org/home), 2021-

::: {style="font-size:80%"}
* [Allegation](https://datacolada.org/109): Data tampering in several of her work on dishonesty (incl. with Dan Ariely!)  
* HBS: Put Gino on an unpaid admin leave, hired file forensic investigators, sent retraction requests to journals  
* Gino: Sued these folks and Harvard, asking for "at least $25 million" ([para 24, main document](https://www.courtlistener.com/docket/67659904/1/gino-v-president-and-fellows-of-harvard-college/))  
* [The problem, though, is that it will take years --- and be extraordinarily expensive --- to settle](https://www.vox.com/future-perfect/2023/8/9/23825966/francesca-gino-honesty-research-scientific-fraud-defamation-harvard-university#:~:text=The%20problem%2C%20though%2C%20is%20that%20it%20will%20take%20years%20%E2%80%94%20and%20be%20extraordinarily%20expensive%20%E2%80%94%20to%20settle)
* [150人あまりの共著者たちはどの論文は不正がないか調べる羽目に](https://www.science.org/content/article/after-honesty-researcher-s-retractions-colleagues-expand-scrutiny-her-work#:~:text=This%20week%2C%20Simonsohn%20and%20five%20other%20Gino%20co%2Dauthors%20emailed)  
* [ハーバード: Ginoを解雇 (2025年5月25日)](https://www.wgbh.org/news/education-news/2025-05-25/in-extremely-rare-move-harvard-revokes-tenure-and-cuts-ties-with-star-business-professor)
:::

<Orange>例4:</Orange> [国立医薬品食品衛生研究所職員による研究活動上の不正行為及びISO/IEC 17025認定審査対応における不適切行為に関する調査結果の公表](https://www.nihs.go.jp/oshirasejoho/Press-Release/20231226/index.html)

::: {style="font-size:80%"}
* 2023年12月26日、食品衛生部長によるFabrication, Falsificationを研究所が認定
:::

----

## なぜFFPをするのか{background-color="#010638"}

Gino: HBS給与は1.4億円+印税+講演料、ばれなければ大金

* 日本の文系(<2000万円)だとFFPは割に合いません  
* 国立医薬品食品衛生研究所所長で1900万円くらい

結構ばれる: Ginoは追試、食品衛生部長は内部通報

<br>

貧乏くじ

<br>

オッズをよく知らないまま、自ら進んで引くのは賢くない

> The university’s top governing board, the Harvard Corporation, decided this month to revoke Francesca Gino’s tenure and end her employment at Harvard Business School. ([WGBH news](https://www.wgbh.org/news/education-news/2025-05-25/in-extremely-rare-move-harvard-revokes-tenure-and-cuts-ties-with-star-business-professor))

----

## 自分は本当に大丈夫か?{background-color="#010638"}

:::::::::: {.columns}
::::: {.column width = "40%"}
<White>アジ研: 外部から研究者成果に剽窃を指摘される(外部委員[2018年](https://www.ide.go.jp/Japanese/New/2018/20180124.html)、内部委員[2024年](https://www.mext.go.jp/a_menu/jinzai/fusei/1360847_00045.html))  </White>
:::::

::::: {.column width = "60%"}
:::: {.fragment}
* 対応: 調査委員会設置、事実認定、事実公表、処分、(科研費)文科省への事実・原因・処分の報告
::::
::::: 
::::::::::

:::::::::: {.columns}
::::: {.column width = "40%"}
:::: {.fragment}
<White>公正な研究? QRP? <White>
::::
:::: {.fragment}
 <White>和文単行書の成果検討では...</White>
::::
:::::

::::: {.column width = "60%"}
:::: {.incremental}
 * 成果を冗長に書く  
 * 要約が不正確  
 * どのような過程を踏んだか明記しない  
 * 図表の脚注に留意事項を説明するが本文は無記載  
 * 研究なのにダークパタン  
 * 過去著作の数段落を明示せずコピー、ただし、参考文献には掲載  
::::
::::: 
::::::::::

:::: {.fragment}
<White>Githubで研究過程を公開=ピア・レビューを活用  </White>
::::

:::: {.fragment}
<White>[こんな感じ](https://github.com/SeiroIto)です</White>
::::

----

## なぜQRPをするのか{background-color="#010638"}

 <White>
成果を出したい、でも、今のままだと成果にならないと考えている
 </White>

:::: {.fragment}
 <White>
 内容が不十分
 </White>
::::

:::::::::: {.columns}
::::: {.column width = "50%"}
:::: {.incremental}
* 結果がはっきりしない  
* 2つの結果が矛盾する  
* 何を議論すべきかもよく分からない  
::::
:::::

::::: {.column width = "50%"}
:::: {.incremental}
* 不適切な統計手法  
* データの未保存や再現性の欠如  
* 曖昧な記述と結論...QRP?  
   * 結論は曖昧です、と明確に書けばok  
   * 単に記述を不明瞭にするのはごまかし=QRP  
::::
::::: 
::::::::::

:::: {.fragment}
* 過程(データとコード)を公開  
* Pre analysis planを公開・登録  
::::

----

## なぜQRPをするのか{background-color="#010638"}

:::::::::: {.columns}
::::: {.column width = "50%"}
:::: {.fragment}
 <White>
自信、発信力が無い
 </White>
::::
:::: {.incremental}
* ビッグ・ネームの推薦を利用したい  
::::

:::::

::::: {.column width = "50%"}
:::: {.incremental}
* ギフト・オーサーシップ  
* 反例: ネットワーク理論: Erdős-Rényiモデル. [Erdős](https://en.wikipedia.org/wiki/Paul_Erd%C5%91s)は1525論文掲載  
::::
::::: 
::::::::::

:::: {.fragment}
 <White>
時間がない
 </White>
::::
:::::::::: {.columns}
::::: {.column width = "50%"}
:::: {.incremental}
* 今までの研究に手を加えるだけ  
* 他者の成果・アイディアを取り込む
::::
:::::

::::: {.column width = "50%"}
:::: {.incremental}
* 多重投稿、多重出版  
   * 機械学習で簡単に判明  
* 剽窃
::::
::::: 
::::::::::

:::: {.fragment}
 <White>
利益相反
 </White>
::::
:::::::::: {.columns}
::::: {.column width = "50%"}
:::: {.incremental}
* スポンサーの利益と研究結果が一致
* 余計な詮索をされたくない
::::
:::::

::::: {.column width = "50%"}
:::: {.incremental}
* 利益相反があることを表明しない  
   * 簡単に判明  
* 研究が健全ならば、利益相反表明しない理由がない
::::
::::: 
::::::::::

----

##  大規模言語モデル(LLM)時代のリスク{background-color="#010638"}

<White>Fabrication, falsification、plagerismのリスク</White>

:::::::::: {.columns}
::::: {.column width = "50%"}
:::: {.fragment}
1. 文章出力をそのまま使ってはいけない
::::
:::::

::::: {.column width = "50%"}
:::: {.incremental}
* FFP
* Authorship
* [「Certainly! Here is a more detailed version of...」と本文に書いてある論文](https://pubpeer.com/publications/01593D32DC9AD348EA64D7BF9ABE0C)  
::::
:::::

::::: {.column width = "50%"}
:::: {.fragment}
2. 図表や推計は与えたデータのみから出力する
::::
:::::

::::: {.column width = "50%"}
:::: {.incremental}
* FF
* 自分でコードを書く、自分でexcelを操作する方が安心できます
::::
:::::

::::: {.column width = "50%"}
:::: {.fragment}
3. 与えた以外の外部情報を用いるかは必ず指定する
::::
:::::

::::: {.column width = "50%"}
:::: {.incremental}
* FFP
* [ICMJE推奨事項](https://www.icmje.org/news-and-editorials/icmje-recommendations_annotated_jan24.pdf#page=3)ではどのように使ったかを具体的に明示するように要求  
::::
::::: 
::::::::::

----

##  大規模言語モデル(LLM)時代のリスク{background-color="#010638"}

<White>作業委託を効率的にするための方針</White>

:::::::::: {.columns}
::::: {.column width = "50%"}
:::: {.fragment}
<White>真偽を確かめやすい単純労働を担わせる</White>
::::
:::::

::::: {.column width = "50%"}
:::: {.incremental}
* 「以下の出来事を時系列順に並べ替えて、日付と出来事の2列からなるtab separated text形式で出力して」  
* 「この論文の数式を全部抜き出して」
* 「変数と定義をmdの表にして出力して」
::::
:::::

::::: {.column width = "50%"}
:::: {.fragment}
<White>真偽を確かめにくいが、作業の一部を担わせ、自分がその内容を確認する</White>
::::
:::::

::::: {.column width = "50%"}
:::: {.incremental}
* 「この論文の内容をこういうふうに要約して」
* 「(3)式から(5)式までの導出過程を詳細に説明して」
* 「この文章をもう少し簡単に言い換えて」
::::
::::: 
::::::::::

----

## {background-color="#010638"  background-image="./JAXASumimasen.jpg" background-size="100%"}

::: {.center2}
::: {style="font-size:160%; color:#33E7FF;"}
<span style="color:LightBlue;">研究倫理 research ethics</span> 
:::
:::

# 研究対象(者の人権)保護、研究コミュニティの評判維持{background-color="#010638"}

# 研究倫理は、守るべきことで定義されます{background-color="#010638"}

# {background-color="#010638"}

:::::::::: {.columns}
::::: {.column width = "50%"}
:::{style="text-align: center;"}
![](the-constant-gardener-original.jpg){height="23cm"}
:::
:::::

::::: {.column width = "50%"}
:::{style="text-align: center;"}
![](TheExperiment.jpg){height="23cm"}
:::
::::: 
::::::::::

----

## {background-color="#010638"}

### 3つのレイヤー `r Citep(myBib, "Mustajoki2017", after = ", p.37")`

:::::::::: {.columns}
::::: {.column width = "30%"}
 1. 法的義務  
:::::

::::: {.column width = "70%"}
:::: {.fragment}
個人情報保護、人権・動物の権利保護など
::::
::::: 
::::::::::

:::::::::: {.columns}
::::: {.column width = "30%"}
:::: {.fragment}
2. ガイドライン  
::::
:::::

::::: {.column width = "70%"}
:::: {.fragment}
インフォームド・コンセント、研究の正確さ、貢献の正確さ、研究協力者の表示、など
::::
::::: 
::::::::::

:::::::::: {.columns}
::::: {.column width = "30%"}
:::: {.fragment}
3. 暗黙の了解  
   * ? (分野や組織ごとにあるルールなど、らしい)  
::::
:::::

::::: {.column width = "70%"}
:::: {.incremental}
* 論文への質問に著者は真摯に答える  
* 内容を批判しても著者の人格まで批判しない  
* 読む前後で自分の知識がどう変わったか書く  
* 研究をより良くするため、読者がより正確に理解するためにコメント・批判をする  
* 問題の指摘で終わらず結論にどう影響するかまで書く
::::
::::: 
::::::::::

:::: {.fragment}
最後の3つはMacartan Humphreysの[How to critique](https://macartan.github.io/teaching/how-to-critique)から抜粋  
::::

----

## {background-color="#010638"}

途上国研究における研究倫理8ポイント `r Citep(myBib, "Emanuel2004")`

:::::::::: {.columns}
::::: {.column width = "60%"}
:::: {.incremental}
1. <Orange>Collaborative partnership</Orange>	対象個人や社会が便益を受ける研究か。 
1. <Orange>Social value</Orange>	社会的に価値のある研究か。 
1. <Orange>Scientific validity</Orange>	<LightBlue>科学的裏付け</LightBlue>(知見を得る方法論)はあるか。 
1. <Orange>Fair subject selection</Orange>	 参加者は公平に選ばれたか。 
1. <Orange>Favorable risk-benefit ratio</Orange>	リスク相応の便益はあるか。個人リスク、社会便益。 
1. <Orange>Independent review</Orange>	<LightBlue>独立機関が審査</LightBlue>したか。 
1. <Orange>Informed consent</Orange>	<LightBlue>インフォームド・コンセント</LightBlue>を得たか。 
1. <Orange>Respects for recruited participants and study, communities</Orange> 個人や対象社会の<LightBlue>意思を尊重</LightBlue>し<LightBlue>経過・結果を伝達</LightBlue>したか。 
::::
:::::

::::: {.column width = "40%"}
:::: {.fragment}
殆どの社会科学研究(=観察研究)  
::::

:::: {.incremental}
* 便益は無く、リスクも個人情報漏洩以外は僅少、社会的価値もよく分からないので、参加者を公平に選ぶべき理由は倫理に照らしてあまりない。  
* 審査の独立性、インフォームド・コンセント、個人情報保護、意思の尊重は手続きに関する考慮。科学的裏付けは研究をする動機。 
* あらゆる調査・研究で満たすべき項目であり、研究倫理審査が特別なことをしているわけではない。 
::::
::::: 
::::::::::

----

## {background-color="#010638"}

::: {.panel-tabset}

## 研究倫理チェックリスト(あらすじ)

:::::::::: {.columns}
::::: {.column width = "15%"}
:::: {.fragment}
企画段階:  
::::
:::::

::::: {.column width = "85%"}
:::: {.incremental}
* やる科学的価値=社会的価値はあるか...既存研究と違いはあるか?  
* 対象選定、標本サイズ、質問票などの研究デザインは十分か=知見が得られる見込みがあるか?  
* 個人情報保護の手段は考えたか?  
::::
:::::

::::: {.column width = "15%"}
:::: {.fragment}
実施段階:  
::::
:::::

::::: {.column width = "85%"}
:::: {.incremental}
* 研究説明書(research information sheet): 研究の概要と連絡先を説明   
* 研究同意書(informed consent): 参加者の権利+何に同意したのか+署名   
   * いつでも参加拒否できる、いかなる質問にも回答拒否できるなど  
* 必要に応じて研究倫理審査委員会の推薦状、現地IRB (institutional review board)の推薦状・許可状  
::::
:::::

::::: {.column width = "15%"}
:::: {.fragment}
分析段階:  
::::
:::::

::::: {.column width = "85%"}
:::: {.incremental}
* ハードコピーは鍵付き書庫に保存したか  
* 分析用データは個人識別子と分離したか  
::::
::::: 
::::::::::

## アジ研の研究倫理審査

* おもに介入研究が対象: 本審査  
* 雑誌によっては観察研究でも要求する場合がある: 迅速審査  
* 判断に迷う場合は研究倫理審査委員会事務局(研究企画課)へ相談を  
* 研究倫理審査委員会は研究倫理に適合しながら研究実施できる方法を共に考える場所です。  
   * 例: COVID-19下で感染リスクを日常生活以上に高めずに調査する方法などを検討

:::

----

## 経営者が研究倫理審査や研究能力を軽視すると、配下の研究者が研究倫理・研究公正から逸脱するのを防げない{background-color="#010638"}

----

## {background-color="#010638"}

:::::::::: {.columns}
::::: {.column width = "70%"}
JAXAの研究倫理・研究公正違反 (2016-2017)  
:::::

::::: {.column width = "30%"}
![](JAXAlogo.jpg){height="50px"}![](JAXA_pic.jpg){height="50px"}
:::::

::::: {.column width = "30%"}
:::: {.incremental}
* 閉鎖環境負荷試験: 2週間×5組、計42人の研究対象者  ... [宇宙兄弟(3)](https://www.amazon.co.jp/%E5%AE%87%E5%AE%99%E5%85%84%E5%BC%9F%EF%BC%88%EF%BC%93%EF%BC%89-%E3%83%A2%E3%83%BC%E3%83%8B%E3%83%B3%E3%82%B0%E3%82%B3%E3%83%9F%E3%83%83%E3%82%AF%E3%82%B9-%E5%B0%8F%E5%B1%B1%E5%AE%99%E5%93%89-ebook/dp/B009KWUFRC?ref_=saga_dp_ss_dsk_dp&qid=1750266783&sr=1-290)
* ストレス反応を測定、ストレスマーカー候補の絞込  
* 1.9億円
::::
:::::

::::: {.column width = "5%"}
:::::

::::: {.column width = "65%"}
:::: {.incremental}
1. Fabrication (nonexisting data)  
1. Falsification (edited data)  
1. Insufficient scientific validity  
1. Lack of saved data and research notes (for reproducibility)  
1. Lack of informed consent after a protocol change  
::::
:::::

::::::::::

:::: {.incremental}
* この実験で新たなストレス・マーカーが得られるのか? 得てどうするのか?  
* 被験者負担、多額の資源に見合う新規の科学的価値があるか (NY Times記事)?  
::::

::: {style="font-size:60%; color:#33E7FF;"}
:::::::::: {.columns}
::::: {.column width = "70%"}
:::: {.incremental}
* [HI-SEAS](https://www.hi-seas.org/about-hi-seas) (Hawaii Space Exploration Analog and Simulation, 2013-17, NASA)、最大12ヶ月  
* [MARS500](https://www.esa.int/Science_Exploration/Human_and_Robotic_Exploration/Mars500/Mars500_quick_facts) (Russian Institute for Biomedical Problems, 2010-11, ロシア)、5名を520日間  
* [HERA](https://www.nasa.gov/mission/hera/) (Human Exploration Research Analog, 2024-, NASA)、45日  
* [CHAPEA](https://www.nasa.gov/humans-in-space/chapea/about-chapea/) (Crew Health and Performance Exploration Analog, 2023-), 宇宙船模倣空間で3名を378日  
::::
:::::

::::: {.column width = "30%"}
:::: {.incremental}
* CHAPEAにも「今更やる価値は」批判
* NASA宇宙システム部門副部長(Rachel MacCauley):  「精神安定のためにどれくらいスナックを載せる必要があるのかも技術計算に重要」
::::
::::: 
::::::::::

:::

----

## {background-color="#010638"}

[JAXA調査報告書 (2022年11月25日脚注3)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=3):

> _2014年に、<LightBlue>分科会の意見に耳を傾けず</LightBlue>有人部門のみの判断で、審査迅速化を目的に、評価要領が改訂され、<LightBlue>分科会の進捗評価は(引用者追記: 不承認が削除され)継続妥当・見直し必要の2択となっていた。</LightBlue>継続妥当となった場合でも多くの重要な指摘事項があったが、それらの指摘事項を担保する仕組みが無かった。_

:::: {.incremental}
* 有人技術宇宙部門 &xlarr; 研究開発「倫理審査委員会」+有人サポート委員会宇宙医学研究推進「分科会」が審査  
* 有人技術宇宙部門 &xrarr; 分科会の評価要領を改訂、<LightBlue>不承認できなくさせる</LightBlue>  
* 評価要領の改訂理由は「審査迅速化」、誰が承認したか報告書に記載なし  
::::

:::: {.fragment}
### JAXA事件の特徴: 組織的な研究倫理審査の軽視   
::::

:::: {.incremental}
* 理事長、副理事長、担当理事は[厳重注意と訓告(非懲戒)と1ヶ月役員報酬10%自主返納](https://www.jaxa.jp/press/2023/01/20230111-2_j.html#:~:text=%E7%B5%8C%E5%96%B6%E8%B2%AC%E4%BB%BB%E3%82%92%E6%98%8E%E7%A2%BA%E3%81%AB%E3%81%99%E3%82%8B%E3%81%9F%E3%82%81%E3%80%81%E5%B1%B1%E5%B7%9D%E7%90%86%E4%BA%8B%E9%95%B7%E3%80%81%E9%88%B4%E6%9C%A8%E5%89%AF%E7%90%86%E4%BA%8B%E9%95%B7%E5%8F%8A%E3%81%B3%E4%BD%90%E3%80%85%E6%9C%A8%E7%90%86%E4%BA%8B%E3%81%AF%E3%80%812023%E5%B9%B41%E6%9C%88%E3%81%AE1%E7%AE%87%E6%9C%88%E9%96%93%E3%80%81%E7%B5%A6%E4%B8%8E%E6%9C%88%E9%A1%8D%E3%81%AE10%25%E3%81%AE%E5%8F%97%E5%8F%96%E3%82%92%E8%BE%9E%E9%80%80%E3%81%84%E3%81%9F%E3%81%97%E3%81%BE%E3%81%99)([非懲戒](https://www.jaxa.jp/about/disclosure/data/disciplinary_fy2022.pdf))、[研究不正を実施した職員の14日停職](https://www.jaxa.jp/about/disclosure/data/disciplinary_20230110.pdf)(懲戒)、研究実施責任者への戒告(懲戒)  
* 末端が一番厳罰  
::::

----

## {background-color="#010638"}

::: {style="font-size:60%;"}
* (2015年1月9日: 分科会不承認)  
* (2015年3月6日: 分科会不承認)  
* (2016年2月5日: 研究開始)  
* 2017年11月27日: <Green>データ取り違え事案</Green>
   * 同時期に<Green>利害相反不適切処理事案</Green>  
   * 公表前成果を組織内手続きを経ずにスポンサー企業の[商品と合わせて学会で公表・共同プレスリリース(p.16)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=16)、[その調整を利益相反マネジメント委員長である事業部長が実施(p.26)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=26)  
* 研究倫理審査委員会+分科会: <Orange>「研究の適正性の検討が必要」</Orange>と報告  
* 2018年10月: 研究実施主体の有人部門が調査開始  
   * 2019年6月: JAXA職員Hに実験参加者Aから<Green>実験プロトコルからの逸脱者</Green>の連絡 
* 2019年7月: 有人部門が研究活動改善に向けたアクション・プラン作成  
* 2019年11月: 理事長が<Yellow>研究中止決定</Yellow>、有人部門長に成果とりまとめ、分科会評価と倫理審査委員会報告を指示  
* 2020年9月23日: 分科会+倫理審査委員会が理事長に対し「多数の研究倫理違反、研究遂行上不適切な事実、研究計画/手法上の問題に加え、データの改ざんが強く疑われる事実が判明」したため<Orange>「調査の実施の検討が必要」</Orange>と報告
* 2021年3月-4月: 委託業者による調査  
   * 2021年9月29日: 有人部門長決定による対策検討チームの聞き取りで職員Hが「外部通報を共有しなかった」と報告、有人部門が通報に関して調査開始  
* 倫理審査委員会が委託業者調査報告を審議、理事長に対し厚労相と文科相に<Orange>報告、公表を提言</Orange>
* 2022年11月25日: 厚労相と文科相に報告書提出、<Yellow>報告書公開</Yellow>
   * 2022年11月25日: 実験参加者AがJAXAへメール通報「2019年にJAXA職員Hに連絡した」
   * 2023年4月28日: 「研究対象者からの連絡に基づく調査報告書」<Yellow>公表</Yellow>
:::

----

## {background-color="#010638"}

### 最初の事案発生から研究中止まで2年、報告書公開まで5年

* 報告書執筆者は第三者委員会ではありません

### 研究倫理審査委員会+分科会は繰り返し調査を要請  
### 外部通報者への連絡も1年以上放置

----

## 経営者と管理職の弁 {background-color="#010638"}

佐々木理事

> _JAXAが<LightBlue>しっかり認識し、ルールを作ったことは間違いない</LightBlue>。一方、研究者一人一人が理解すべきだったが、末端まで伝わっていなかった。教育、指導を働きかけねばならなかった。_

:::: {.fragment}
* 分科会評価要領で不承認判断を許さなかったのに「しっかり認識し、ルールを作ったことは間違いない」?  
::::

----

## 経営者と管理職の弁 {background-color="#010638"}

:::: {.fragment}
有人宇宙技術部門小川事業推進部長(当該研究の研究実施責任者(古川宇宙飛行士)の上司)

>_JAXAとして、人の安全を守るという<LightBlue>医学研究の理解度、練度が本当に低かった</LightBlue>のでは。STAP問題があった時、（研究者は）自分のこととして納得しなければならなかったが、果たして医学研究ではどうだったか。深堀りして確認していなかったからこそ、この状態になった。<LightBlue>今こそ振り返り、是正すべきだ</LightBlue>。_
::::

:::: {.incremental}
* 有人部門が反対を無視して評価要領を変更したのは「医学研究の理解度、練度が本当に低かった」から?  
* 「今こそ振り返り、是正すべきだ」...他人事のような...  
::::

:::: {.fragment}
### 「JAXA事件の特徴: 組織的な研究倫理審査の軽視」という認識は当事者にない   
::::

----

## 経営者と管理職の弁 {background-color="#010638"}

[JAXA経営陣の説明](https://scienceportal.jst.go.jp/newsflash/20221128_n02/)

> _研究費は文科省科学研究費補助金（科研費）とJAXA予算を合わせ、約1億9000万円。佐々木理事はこの研究は<LightBlue>成果として発表しておらず、特定不正行為には当たらない</LightBlue>とした。その上で「<LightBlue>研究計画が稚拙で、十分な科学的合理性がなかった。</LightBlue>JAXAに医学研究の難しさの知見がなく、研究チームの人数が少なく体制を整えず、無理になってしまった。実験データが個人情報のためアクセスできる人を限ってしまい、周りが十分にチェックできなかった。国民の負託に応えられず、実験協力者の善意も裏切り深くお詫びする。一丸となって再発防止に取り組む」とした。_

:::: {.fragment}
* 成果公表しなければFFPではない、は誤解です。学会でも報告済み。  
::::

:::: {.fragment}
> _この指針は、我が国の研究者等 により実施され、又は日本国内において実施される人を対象とする生命科学・医学系研究 を対象とする。_ [厚労省「指針」第3「適用範囲」1「適用される研究」](https://www.mhlw.go.jp/content/001077424.pdf#page=10)
::::

----

## 経営者と管理職の弁 {background-color="#010638"}

:::: {.fragment}
* 科学的意義の低い研究と分科会が判断  
::::

:::: {.fragment}
> _最も問題なのは、科学的合理性のある研究計画が検討されておらず、かつ科学的に適切な方法で遂行されなかったことであり、またそのように努めなかったこと_ ([調査報告書](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=7))
::::

### 意義の低い研究の実施=研究倫理(Emanuel et al. 2, 3, 6)の逸脱

### そう指摘をする分科会を沈黙させた組織運営

----

## JAXA事件の背景([報告書, p.30](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=30)){background-color="#010638"}

### 組織として閉鎖環境実験成果を出すことを目指す

* 1985年から莫大な国費を用いてISS計画を推進  
* ISSでの成果の創出が有人部門にとっての優先事項となっていた  
* 2015年時点で、5年後の目指す成果の出口は「地上の閉鎖環境等における実証とISSにおける一部の検証を終え、火星に向けた研究を開始」すること  
* 2015年1月の第45回分科会(不承認)、第46回分科会(不承認)、同年9月第47回分科会(承認)、倫理審査委員会同年12月の第75回倫理審査委員会、2016年1月第76回倫理審査委員会(承認)

----

## {background-color="#010638"}

### 2回も科学的意義で落第点

### なのに無理やり実施: 経営者もそうだが、提案者も相当

>_倫理審査委員会や分科会での<Orange>指摘事項に十分応えることなく研究を進めようとした研究チームの不誠実な態度</Orange>、それを容認してしまう倫理意識の欠如も背後要因となった。_ [調査報告書(p.19)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=19)

> _倫理審査委員会に提出した研究等実施計画書（第 3 回閉鎖試験）に「分科会において、妥当な計画であるとの評価を受けた」と<Orange>虚偽の事実が記載</Orange>されていた_ [同(p.22)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=22)

### 「[専門家に任せる信頼の気持ちが勝ってしまった](https://scienceportal.jst.go.jp/newsflash/20230113_n01/#:~:text=%E5%B0%82%E9%96%80%E5%AE%B6%E3%81%AB%E4%BB%BB%E3%81%9B%E3%82%8B%E4%BF%A1%E9%A0%BC%E3%81%AE%E6%B0%97%E6%8C%81%E3%81%A1%E3%81%8C%E5%8B%9D%E3%81%A3%E3%81%A6%E3%81%97%E3%81%BE%E3%81%A3%E3%81%9F%E3%80%82)」ギフト・オーサーシップではないが、共著者なのに分担を超えて確認せず

### 研究体制の不備: 能力不足ならシニア研究者とペアにすべき

> _研究者としての<Orange>適性を見極めず医師であることを理由に研究計画の策定から実行まで</Orange>を実施させた_ [同(p.24)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=24)

----

## JAXA当事者はなぜFF(P)やRE違反をしたのか?{background-color="#010638"}

* 非研究者なので研究倫理について考えていなかった
* 経営者が非研究者(研究能力の乏しい人)に多額予算の研究を実施させ、成果を出させようとした
* 分科会審査という安全装置が外された(邪魔者扱い)
* 人の配置、権限配分、予算、期間を経営者が考えていなかった
* 経営者が研究の見識を(経営者に相応しいほど)持っていなかった

----

## JAXA当事者はなぜFF(P)やRE違反をしたのか?{background-color="#010638"}

* Gino事件とは違い、個人的な金銭利得はほぼゼロ  
* [組織の目標(成果を出す)と手段(人員配置、チェック機能、予算や期間)が整合的ではなかったために、末端の人員が帳尻あわせに研究不正をしたのでは?](https://scienceportal.jst.go.jp/newsflash/20230113_n01/#:~:text=%E9%99%90%E3%82%89%E3%82%8C%E3%81%9F%E4%BA%BA%E5%93%A1%E3%81%A8%E5%8E%B3%E3%81%97%E3%81%84%E3%82%B9%E3%82%B1%E3%82%B8%E3%83%A5%E3%83%BC%E3%83%AB%E3%81%AE%E4%B8%AD%E3%81%A7%E3%81%84%E3%81%8B%E3%81%AB%E7%A0%94%E7%A9%B6%E3%82%92%E9%80%B2%E3%82%81%E3%82%8B%E3%81%8B%E3%80%81%E5%BF%85%E6%AD%BB%E3%81%AB%E3%81%AA%E3%81%A3%E3%81%A6%E3%81%97%E3%81%BE%E3%81%A3%E3%81%9F%E3%80%82%E3%83%9E%E3%83%8D%E3%82%B8%E3%83%A1%E3%83%B3%E3%83%88%E3%81%AB%E9%81%A9%E5%88%87%E3%81%AB%E5%AE%9F%E6%83%85%E3%82%92%E4%BC%9D%E3%81%88%E3%81%A6%E8%AA%BF%E6%95%B4%E3%81%99%E3%81%B9%E3%81%8D%E3%81%A0%E3%81%A3%E3%81%9F%E3%80%82)
* サラリーマン研究者解釈=日本ぽい
   * ダイハツ工業、豊田自動織機、三菱自動車工業、ビッグモーターの検査不正  
   * 研究生産性を高めるのは良いことですが、資源投入を減らし過ぎると弊害があります  
   * 研究予算などの入り口で搾り、人事評価などの出口で圧迫すると、研究倫理を逸脱する可能性を高めます

----

## アジ研には研究倫理を尊重する文化があります{background-color="#010638"}

## 今年の研究成果に不安・不透明感を持つならば、早めに上長に相談しましょう{background-color="#010638"}

## 適度な競争や批判の圧力の下、若手を育てます{background-color="#010638"}

----

# まとめ: 社会から託された信任を裏切らない  {background-color="#010638"}

References

::: {#refs}
:::

### 2025-06-18 posts/ethics/KenkyuRinriSlides.rmd
added DelabastitaRubens repo and ethics repo

---
title: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "fonts.css", "ninjutsu", "xaringan-themer.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
header-includes:
    - \usepackage{amsmath}
    - \usepackage{RefManageR}
# setwd(path <- "c:/seiro/docs/IDE/jinzai/kensyu/2024/"); rmarkdown::render("KenkyuRinriSlides.Rmd")
# git status
# git config --global user.name "SeiroIto"
# git add c:/seiro/docs/IDE/jinzai/kensyu/2024
## I added 2023/ in exclude to untrack 2023 folder
# git commit -m "Adding kenkyuu rinri rmd file in 2024"
# git branch -M main
# git remote add origin https://github.com/SeiroIto/Miscellaneous.git
# git push -u origin main
---
class: inverse, middle, center
```{r setting, eval = T, cache = T, warning = F, message = F, echo = F}
library(RefManageR)
myBib <- ReadBib("c:/seiro/settings/TeX/seiro.bib")
BibOptions(max.names = 3)
```
```{r xaringan setup, include=FALSE, warning=FALSE}
library(RefManageR)
library(xaringanthemer)
####style_mono_dark(base_color = "#cbf7ed")
####style_mono_light(base_color = "#23395b")
style_solarized_light(
  text_color = "black", 
  title_slide_background_color = "darkblue",
  link_color = "green")
#### xaringancolor setup
library(xaringancolor)
xaringancolor::setup_colors(
  red ="red",
  green = "green",
  blue = "blue"
)
```
```{r xaringan-panelset, echo=FALSE, warning = F}
library(xaringanExtra)
use_panelset()
```
<style>
.red {color: #FF0000;}
.green {color: #00FF00;}
.blue {color: #0000FF;}
.lightblue {color: #33E7FF;}
</style>
```{r flair_color, echo=FALSE}
library(flair)
red <- "#FF0000"
green <- "#00FF00"
blue <- "#0000FF"
```
```{css echo=FALSE}
.highlight-last-item > ul > li,
.highlight-last-item > ol > li {
  opacity: 0.5;
}
.highlight-last-item > ul > li:last-of-type,
.highlight-last-item > ol > li:last-of-type {
  opacity: 1;
}
.inverse {
  background-color: #00003c;
  color: #d6d6d6;
  text-shadow: 0 0 20px #333;
}
mark.red {
    color:#ff0000;
    background: none;
}
mark.blue {
    color:#0000A0;
    background: none;
}
.my-style {
  font-weight: bold;
  font-style: italic;
  font-size: 1.5em;
  color: red;
}
.large {
  font-size: 2em;
}
blockquote {
  border-left: .2px solid #275d38;
  margin: -5px 80px -5px 20px;
  padding-top: -0.5px;
  padding-bottom: -0.5px;
  line-height: 1.35em;
}
.pull-left60 {
  float: left;
  width: 57%;
  padding-right: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-right40 {
  float: right;
  width: 37%;
  padding-left: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-left70 {
  float: left;
  width: 67%;
  padding-right: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-right30 {
  float: right;
  width: 29%;
  padding-left: -4px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-left75 {
  float: left;
  width: 73%;
  padding-right: 2px
  margin-top: -5em;
  margin-bottom: 0em;
}
.pull-right25 {
  float: right;
  width: 22%;
  padding-left: 2px
  margin-top: -5em;
  margin-bottom: 0em;
}
.pull-left50 {
  float: left;
  width: 50%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: 0em;
}
.pull-right50 {
  float: right;
  width: 50%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: 0em;
}
.pull-right50jaxa {
  float: right;
  width: 50%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: -5em;
}
.pull-left20 {
  float: left;
  width: 18%;
  padding-right: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-right80 {
  float: right;
  width: 79%;
  padding-left: -4px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-left100 {
  float: left;
  width: 100%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: 0em;
}
.pull-right {
  float: right;
  width: 100%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: 0em;
  margin-bottom: 0em;
}
.center2 {
  margin: 0;
  position: absolute;
  top: 50%;
  left: 50%;
  -ms-transform: translate(-50%, -50%);
  transform: translate(-50%, -50%);
}
```
# 研究倫理

## 2024年度新入職員研修

<br>
### 開発研究センター 
### 伊藤成朗
### 2024年4月17日(水)
---
class: inverse, middle, center

### https://seiroito.github.io/Miscellaneous/2024/KenkyuRinriSlides.html

---
class: inverse, middle, center
## 仕事は社会・業界の行動規範の範囲内で実施します

## 法令、組織の規則、社会通念、業界通念

## 研究業界(学界+その周辺)にも特有の規範=研究倫理があります
---
class: inverse, middle, center
## 本講義の目的: 研究コミュニティに身を置く者が研究倫理逸脱で失敗しないようにすること

## 崇高な研究哲学論ではなく、実際的な自衛の話
---
class: inverse, middle, center
# 研究倫理 = {狭義の研究倫理,  研究公正}
---
class: inverse
.pull-left75[]
.pull-right25[### 極言すると  ]

--

### 研究倫理とは

--

## 研究で対象社会に.red[迷惑をかけない]という原則

--

### 研究公正とは

--

## 研究で.red[ごまかしをしない]という原則

--

### 対象社会は日本や学界も含みます
---
class: inverse
<p style="margin-top:0cm;">
</p>

### 研究は社会と契約を結んで実施しています

--

### よって、これらは最低限の原則です

--

### それでも足をすくわれる研究者はいます

--

### どう判断すれば分からなくなったとき、大原則に戻ると考えやすいです

--

## .red[社会から託された信任を裏切らない]

---
class: inverse
name: REAndRI

.pull-left70[
<img src="REAndRI.jpg" height="400px">
]

--

.pull-right30[
<br>
* Do good for the society (not just for oneself/clients).  
]

--

.pull-right30[
<br>
* Be honest to the scientific standard (not just to oneself).  
]

--

.pull-right30[
ごまかしのある研究は社会に迷惑をかけるように、研究公正と研究倫理はきれいに分かれず、相互に関連する部分があります
]

--
<p style="margin-top:3cm;">
.pull-left100[

]
</p>

.pull-left100[
]

.pull-left100[
[研究公正の鳥瞰図研究](https://www.vitae.ac.uk/vitae-publications/research-integrity-a-landscape-study) が示す研究公正のキーワード: honest, rigorous, transparent, open, respect and stewardship of research, care for the people of research.
]
---
class: inverse, middle, center
background-image: url("JAXASumimasen.jpg")
background-size: 100%
# .lightblue[研究公正 research integrity]
---
class: inverse, middle, center

写真はJAXA記者会見映像からキャプチャしました
---
class: inverse, middle, center
# 科学的根拠をもとに研究し、得た知識を歪めずに発表すること。
---
class: inverse, middle, center
# 研究公正は、やってはいけないことで定義されます。
---
class: inverse
<p style="margin-top:-1cm;">
</p>
## .red[一発退場 Research misconducts] (.lightblue[FFP], [厚労省研究不正ガイドライン「特定不正行為」](https://www.mhlw.go.jp/file/06-Seisakujouhou-10600000-Daijinkanboukouseikagakuka/0000152685.pdf#page=8))

--

<p style="margin-top:-2cm;">
</p>
.pull-left70[
### .lightblue[F].orange[abrication ねつ造]「存在しないデータや研究成果を作成すること。」  
### .lightblue[F].orange[alsification 改ざん]「操作」によって「データ」や「成果などを真正でないものに加工すること」。  
### .lightblue[P].orange[lagerism 剽窃、盗用] 他の研究者の「アイディア」「方法」「データ」「結果」を「了解または適切な表示なく流用すること」。
]
--
.pull-right30[
* やると研究者人生が終わります。  
]
--
.pull-right30[
* 理由: 研究者社会は相互信頼で成立しています。ウソをつかないことを前提に評価がされて、知識が共有されます。裏切ると、研究者社会全体が資源を浪費するので、厳罰処分されます。不正が露見するまで、他の研究者は存在しない現象を追求して苦しみます。  
]
--
.pull-right30[
   * STAP論文: 追試資源や継続研究資源の浪費、周辺研究まで疑いの目
]
---
class: inverse
<p style="margin-top:-1cm;">
</p>
## .yellow[信用失墜] .orange[Questionable research practices (QRPs)]
<p style="margin-top:0cm;">
</p>
.pull-left60[
  .large[
* データの未保存や再現性の欠如  
* 不適切な統計手法  
* ギフト・オーサーシップ ([CRediT](https://www.elsevier.com/authors/policies-and-guidelines/credit-author-statement))  
* 論文の多重投稿  
* 利益相反によるバイアス  
  ]
]

--

<p style="margin-top:1cm;">
</p>
.pull-right40[
* 識者によって内容が異なるので幾つか列記  
* $P$-hacking: E.g.,  望んだ結果が出るまで推計を試行し、その結果のみ提示すること  
* GAは贈賄双方に便益あり、[なくなりにくい、でも、著者責任も付いてくる](https://www.bmj.com/content/309/6967/1456#:~:text=Perhaps%20this%20scandal%20will%20finally%20undermine%20gift%20authorship.%20At%20the%20very%20least%20it%20should%20make%20researchers%20think%20hard%20about%20the%20responsibilities%20that%20come%20with%20putting%20their%20names%20on%20papers.)   
* 刊行された論文を翻訳して投稿してはいけません ([International Committee of Medical Journal Editors](https://www.icmje.org/recommendations/browse/publishing-and-editorial-issues/overlapping-publications.html#:~:text=Authors%20should%20not%20submit%20the%20same%20manuscript%2C%20in%20the%20same%20or%20different%20languages%2C%20simultaneously%20to%20more%20than%20one%20journal.))、翻訳であることを明示しても危ういので、翻訳論文はワーキング・ペーパーに留めましょう  
]

--

.pull-left100[
See a list of QRPs used in Dutch national survey (DNSRI) summarised [here](https://htmlpreview.github.io/?https://github.com/SeiroIto/PaperNotes/blob/main/QRPsInDutchNationalSurvey_Tufte.html)
]
---
class: inverse
### 蔓延度は不明ですが、結構深刻では?

.orange[例1:] Dutch National Survey on Research Integrity [`r Citep(myBib, "Gopalakrishnaetal2022")`](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0263023&fbclid=IwAR17MW5qz2wpZXQz6crAQVrsDbGsx0DKeE5ZIuvIywF137rYlf_5hXz1Alw):

* 15大学+7医学研究所、大学院生以上、`r 63778-1913`人メール送達、6813人回答(`r round(100*6813/(63778-1913), 0)`%)、過去3年  
* 人文科学(全体の`r round(100*636/6813, 0)`%)、社会・行動科学(同`r round(100*1965/6813, 0)`%)、生命・医学科学(同`r round(100*2747/6813, 0)`%)、自然・工学科学(同`r round(100*1465/6813, 0)`%)  
* 大学院生(全体の`r round(100*2013/6813, 0)`%)でも準教授以上(同`r round(100*2066/6813, 0)`%)でも傾向はほぼ同じ

<p style="margin-top:-.5cm;">
</p>
.pull-left50[
FF(P) 8.3% engaged in any FF(P).  
QRP 51.3% engaged frequently in at least 1 QRP.  
]
--
.pull-right50[
Beh&Soc.Sci 5.7%  
Beh&Soc.Sci 50.2%
]

--

.orange[例2:] The International Research Integrity Survey `r Citep(myBib, "Allumetal2022")`:

* Web of Scienceのジャーナル21894誌・本、2016-2020年、816万 &rarr; 392万 &rarr; 375万 &rarr; 307万2372 (対象国: 米欧)  
* Stratified random sampling (strata=natural, medical, social sciences and humanities)  
* (おそらく)307万にemail送付  &rarr; 73757が回答 (ウェイト後回答率=7.2%)  &rarr; 64074 (分析標本サイズ)  
* 過去3年のQRP

--

<p style="margin-top:-1.0cm;">
</p>
.pull-left50[
ギフト・オーサーシップ: 69%(欧)、55%(米)  
予想と矛盾する結果の秘匿: 25%以上(欧)、20%以上(米)  
剽窃: 10%未満(欧)、10%未満(米)  
]

--
.pull-right50[
自然科学: 30%以上(欧)、30%以上(米)  
医学科学: 30%以上(欧)、25%以上(米)  
社会科学: 30%以上(欧)、20%以上(米)  
人文科学: 20%以上(欧)、15%以上(米)  
]

---
class: inverse
### 蔓延度は不明ですが、結構深刻では?

.orange[例3:] [Data Collada (psychologist blog)](https://datacolada.org/) vs. Francesca Gino (HBS prof), 2021-

* Convincing evidence of data tampering in several of her work on dishonesty (incl. with Dan Ariely!)  
* HBS put Gino on an admin leave, hired file forensic investigators, and sent retraction requests to journals  
* Gino sued these folks and Harvard for lost incomes and damages which is costing them exhorbitant legal fees, and asking for "not less than $25 million"  
* [The problem, though, is that it will take years --- and be extraordinarily expensive --- to settle](https://www.vox.com/future-perfect/2023/8/9/23825966/francesca-gino-honesty-research-scientific-fraud-defamation-harvard-university#:~:text=The%20problem%2C%20though%2C%20is%20that%20it%20will%20take%20years%20%E2%80%94%20and%20be%20extraordinarily%20expensive%20%E2%80%94%20to%20settle)
* [150人あまりの共著者たちはどの論文は不正がないか調べる羽目に](https://www.science.org/content/article/after-honesty-researcher-s-retractions-colleagues-expand-scrutiny-her-work#:~:text=This%20week%2C%20Simonsohn%20and%20five%20other%20Gino%20co%2Dauthors%20emailed)

--

.orange[例4:] [国立医薬品食品衛生研究所職員による研究活動上の不正行為及びISO/IEC 17025認定審査対応における不適切行為に関する調査結果の公表](https://www.nihs.go.jp/oshirasejoho/Press-Release/20231226/index.html) 2023年12月26日、食品衛生部長によるFabrication, Falsification認定
---
class: inverse

## なぜFFPをするのか

--

### Gino: HBS給与は1.4億円+印税+講演料、ばれなければ大金

--

* 日本の文系(<2000万円)だとFFPは割に合いません  
* 国立医薬品食品衛生研究所所長で1900万円くらい

--

## 結構ばれる: Ginoは追試、食品衛生部長は内部通報

--

## 貧乏くじ

--

### オッズをよく知らないまま、自ら進んで引くのは賢くない
---
class: inverse

## 自分は本当に大丈夫か?

--

<p style="margin-top:-1.5cm;">
</p>
.pull-left50[
### アジ研: 外部から内部研究者成果に剽窃を指摘される事例(2010年代、2024)

### 公正な研究?  
]

--

.pull-right50[
<br>
* どのように組織が対処したか知ると学びの機会になります  
]

--

.pull-right50[
* 成果を冗長に書く  
* 要約が不正確  
* どのような過程を踏んだか明記しない  
* 図表の脚注に留意事項を説明するが本文は無記載  
* 研究なのにダークパタン  
* 過去著作の数段落を明示せずコピー、参考文献に掲載  
]

--
.pull-left100[
]
.pull-left50[
### Githubで研究過程を公開

### ピア・レビューを活用  
]

--
<p style="margin-top:1.0cm;">
</p>
.pull-right50[

### [こんな感じ](https://github.com/SeiroIto)です
]
---
class: inverse

## なぜQRPをするのか

--

## 成果を出したい、でも、今のままだと成果にならないと考えている

--
.pull-left50[
### 内容が不十分な場合

* 結果がはっきりしない  
* 2つの結果が矛盾する  
* 何を議論すべきかもよく分からない  
]

--

.pull-right50[
<p style="margin-top:.5cm;">
</p>

* 不適切な統計手法  
* データの未保存や再現性の欠如  
* 曖昧な記述と結論...QRP?  
   * 結論は曖昧です、と明確に書けばok  
   * 単に記述を不明瞭にするのはごまかし=QRP

* 過程(データとコード)を公開  
* Pre analysis planを公開・登録  
]
---
class: inverse

## なぜQRPをするのか

--

<p style="margin-top:-1.5cm;">
</p>

.pull-left70[
### 自信、発信力が無い場合

* ビッグ・ネームの推薦を利用したい  
]

--

.pull-right30[

<p style="margin-top:1.5cm;">
</p>

* ギフト・オーサーシップ  
* 反例: ネットワーク理論: Erdős-Rényiモデル. [Erdős](https://en.wikipedia.org/wiki/Paul_Erd%C5%91s)は1525論文掲載  
]

--

<p style="margin-top:-2.5cm;">
</p>

.pull-left100[
]
.pull-left70[
### 時間がない場合

* 今までの研究に手を加えれば新しいものに?  
]

--

.pull-right30[

<p style="margin-top:1.5cm;">
</p>

* 多重投稿、多重出版  
* 機械学習で簡単に判明  
]

<p style="margin-bottom:-1.5cm;">
</p>

--

<p style="margin-top:-1.5cm;">
</p>
.pull-left100[
]
.pull-left70[
### スポンサーと利益相反するが、研究は健全で余計な詮索をされたくないと考える場合

* 利益相反があることを表明しない  
]

--

.pull-right30[

<p style="margin-top:1.5cm;">
</p>

* 利益相反

<p style="margin-top:0.5cm;">
</p>

* 簡単に判明  
]

---
class: inverse
## おまけ: 大規模言語モデル(LLM)時代のリスク

--

### ChatGPT謹製論文「この論文書いて」?

* [「ChatGPTでーす、もう一回答えますか?」チャットボット入りの論文](https://pubpeer.com/publications/A9C4E7545D0D3D00A561576BC3B603#)  
* [「AIなので私はアクセスできないのですが」と本文に書いてある論文](https://pubpeer.com/publications/24174D03E640768EDD528C01DD4B8E)

--

### 「この分野の研究教えて」「この論文の評価を教えて」とLLMを使って良いのか?

--

* 学生が指導教官に尋ねているのと同じです

--

* 返答をそのまま本文に使うと問題です

--

   * Ghost authorshipになります

--

* 仮に返答に剽窃がない場合でも...返答が正しいか自信持てますか?...Hallucinations  
---
class: inverse
## おまけ: 大規模言語モデル(LLM)時代のリスク

### 「この推計コード書いて」「違う表現で書いて」とLLMを使って良いのか?

--

* 研究者がRAを使っているのと同じです

--

* LLMに謝辞を書けば良い...のか? よく分からない  
   * [ICMJE推奨事項](https://www.icmje.org/news-and-editorials/icmje-recommendations_annotated_jan24.pdf#page=3)ではどのように使ったかを具体的に明示するように要求

--

* ...出力が正しいか自信持てますか?  
   * 最初のとっかかりを書いてもらうのには使えるか  
   * 他人のコードが何をやっているのかを教えてもらうことはできます

---
class: inverse
## おまけ: 大規模言語モデル(LLM)時代のリスク

### LLMとどう付き合うか

* 自信過剰な物知りのRA(無休・無給)程度に考えておけばいいのでは  
   * 当然、知ったかぶりして間違えます

--

* 利用者全員が同じ視点を前提にするようになり、LLMをベースにすると独創性が失われます  
   * 同じLLMでは、類似の質問への回答内容は収斂するため

--

* 真偽を確かめやすい単純労働を担わせるには良い相手かも  
   * 「以下の出来事を時系列順に並べ替えて、日付と出来事の2列からなるtab separated text形式で出力して」

---
class: inverse
## おまけ: 大規模言語モデル(LLM)時代のリスク

### 自分の公開論文が学習されて謹製されてしまう可能性

* Github公開は版管理機能があるのでオリジナル論文の証拠になります  
* アジ研DPも同様ですが手間がかかります  
---
class: inverse, middle, center
background-image: url("JAXASumimasen.jpg")
background-size: 100%
# .lightblue[研究倫理 research ethics]
---
class: inverse, middle, center
# 研究対象(者の人権)保護、研究コミュニティの評判維持
---
class: inverse, middle, center
# 研究倫理は、守るべきことで定義されます
---
class: inverse
.pull-left50[
<img src="the-constant-gardener-original.jpg" width="400px">
]
.pull-right50[
<img src="TheExperiment.jpg" width="400px">
]
---
class: inverse
### 3つのレイヤー `r Citep(myBib, "Mustajoki2017", after = ", p.37")`

.pull-left20[
 1. 法的義務  
]
.pull-right80[
個人情報保護、人権・動物の権利保護など
]
.pull-left20[
<ol start="2">
<li> ガイドライン  
</ol>
]
.pull-right80[
インフォームド・コンセント、研究の正確さ、貢献の正確さ、研究協力者の表示、など
]
.pull-left20[
<ol start="3">
<li>  暗黙の了解  
</ol>
]
.pull-right80[
? (分野や組織ごとにあるルールなど、らしい)

* 論文への質問に著者は真摯に答える  
* 内容を批判しても著者の人格まで批判しない  
* 読む前後で自分の知識がどう変わったか書く  
* 研究をより良くするため、読者がより正確に理解するためにコメント・批判をする  
* 問題の指摘で終わらず結論にどう影響するかまで書く

最後の3つはMacartan Humphreysの[How to critique](https://macartan.github.io/teaching/how-to-critique)から抜粋  
]

---
class: inverse
途上国研究における研究倫理8ポイント `r Citep(myBib, "Emanuel2004")`

1. .orange[Collaborative partnership]	対象個人や社会が便益を受ける研究か。 
1. .orange[Social value]	社会的に価値のある研究か。 
1. .orange[Scientific validity]	.lightblue[科学的裏付け](知見を得る方法論)はあるか。 
1. .orange[Fair subject selection]	 参加者は公平に選ばれたか。 
1. .orange[Favorable risk-benefit ratio]	リスク相応の便益はあるか。個人リスク、社会便益。 
1. .orange[Independent review]	.lightblue[独立機関が審査]したか。 
1. .orange[Informed consent]	.lightblue[インフォームド・コンセント]を得たか。 
1. .orange[Respects for recruited participants and study, communities] 個人や対象社会の.lightblue[意思を尊重]し.lightblue[経過・結果を伝達]したか。

--

殆どの社会科学研究(=観察研究)

* 便益は無く、リスクも個人情報漏洩以外は僅少、社会的価値もよく分からないので、参加者を公平に選ぶべき理由は倫理に照らしてあまりない。  
* 審査の独立性、インフォームド・コンセント、個人情報保護、意思の尊重は手続きに関する考慮。科学的裏付けは研究をする動機。 
* あらゆる調査・研究で満たすべき項目であり、研究倫理審査が特別なことをしているわけではない。 
---
class: inverse
.panelset[
	.panel[.panel-name[研究倫理チェックリスト(あらすじ)]

企画段階:  
* やる科学的価値=社会的価値はあるか...既存研究と違いはあるか?  
* 対象選定、標本サイズ、質問票などの研究デザインは十分か=知見が得られる見込みがあるか?  
* 個人情報保護の手段は考えたか?

実施段階:  
* 研究説明書(research information sheet): 研究の概要と連絡先を説明   
* 研究同意書(informed consent): 参加者の権利+何に同意したのか+署名   
   * いつでも参加拒否できる、いかなる質問にも回答拒否できるなど  
* 必要に応じて研究倫理審査委員会の推薦状、現地IRB (institutional review board)の推薦状・許可状

分析段階:  
* ハードコピーは鍵付き書庫に保存したか  
* 分析用データは個人識別子と分離したか  
	]
	.panel[.panel-name[アジ研の研究倫理審査]

* おもに介入研究が対象: 本審査  
* 雑誌によっては観察研究でも要求する場合がある: 迅速審査  
* 判断に迷う場合は研究倫理審査委員会事務局(研究企画課)へ相談を  
* 研究倫理審査委員会は研究倫理に適合しながら研究実施できる方法を共に考える場所です。  
   * 例: COVID-19下で感染リスクを日常生活以上に高めずに調査する方法などを検討  
	]
]

---
class: inverse, center, middle
## 経営者が研究倫理審査や研究能力を軽視すると、配下の研究者が研究倫理・研究公正から逸脱するのを防げない
---
class: inverse
.pull-left50[
JAXAの研究倫理・研究公正違反 (2016-2017)  
]
.pull-right50[
<img src="JAXAlogo.jpg" height="75px">
<img src="JAXA_pic.jpg" height="75px">
]

<p style="margin-top:-1cm;">
</p>
.pull-left50[
<br><br><br>
* 5回の閉鎖試験、計42人の研究対象者、2週間の閉鎖環境負荷  
* ストレス反応を測定、ストレスマーカー候補の絞込  
* 1.9億円
]

<p style="margin-top:-1cm;">
</p>
.pull-right50[
1. Fabrication (nonexisting data)  
1. Falsification (edited data)  
1. Insufficient scientific validity  
1. Lack of saved data and research notes (for reproducibility)  
1. Lack of informed consent after a protocol change  
]

--

.pull-left100[
* この実験で新たなストレス・マーカーが得られるのか? 得てどうするのか?  
* 被験者に負担をかけ、多額の資源を使っても類似実験にはない科学的価値があるか (NY Times記事)?  
  * [HI-SEAS](https://www.hi-seas.org/about-hi-seas) (Hawaii Space Exploration Analog and Simulation, 2013-17, NASA)、最大12ヶ月  
  * [MARS500](https://www.esa.int/Science_Exploration/Human_and_Robotic_Exploration/Mars500/Mars500_quick_facts) (Russian Institute for Biomedical Problems, 2010-11, ロシア)、5名を520日間  
  * [HERA](https://www.nasa.gov/mission/hera/) (Human Exploration Research Analog, 2024-, NASA)、45日  
  * [CHAPEA](https://www.nasa.gov/humans-in-space/chapea/about-chapea/) (Crew Health and Performance Exploration Analog, 2023-), 宇宙船模倣空間で3名を378日  
     * CHAPEAにも「今更やる価値は」批判があります  
     * 「精神安定のためにどれくらいスナックを載せる必要があるのかも技術計算に重要」NASA宇宙システム部門副部長(Rachel MacCauley):  
]
---
class: inverse
[JAXA調査報告書 (2022年11月25日脚注3)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=3):

> _2014年に、.lightblue[分科会の意見に耳を傾けず]有人部門のみの判断で、審査迅速化を目的に、評価要領が改訂され、.lightblue[分科会の進捗評価は(引用者追記: 不承認が削除され)継続妥当・見直し必要の2択となっていた。]継続妥当となった場合でも多くの重要な指摘事項があったが、それらの指摘事項を担保する仕組みが無かった。_

--

* 有人技術宇宙部門 &xlarr; 研究開発「倫理審査委員会」+有人サポート委員会宇宙医学研究推進「分科会」が審査

--

* 有人技術宇宙部門 &xrarr; 分科会の評価要領を改訂、不承認できなくさせる

--

* 評価要領の改訂理由は「審査迅速化」、誰が承認したか報告書に記載なし

--

### JAXA事件の特徴: 組織的な研究倫理審査の軽視

--

* 理事長、副理事長、担当理事は[厳重注意と訓告(非懲戒)と1ヶ月役員報酬10%自主返納](https://www.jaxa.jp/press/2023/01/20230111-2_j.html#:~:text=%E7%B5%8C%E5%96%B6%E8%B2%AC%E4%BB%BB%E3%82%92%E6%98%8E%E7%A2%BA%E3%81%AB%E3%81%99%E3%82%8B%E3%81%9F%E3%82%81%E3%80%81%E5%B1%B1%E5%B7%9D%E7%90%86%E4%BA%8B%E9%95%B7%E3%80%81%E9%88%B4%E6%9C%A8%E5%89%AF%E7%90%86%E4%BA%8B%E9%95%B7%E5%8F%8A%E3%81%B3%E4%BD%90%E3%80%85%E6%9C%A8%E7%90%86%E4%BA%8B%E3%81%AF%E3%80%812023%E5%B9%B41%E6%9C%88%E3%81%AE1%E7%AE%87%E6%9C%88%E9%96%93%E3%80%81%E7%B5%A6%E4%B8%8E%E6%9C%88%E9%A1%8D%E3%81%AE10%25%E3%81%AE%E5%8F%97%E5%8F%96%E3%82%92%E8%BE%9E%E9%80%80%E3%81%84%E3%81%9F%E3%81%97%E3%81%BE%E3%81%99)([非懲戒](https://www.jaxa.jp/about/disclosure/data/disciplinary_fy2022.pdf))、[研究不正を実施した職員の14日停職](https://www.jaxa.jp/about/disclosure/data/disciplinary_20230110.pdf)(懲戒)、研究実施責任者への戒告(懲戒)  
* 末端が一番厳罰

---
class: inverse
* (2015年1月9日: 分科会不承認)  
* (2015年3月6日: 分科会不承認)  
* (2016年2月5日: 研究開始)  
* 2017年11月27日: .green[データ取り違え事案]  
   * 同時期に.green[利害相反不適切処理事案]  
   * 公表前成果を組織内手続きを経ずにスポンサー企業の[商品と合わせて学会で公表・共同プレスリリース(p.16)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=16)、[その調整を利益相反マネジメント委員長である事業部長が実施(p.26)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=26)  
* 研究倫理審査委員会+分科会: .orange[「研究の適正性の検討が必要」]  
* 2018年10月: 研究実施主体の有人部門が調査開始  
   * 2019年6月: JAXA職員Hに実験参加者Aから.green[実験プロトコルからの逸脱者]の連絡  
* 2019年7月: 有人部門が研究活動改善に向けたアクション・プラン作成  
* 2019年11月: 理事長が.yellow[研究中止決定]、有人部門長に成果とりまとめ、分科会評価と倫理審査委員会報告を指示  
* 2020年9月23日: 分科会+倫理審査委員会が理事長に対し.orange[「調査の実施の検討が必要」] 
* 2021年3月-4月: 委託業者による調査  
   * 2021年9月29日: 有人部門長決定による対策検討チームの聞き取りで職員Hが「外部通報を共有しなかった」と報告、有人部門が通報に関して調査開始  
* 倫理審査委員会が委託業者調査報告を審議、理事長に対し厚労相と文科相に.orange[報告、公表を提言]  
* 2022年11月25日: 厚労相と文科相に報告書提出、.yellow[報告書公開]  
   * 2022年11月25日: 実験参加者AがJAXA職員Hへの連絡したことをJAXAへメール通報
   * 2023年4月28日: 「研究対象者からの連絡に基づく調査報告書」.yellow[公表]  
---
class: inverse

### 最初の事案発生から研究中止まで2年、報告書公開まで5年

* 報告書執筆者は第三者委員会ではありません

### 研究倫理審査委員会+分科会は繰り返し調査を要請  
### 外部通報者への連絡も1年以上放置
---
class: inverse
佐々木理事  
> _JAXAが.lightblue[しっかり認識し、ルールを作ったことは間違いない]。一方、研究者一人一人が理解すべきだったが、末端まで伝わっていなかった。教育、指導を働きかけねばならなかった。_

--

* 分科会評価要領で不承認判断を許さなかったのに「しっかり認識し、ルールを作ったことは間違いない」?

--

<br>
有人宇宙技術部門小川事業推進部長(当該研究の研究実施責任者(古川宇宙飛行士)の上司)  
>_JAXAとして、人の安全を守るという.lightblue[医学研究の理解度、練度が本当に低かった]のでは。STAP問題があった時、（研究者は）自分のこととして納得しなければならなかったが、果たして医学研究ではどうだったか。深堀りして確認していなかったからこそ、この状態になった。.lightblue[今こそ振り返り、是正すべきだ]。_

--

* 有人部門が反対を無視して評価要領を変更したのは「医学研究の理解度、練度が本当に低かった」から?

--

* 「今こそ振り返り、是正すべきだ」...他人事のような...

--

### 「JAXA事件の特徴: 組織的な研究倫理審査の軽視」という認識は当事者にない

---
class: inverse

[JAXA経営陣の説明](https://scienceportal.jst.go.jp/newsflash/20221128_n02/)

> _研究費は文科省科学研究費補助金（科研費）とJAXA予算を合わせ、約1億9000万円。佐々木理事はこの研究は.lightblue[成果として発表しておらず、特定不正行為には当たらない]とした。その上で「.lightblue[研究計画が稚拙で、十分な科学的合理性がなかった。]JAXAに医学研究の難しさの知見がなく、研究チームの人数が少なく体制を整えず、無理になってしまった。実験データが個人情報のためアクセスできる人を限ってしまい、周りが十分にチェックできなかった。国民の負託に応えられず、実験協力者の善意も裏切り深くお詫びする。一丸となって再発防止に取り組む」とした。_

--

* 成果公表しなければFFPではない、は誤解です。学会でも報告済み。

> _この指針は、我が国の研究者等 により実施され、又は日本国内において実施される人を対象とする生命科学・医学系研究 を対象とする。_ [厚労省「指針」第3「適用範囲」1「適用される研究」](https://www.mhlw.go.jp/content/001077424.pdf#page=10)

--

* 科学的意義の低い研究と分科会が判断

> _最も問題なのは、科学的合理性のある研究計画が検討されておらず、かつ科学的に適切な方法で遂行されなかったことであり、またそのように努めなかったこと_ ([調査報告書](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=7))

--

### 意義の低い研究の実施=研究倫理(Emanuel et al. 2, 3, 6)の逸脱

--

### そう指摘をする分科会を沈黙させた組織運営
---
class: inverse
## JAXA事件の背景([報告書, p.30](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=30))

### 組織として閉鎖環境実験成果を出すことを目指す

--

* 1985年から莫大な国費を用いてISS計画を推進  
* ISSでの成果の創出が有人部門にとっての優先事項となっていた  
* 2015年時点で、5年後の目指す成果の出口は「地上の閉鎖環境等における実証とISSにおける一部の検証を終え、火星に向けた研究を開始」すること  
* 2015年1月の第45回分科会(不承認)、第46回分科会(不承認)、同年9月第47回分科会(承認)、倫理審査委員会同年12月の第75回倫理審査委員会、2016年1月第76回倫理審査委員会(承認)

---
class: inverse

### 2回も科学的意義で落第点

--

### なのに無理やり実施: 経営者もそうだが、提案者も相当

>_倫理審査委員会や分科会での.orange[指摘事項に十分応えることなく研究を進めようとした研究チームの不誠実な態度]、それを容認してしまう倫理意識の欠如も背後要因となった。_ [調査報告書(p.19)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=19)

> _倫理審査委員会に提出した研究等実施計画書（第 3 回閉鎖試験）に「分科会において、妥当な計画であるとの評価を受けた」と.orange[虚偽の事実が記載]されていた_ [同(p.22)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=22)

--

### 「[専門家に任せる信頼の気持ちが勝ってしまった](https://scienceportal.jst.go.jp/newsflash/20230113_n01/#:~:text=%E5%B0%82%E9%96%80%E5%AE%B6%E3%81%AB%E4%BB%BB%E3%81%9B%E3%82%8B%E4%BF%A1%E9%A0%BC%E3%81%AE%E6%B0%97%E6%8C%81%E3%81%A1%E3%81%8C%E5%8B%9D%E3%81%A3%E3%81%A6%E3%81%97%E3%81%BE%E3%81%A3%E3%81%9F%E3%80%82)」ギフト・オーサーシップではないが、共著者なのに分担を超えて確認せず

--

### 研究体制の不備: 能力不足ならシニア研究者とペアにすべき

> _研究者としての.orange[適性を見極めず医師であることを理由に研究計画の策定から実行まで]を実施させた_ [同(p.24)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=24)
---
class: inverse

## JAXA当事者はなぜFF(P)やRE違反をしたのか?

--

### 非研究者なので研究倫理について考えていなかった

--

### 経営者が非研究者(研究能力の乏しい人)に多額予算の研究を実施させ、成果を出させようとした

--

### 分科会審査という安全装置が外された(邪魔者扱い)

--

### 人の配置、権限配分、予算、期間を経営者が考えていなかった

--

### 経営者が研究の見識を(経営者に相応しいほど)持っていなかった

---
class: inverse
## JAXA当事者はなぜFF(P)やRE違反をしたのか?

### Gino事件とは違い、個人的な金銭利得はほぼゼロ

--

### [組織の目標(成果を出す)と手段(人員配置、チェック機能、予算や期間)が整合的ではなかったために、末端の人員が帳尻あわせに研究不正をしたのでは?](https://scienceportal.jst.go.jp/newsflash/20230113_n01/#:~:text=%E9%99%90%E3%82%89%E3%82%8C%E3%81%9F%E4%BA%BA%E5%93%A1%E3%81%A8%E5%8E%B3%E3%81%97%E3%81%84%E3%82%B9%E3%82%B1%E3%82%B8%E3%83%A5%E3%83%BC%E3%83%AB%E3%81%AE%E4%B8%AD%E3%81%A7%E3%81%84%E3%81%8B%E3%81%AB%E7%A0%94%E7%A9%B6%E3%82%92%E9%80%B2%E3%82%81%E3%82%8B%E3%81%8B%E3%80%81%E5%BF%85%E6%AD%BB%E3%81%AB%E3%81%AA%E3%81%A3%E3%81%A6%E3%81%97%E3%81%BE%E3%81%A3%E3%81%9F%E3%80%82%E3%83%9E%E3%83%8D%E3%82%B8%E3%83%A1%E3%83%B3%E3%83%88%E3%81%AB%E9%81%A9%E5%88%87%E3%81%AB%E5%AE%9F%E6%83%85%E3%82%92%E4%BC%9D%E3%81%88%E3%81%A6%E8%AA%BF%E6%95%B4%E3%81%99%E3%81%B9%E3%81%8D%E3%81%A0%E3%81%A3%E3%81%9F%E3%80%82)

--

### サラリーマン研究者解釈=日本ぽい

--

* ダイハツ工業、豊田自動織機、三菱自動車工業、ビッグモーターの検査不正  
* 研究生産性を高めるのは良いことですが、資源投入を減らし過ぎると弊害があります  
* 研究予算などの入り口で搾り、人事評価などの出口で圧迫すると、研究倫理を逸脱する可能性を高めます

---
class: inverse, middle, center

## アジ研には研究倫理を尊重する文化があります

--

## 今年の研究成果に不安・不透明感を持つならば、早めに上長に相談しましょう

--

## 適度な競争や批判の圧力の下、若手を育てます

---
class: inverse, middle, center

# まとめ: 社会から託された信任を裏切らない

---
class: inverse
```{r results = "asis", echo = FALSE, warning = F}
PrintBibliography(myBib, .opts = list(check.entries = FALSE, sorting = "nyt",
  no.print.fields=c("eprint", "URL")),
  start = 1, end = 7)
```

### 2025-06-18 posts/ethics/KenkyuuRinri_slides.rmd
added DelabastitaRubens repo and ethics repo

---
title: ""
header_includes:
    - \usepackage{tikz}
    - \usepackage{adjustbox}
    - \usepackage{seiro_minimal}
    - \usepackage{xltxtra} 
    - \usepackage{zxjatype} 
    - \usepackage[ipa]{zxjafont} 
    - \usepackage{amsmath}
    - \usepackage{amssymb}
    - \usepackage{amsfonts}
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "default-fonts", "font.css", "xaringan-themer.css"]
    nature:
      highlightStyle: github
      countIncrementalSlides: false
      ratio: "16:9"
    citation_package: natbib
    self_contained: true
####    includes:
####      before_body: mathabbreviation.html
bibliography: c:/seiro/settings/Tex/seiro.bib
self_contained: true
# setwd(path <- "c:/seiro/docs/IDE/jinzai/kensyu/2024/"); rmarkdown::render("KenkyuuRinri_Slides.Rmd")
#### renderthis::to_pdf("docs/GNR_slides.Rmd")
#### git init
#### git add /docs    ## this adds a folder to track
#### git commit -am "initial commit"
#### git config --global user.name "SeiroIto"
#### git remote add origin https://github.com/SeiroIto/MinWageMarketPower.git  ## specify remote
#### git push -f origin main  ## push files to remote
---
class: inverse
```{r xaringan setup, include=FALSE, warning=FALSE}
xaringanExtra::use_panelset()
library(xaringanthemer)
####style_mono_dark(base_color = "#cbf7ed")
####style_mono_light(base_color = "#23395b")
style_solarized_light(
  text_color = "black", 
  title_slide_background_color = "darkblue",
  link_color = "green")
#### xaringancolor setup
library(xaringancolor)
xaringancolor::setup_colors(
  red ="red",
  green = "green",
  blue = "blue"
)
library(RefManageR)
```
```{r xaringan bib, include=FALSE, cache=T}
#### bibliography setup
BibOptions(check.entries = FALSE, style = "markdown", cite.style = "authoryear",
  bib.style = "authoryear", dashed = T, hyperlink = "to.doc")
##### (old) Uncomment below for the 1st run of each R session. 2nd run will get cached results.
##### (new) Delete cache if you add an entry to seiro.bib.
myBib <- ReadBib("c:/seiro/settings/TeX/seiro.bib")
```
<div style = "position:fixed; visibility: hidden">
$$\require{color}\definecolor{red}{rgb}{1, 0, 0}$$
$$\require{color}\definecolor{green}{rgb}{0, 1, 0}$$
$$\require{color}\definecolor{blue}{rgb}{0, 0, 1}$$
</div>
<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      red: ["{\\color{red}{#1}}", 1],
      green: ["{\\color{green}{#1}}", 1],
      blue: ["{\\color{blue}{#1}}", 1]
    },
    loader: {load: ['[tex]/color']},
    tex: {packages: {'[+]': ['color']}}
  }
});
</script>
<style>
.red {color: #FF0000;}
.green {color: #00FF00;}
.blue {color: #0000FF;}
</style>
```{r flair_color, echo=FALSE}
library(flair)
red <- "#FF0000"
green <- "#00FF00"
blue <- "#0000FF"
```
```{css echo=FALSE}
.highlight-last-item > ul > li,
.highlight-last-item > ol > li {
  opacity: 0.5;
}
.highlight-last-item > ul > li:last-of-type,
.highlight-last-item > ol > li:last-of-type {
  opacity: 1;
}
.inverse {
  background-color: #272822;
  color: #d6d6d6;
  text-shadow: 0 0 20px #333;
}
mark.red {
    color:#ff0000;
    background: none;
}
mark.blue {
    color:#0000A0;
    background: none;
}
.my-style {
  font-weight: bold;
  font-style: italic;
  font-size: 1.5em;
  color: red;
}
blockquote {
  border-left: .2px solid #275d38;
  margin: -5px 80px -5px 20px;
  padding-top: -0.5px;
  padding-bottom: -0.5px;
  line-height: 1.35em;
}
.pull-left60 {
  float: left;
  width: 57%;
  padding-right: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-right40 {
  float: right;
  width: 37%;
  padding-left: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-left70 {
  float: left;
  width: 67%;
  padding-right: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-right30 {
  float: right;
  width: 27%;
  padding-left: 2px
  margin-top: -1em;
  margin-bottom: 0em;
}
.pull-left50 {
  float: left;
  width: 50%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: -2em;
  margin-bottom: -1em;
}
.pull-right50 {
  float: right;
  width: 50%;
  padding-right: 0px;
  padding-left: 0px;
  margin-top: -2em;
  margin-bottom: -1em;
}
.center2 {
  margin: 0;
  position: absolute;
  top: 50%;
  left: 50%;
  -ms-transform: translate(-50%, -50%);
  transform: translate(-50%, -50%);
}
```

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      E: "{\\Large\\varepsilon}",
      bfx: "{\\mathbf{x}}",
      bfX: "{\\mathbf{X}}",
      bfalpha: "{\\boldsymbol{\\alpha}}",
      bfbeta: "{\\boldsymbol{\\beta}}",
      st: "{\\mbox{s.t.}}"
    }
  }
});
</script>
Any japanese?
---
class: inverse, middle, center
# 研究倫理

## 2024年度新入職員研修

<br>
### 開発研究センター 伊藤成朗
### `r format(Sys.time(), '%Y年%m月%d日 %R')`
---
class: inverse, middle, center

Any japanese?
---
class: inverse, middle, center
## 研究コミュニティに身を置く者が持つべき心構えをお話します。
---
class: inverse

極言すると

.pull-left70[
研究倫理とは

研究で対象社会に迷惑を与えないという原則

研究公正とは  
研究でごまかしをしないという原則
]

.pull-right30[
最低限の原則

それでも足をすくわれる研究者はいる
]
<br>
どう判断すれば分からなくなったとき、大原則に戻ると考えやすくなる

--

研究は社会と契約を結んで実施している

--

社会から託された信任を裏切ってはいけない

`r Citet(myBib, "GNR2020")`
---

![:scale 50%](REAndRI.jpg)

---
class: center
name: 
background-image: url("REAndRI.jpg")
background-size: 100%

---

> the conditions which a value added production function can straightforwardly be derived from a more primitive gross output production function are quite restrictive.

<p style="margin-bottom:2cm;">.pull-left[] .pull-right[`r Citet(myBib, "DeLoeckerSyverson2021", after = ", 5.6.2")`]</p>

---
class: inverse
```{r bib, include=FALSE}
#### create a bib file for the R packages used in this document
knitr::write_bib(x = "rmarkdown", file = 'KenkyuuRinri.bib')
```
```{r results = "asis", echo = FALSE}
PrintBibliography(myBib, .opts = list(check.entries = F, sorting = "nyt",
  no.print.fields=c("eprint", "URL")),
  start = 1, end = 8)
```
---
class: inverse
```{css, eval = F, echo = F}
.remark-slide-content {
  font-size:          12px;
  line-height:        1.2em;
}
```
```{r results = "asis", echo = FALSE, warning = F}
PrintBibliography(myBib, .opts = list(check.entries = FALSE, sorting = "nyt",
  no.print.fields=c("eprint", "URL")),
  start = 9, end = 12)
```

### 2025-06-18 posts/ethics/testslides.rmd
added DelabastitaRubens repo and ethics repo

--- 2025-06-18 posts/ethics/KenkyuRinriSlides.rmd
@@
     - \usepackage{RefManageR}
-# setwd(path <- "c:/seiro/docs/IDE/jinzai/kensyu/2024/"); rmarkdown::render("KenkyuRinriSlides.Rmd")
+# setwd(path <- "c:/seiro/docs/IDE/jinzai/kensyu/2024/"); rmarkdown::render("testslides.Rmd")
 # git status
@@
 ### 伊藤成朗
+### `r format(Sys.time(), '%Y年%m月%d日 %R')`
 ### 2024年4月17日(水)
@@
 
-### https://seiroito.github.io/Miscellaneous/2024/KenkyuRinriSlides.html
+### https://seiroito.github.io/Miscellaneous/2024/testslides.html
 
@@
 
-## 崇高な研究哲学論ではなく、実際的な自衛の話
+## 崇高な精神論ではなく、自衛の話
 ---
@@
 
-.orange[例1:] Dutch National Survey on Research Integrity [`r Citep(myBib, "Gopalakrishnaetal2022")`](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0263023&fbclid=IwAR17MW5qz2wpZXQz6crAQVrsDbGsx0DKeE5ZIuvIywF137rYlf_5hXz1Alw):
+例1: Dutch National Survey on Research Integrity [`r Citep(myBib, "Gopalakrishnaetal2022")`](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0263023&fbclid=IwAR17MW5qz2wpZXQz6crAQVrsDbGsx0DKeE5ZIuvIywF137rYlf_5hXz1Alw):
 
@@
 
-.orange[例2:] The International Research Integrity Survey `r Citep(myBib, "Allumetal2022")`:
+例2: The International Research Integrity Survey `r Citep(myBib, "Allumetal2022")`:
 
@@
 
-.orange[例3:] [Data Collada (psychologist blog)](https://datacolada.org/) vs. Francesca Gino (HBS prof), 2021-
+例3: [Data Collada (psychologist blog)](https://datacolada.org/) vs. Francesca Gino (HBS prof), 2021-
 
@@
 
-.orange[例4:] [国立医薬品食品衛生研究所職員による研究活動上の不正行為及びISO/IEC 17025認定審査対応における不適切行為に関する調査結果の公表](https://www.nihs.go.jp/oshirasejoho/Press-Release/20231226/index.html) 2023年12月26日、食品衛生部長によるFabrication, Falsification認定
+例4: [国立医薬品食品衛生研究所職員による研究活動上の不正行為及びISO/IEC 17025認定審査対応における不適切行為に関する調査結果の公表](https://www.nihs.go.jp/oshirasejoho/Press-Release/20231226/index.html) 2023年12月26日、食品衛生部長によるFabrication, Falsification認定
 ---
@@
 ]
----
-class: inverse
-
-## なぜQRPをするのか
-
---
-
-## 成果を出したい、でも、今のままだと成果にならないと考えている
-
---
-.pull-left50[
-### 内容が不十分な場合
-
-* 結果がはっきりしない  
-* 2つの結果が矛盾する  
-* 何を議論すべきかもよく分からない  
-]
-
---
-
-.pull-right50[
-<p style="margin-top:.5cm;">
-</p>
-
-* 不適切な統計手法  
-* データの未保存や再現性の欠如  
-* 曖昧な記述と結論...QRP?  
-   * 結論は曖昧です、と明確に書けばok  
-   * 単に記述を不明瞭にするのはごまかし=QRP
-
-* 過程(データとコード)を公開  
-* Pre analysis planを公開・登録  
-]
----
-class: inverse
-
-## なぜQRPをするのか
-
---
-
-<p style="margin-top:-1.5cm;">
-</p>
-
-.pull-left70[
-### 自信、発信力が無い場合
-
-* ビッグ・ネームの推薦を利用したい  
-]
-
---
-
-.pull-right30[
-
-<p style="margin-top:1.5cm;">
-</p>
-
-* ギフト・オーサーシップ  
-* 反例: ネットワーク理論: Erdős-Rényiモデル. [Erdős](https://en.wikipedia.org/wiki/Paul_Erd%C5%91s)は1525論文掲載  
-]
-
---
-
-<p style="margin-top:-2.5cm;">
-</p>
-
-.pull-left100[
-]
-.pull-left70[
-### 時間がない場合
-
-* 今までの研究に手を加えれば新しいものに?  
-]
-
---
-
-.pull-right30[
-
-<p style="margin-top:1.5cm;">
-</p>
-
-* 多重投稿、多重出版  
-* 機械学習で簡単に判明  
-]
-
-<p style="margin-bottom:-1.5cm;">
-</p>
-
---
-
-<p style="margin-top:-1.5cm;">
-</p>
-.pull-left100[
-]
-.pull-left70[
-### スポンサーと利益相反するが、研究は健全で余計な詮索をされたくないと考える場合
-
-* 利益相反があることを表明しない  
-]
-
---
-
-.pull-right30[
-
-<p style="margin-top:1.5cm;">
-</p>
-
-* 利益相反
-
-<p style="margin-top:0.5cm;">
-</p>
-
-* 簡単に判明  
-]
-
 ---
@@
 class: inverse, center, middle
-## 経営者が研究倫理審査や研究能力を軽視すると、配下の研究者が研究倫理・研究公正から逸脱するのを防げない
+## 経営者が研究倫理審査や研究能力を軽視すると、研究倫理・研究公正違反を防げない
 ---
@@
 
-* 報告書執筆者は第三者委員会ではありません
+* 第三者委員会ではありません
 

### 2025-06-18 posts/ethics/KenkyuRinri2025_slides.qmd
added ethics pages to readme.md
dropped rmd files

@@
 
+## {background-color="#010638"}
+
+### 蔓延度は不明ですが、結構深刻では?
+
+<Orange>例5:</Orange> ミクロ開発経済学研究者[Asad Islam](https://users.monash.edu.au/~asaduli/index.html)の複数論文で[データ不正(FF)の疑い](https://a-ortmann.medium.com/the-asad-islam-gdri-saga-f96b567c93d6)
+
+[Institute For Replicationによる再現性分析](https://i4replication.org/gdri.html)で指摘
+
+:::::::::: {.columns}
+::::: {.column width = "40%"}
+::: {style="text-align: center;"}
+![](./IslamResearchMisconduct.jpg){width="90%"}
+:::
+:::::
+
+::::: {.column width = "60%"}
+Retracted
+
+* Journal of European Economics Association
+* Economic Development and Cultural Change
+* European Economic Review
+* Economic Journal
+* Journal of Public Economics (under review)
+* (American Journal of Economics: Micro)
+* (Review of Economics and Statistics)
+:::::
+::::::::::
+
+----
+
 ## なぜFFPをするのか{background-color="#010638"}
@@
 ::::::::::
+
+----
+
+## 出版社もFFP, QRP通報窓口を設置{background-color="#010638"}
+
+[COPE](https://publicationethics.org/guidance/case/ethics-approval-and-consent)
 
@@
 
+:::: {.fragment}
 ### 意義の低い研究の実施=研究倫理(Emanuel et al. 2, 3, 6)の逸脱
@@
 ### そう指摘をする分科会を沈黙させた組織運営
+::::
 
@@
 
+:::: {.fragment}
+### その担当者に非研究職の宇宙飛行士を任命、研究の実施と成果を求める
+
+* トップダウンのプロジェクト
+::::
+
 ----
@@
 
+:::: {.fragment}
 ### 「[専門家に任せる信頼の気持ちが勝ってしまった](https://scienceportal.jst.go.jp/newsflash/20230113_n01/#:~:text=%E5%B0%82%E9%96%80%E5%AE%B6%E3%81%AB%E4%BB%BB%E3%81%9B%E3%82%8B%E4%BF%A1%E9%A0%BC%E3%81%AE%E6%B0%97%E6%8C%81%E3%81%A1%E3%81%8C%E5%8B%9D%E3%81%A3%E3%81%A6%E3%81%97%E3%81%BE%E3%81%A3%E3%81%9F%E3%80%82)」ギフト・オーサーシップではないが、共著者なのに分担を超えて確認せず
-
+::::
+
+:::: {.fragment}
 ### 研究体制の不備: 能力不足ならシニア研究者とペアにすべき
@@
 > _研究者としての<Orange>適性を見極めず医師であることを理由に研究計画の策定から実行まで</Orange>を実施させた_ [同(p.24)](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=24)
+::::
 
@@
 
+:::: {.incremental}
 * 非研究者なので研究倫理について考えていなかった
@@
 * 経営者が研究の見識を(経営者に相応しいほど)持っていなかった
+::::
 

### 2025-06-19 posts/ethics/KenkyuRinri2025_slides.qmd
dropped xaringanExtra library

@@
 ---
-```{r xaringan-panelset, echo=FALSE, warning = F}
-library(xaringanExtra)
-use_panelset()
+```{r read RefManageR library, echo=FALSE, warning = F}
 library(RefManageR)
@@
 
+## {background-color="#010638"}
+
 ::: {.center2}
@@
 
-----
-
 # 研究倫理{background-color="#010638"}
 
-----
-
 ## {background-color="#010638"}
@@
 :::
-
-----
 
@@
 :::
-
-----
-
-## 研究公正 research integrity {background-color="DarkBlue"  background-image="./JAXASumimasen.jpg" background-size="100%"}
-
-----
-
-# 研究公正 research integrity {background-color="DarkBlue"  background-image="./JAXASumimasen.jpg" background-size="100%"}
-
-----
-
-## 研究公正 research integrity {background-color="black" background-image="./JAXASumimasen.jpg" background-size="100px" background-repeat="repeat"}
 
@@
 :::::::::: {.columns}
-::::: {.column width = "65%"}
-::: {.fragment}
+::::: {.column width = "60%"}
 ::: {style="font-size:130%"}
+:::: {.fragment}
 <White> <LightBlue>F</LightBlue><Orange>abrication ねつ造</Orange>「存在しないデータや研究成果を作成すること。」  
+::::
+
+:::: {.fragment}
 <LightBlue>F</LightBlue><Orange>alsification 改ざん</Orange>「操作」によって「データ」や「成果などを真正でないものに加工すること」。  
+::::
+
+:::: {.fragment}
 <LightBlue>P</LightBlue><Orange>lagerism 剽窃、盗用</Orange> 他の研究者の「アイディア」「方法」「データ」「結果」を「了解または適切な表示なく流用すること」。</White>
-:::
-:::
-:::::
-
-::::: {.column width = "35%"}
+::::
+:::
+:::::
+
+::::: {.column width = "40%"}
 ::: {style="font-size:90%"}
 ::: {.incremental}
-* やると研究者人生が終わります。  
+* やると研究者人生がほぼ終わります。  
 * 理由: 研究者社会は相互信頼で成立しています。ウソをつかないことを前提に評価がされて、知識が共有されます。裏切ると、研究者社会全体が資源を浪費するので、厳罰処分されます。不正が露見するまで、他の研究者は存在しない現象を追求して苦しみます。  
    * STAP論文: 追試資源や継続研究資源の浪費、周辺研究まで疑いの目
+   * 不正をする人は自分の利得だけで社会への影響をほぼ無視
 :::
@@
 :::::::::: {.columns}
-::::: {.column width = "55%"}
+::::: {.column width = "40%"}
 ::: {.incremental}
@@
 
-::::: {.column width = "45%"}
+::::: {.column width = "2%"}
+:::::
+
+::::: {.column width = "58%"}
 ::: {.incremental}
@@
 * $P$-hacking: E.g.,  望んだ結果が出るまで推計を試行し、その結果のみ提示すること  
-* GAは贈賄双方に便益あり、[なくなりにくい、でも、著者責任も付いてくる](https://www.bmj.com/content/309/6967/1456#:~:text=Perhaps%20this%20scandal%20will%20finally%20undermine%20gift%20authorship.%20At%20the%20very%20least%20it%20should%20make%20researchers%20think%20hard%20about%20the%20responsibilities%20that%20come%20with%20putting%20their%20names%20on%20papers.)
+* GAは贈賄双方に便益あり、[なくなりにくい、でも、著者責任も付いてくる](https://www.bmj.com/content/309/6967/1456#:~:text=Perhaps%20this%20scandal%20will%20finally%20undermine%20gift%20authorship.%20At%20the%20very%20least%20it%20should%20make%20researchers%20think%20hard%20about%20the%20responsibilities%20that%20come%20with%20putting%20their%20names%20on%20papers.)&larr; Asad IsamのFF疑惑
 * 刊行された論文を翻訳して投稿してはいけません ([International Committee of Medical Journal Editors](https://www.icmje.org/recommendations/browse/publishing-and-editorial-issues/overlapping-publications.html#:~:text=Authors%20should%20not%20submit%20the%20same%20manuscript%2C%20in%20the%20same%20or%20different%20languages%2C%20simultaneously%20to%20more%20than%20one%20journal.))、翻訳であることを明示しても危ういので、翻訳論文はワーキング・ペーパーに留めましょう
@@
 ::::: {.column width = "55%"}
-<White>
+:::: {.fragment}
 FF(P) 8.3% engaged in any FF(P).
-</White>
-
-<White>
+
 QRP 51.3% engaged frequently in at least 1 QRP.  
-</White>
+::::
 :::::
@@
 ::::: {.column width = "45%"}
-<White>
+:::: {.fragment}
 Beh&Soc.Sci 5.7%
-</White>
-
-<White>
+
 Beh&Soc.Sci 50.2%
-</White>
+::::
 :::::
@@
 
+:::: {.fragment}
 ::: {style="font-size:80%"}
@@
 :::
-
+::::
+
+:::: {.fragment}
 <Orange>例4:</Orange> [国立医薬品食品衛生研究所職員による研究活動上の不正行為及びISO/IEC 17025認定審査対応における不適切行為に関する調査結果の公表](https://www.nihs.go.jp/oshirasejoho/Press-Release/20231226/index.html)
@@
 :::
+::::
 
@@
 
+:::: {.fragment}
 :::::::::: {.columns}
@@
 * Journal of Public Economics (under review)
-* (American Journal of Economics: Micro)
+* (American Economic Journl: Micro)
 * (Review of Economics and Statistics)
@@
 :::::::::: 
+::::
 
@@
 
-結構ばれる: Ginoは追試、食品衛生部長は内部通報
+結構ばれる: Ginoは追試、食品衛生部長は内部通報、Islamは経済学界の再現可能性指向
 
@@
 
-<White>Fabrication, falsification、plagerismのリスク</White>
+Fabrication, falsification、plagerism回避のルール
 
@@
 * 非研究者なので研究倫理について考えていなかった
-* 経営者が非研究者(研究能力の乏しい人)に多額予算の研究を実施させ、成果を出させようとした
+* 経営者が非研究者(研究能力の乏しい人)に多額予算の研究を実施させ、成果を出させようとした(トップダウン)
 * 分科会審査という安全装置が外された(邪魔者扱い)

### 2025-06-19 posts/ethics/KenkyuRinri2025_slides.qmd
dropped xaringanExtra library

@@
   color:white
+}
+.title{
+  background-color=#010638;
 }
@@
 
-## JAXA事件の背景([報告書, p.30](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=30)){background-color="#010638"}
+## JAXA事件の組織的背景([報告書, p.30](https://www.jaxa.jp/press/2022/11/20221125-2.pdf#page=30)){background-color="#010638"}
 
@@
 
+:::: {.fragment}
 * 1985年から莫大な国費を用いてISS計画を推進  
@@
 * 2015年1月の第45回分科会(不承認)、第46回分科会(不承認)、同年9月第47回分科会(承認)、倫理審査委員会同年12月の第75回倫理審査委員会、2016年1月第76回倫理審査委員会(承認)  
-
-:::: {.fragment}
-### その担当者に非研究職の宇宙飛行士を任命、研究の実施と成果を求める
+::::
+
+:::: {.fragment}
+### その担当者に非研究職(医師、宇宙飛行士)を任命、研究の実施と成果を求める
 
 * トップダウンのプロジェクト
-::::
-
-----
-
-## {background-color="#010638"}
+* 無理な業務命令なので末端が無理をしている
+::::
+
+----
+
+## JAXA事件の組織的背景 {background-color="#010638"}
 
@@
 
-### なのに無理やり実施: 経営者もそうだが、提案者も相当
+### なのに無理やり実施: 経営者が無理解、提案者は忖度?恭順?
 
@@
 
-:::: {.fragment}
-### 「[専門家に任せる信頼の気持ちが勝ってしまった](https://scienceportal.jst.go.jp/newsflash/20230113_n01/#:~:text=%E5%B0%82%E9%96%80%E5%AE%B6%E3%81%AB%E4%BB%BB%E3%81%9B%E3%82%8B%E4%BF%A1%E9%A0%BC%E3%81%AE%E6%B0%97%E6%8C%81%E3%81%A1%E3%81%8C%E5%8B%9D%E3%81%A3%E3%81%A6%E3%81%97%E3%81%BE%E3%81%A3%E3%81%9F%E3%80%82)」ギフト・オーサーシップではないが、共著者なのに分担を超えて確認せず
+----
+
+## JAXA事件の組織的背景 {background-color="#010638"}
+
+:::: {.fragment}
+### FFについて:「[専門家に任せる信頼の気持ちが勝ってしまった](https://scienceportal.jst.go.jp/newsflash/20230113_n01/#:~:text=%E5%B0%82%E9%96%80%E5%AE%B6%E3%81%AB%E4%BB%BB%E3%81%9B%E3%82%8B%E4%BF%A1%E9%A0%BC%E3%81%AE%E6%B0%97%E6%8C%81%E3%81%A1%E3%81%8C%E5%8B%9D%E3%81%A3%E3%81%A6%E3%81%97%E3%81%BE%E3%81%A3%E3%81%9F%E3%80%82)」ギフト・オーサーシップではないが、共著者なのに分担を超えて確認せず
 ::::
@@
 
+:::: {.incremental}
 * Gino事件とは違い、個人的な金銭利得はほぼゼロ  
@@
    * 研究予算などの入り口で搾り、人事評価などの出口で圧迫すると、研究倫理を逸脱する可能性を高めます
-
-----
-
-## アジ研には研究倫理を尊重する文化があります{background-color="#010638"}
-
-## 今年の研究成果に不安・不透明感を持つならば、早めに上長に相談しましょう{background-color="#010638"}
-
-## 適度な競争や批判の圧力の下、若手を育てます{background-color="#010638"}
-
-----
+::::
+
+# アジ研には研究倫理を尊重する文化があります{background-color="#010638"}
+
+# 今年の研究成果に不安・不透明感を持つならば、早めに上長に相談しましょう{background-color="#010638"}
+
+# 適度な競争や批判の圧力の下、若手を育てます{background-color="#010638"}
 
@@
 
+----
+
+## {background-color="#010638"}
+
 References

### 2025-06-19 posts/ethics/KenkyuRinri2025_slides.qmd
dropped xaringanExtra library

@@
     logo: images/quarto.png
-    footer: <http://seiroito.github.io/QuartoFiles/Horton/Horton2025_slides.html>
+    footer: <http://seiroito.github.io/QuartoFiles/ethics/KenkyuRinri2025_slides.html>
     theme: sky
@@
 
+:::: {.fragment}
 See a list of QRPs used in Dutch national survey (DNSRI) summarised [here](https://htmlpreview.github.io/?https://github.com/SeiroIto/PaperNotes/blob/main/QRPsInDutchNationalSurvey_Tufte.html)
+::::
 
@@
 ::: {style="font-size:80%"}
-* [Allegation](https://datacolada.org/109): Data tampering in several of her work on dishonesty (incl. with Dan Ariely!)  
+* [Allegation](https://datacolada.org/109): Data tampering (FF) in several of her work on dishonesty (incl. with Dan Ariely!)  
 * HBS: Put Gino on an unpaid admin leave, hired file forensic investigators, sent retraction requests to journals  
@@
 
+:::: {.incremental}
 * 日本の文系(<2000万円)だとFFPは割に合いません  
 * 国立医薬品食品衛生研究所所長で1900万円くらい
-
+::::
+
+:::: {.fragment}
 結構ばれる: Ginoは追試、食品衛生部長は内部通報、Islamは経済学界の再現可能性指向  
+::::
 
@@
 
+:::: {.fragment}
 貧乏くじ
+::::
 
@@
 
+:::: {.fragment}
 オッズをよく知らないまま、自ら進んで引くのは賢くない
@@
 ([WGBH news](https://www.wgbh.org/news/education-news/2025-05-25/in-extremely-rare-move-harvard-revokes-tenure-and-cuts-ties-with-star-business-professor))
+
+::::
 
